\subsection{Case III}
\label{sec:molecular-algebraic-induction-caseIII}

Case III realizes the reduction move left by Cases~I and~II: splitting off a
degree-$2$ vertex $v$ of a simple minimal $0$-dof-graph $G$ that has no proper
rigid subgraph (KT Lemmas~6.10/6.13, \cref{lem:case-III}). The placement of
\cref{lem:case-II-realization-placement} realizes $G$ at rank $D(|V| - 1) - 1$,
one row short of full rank, and Case III supplies the missing row. That row is
KT Claim~6.11 (eq.\ (6.23)): in the split-off graph $G_v^{ab}$, one rigidity row
at the deleted edge $ab$ is a linear combination of the others, so it is
redundant. This subsection proves Claim~6.11; the candidate-completion
(\cref{sec:molecular-algebraic-induction-candidate-completion}) then rewrites
the redundant row into the missing full-rank row, and Claim~6.12
(\cref{sec:molecular-algebraic-induction-claim612}) discharges the one remaining
conditional.

The realizations throughout Case~III are \emph{free-normal} panel frameworks: a
panel-hinge framework (\cref{def:panel-hinge-framework}) whose body normals are
read directly off a seed $q$, with $n_v = q(v, \cdot) \in \R^{k+2}$, together
with an endpoint selector $\mathit{ends}$ recording each edge's two endpoints.

Claim~6.11 is proved in three steps, in dependency order. The first is pure
matroid theory in $M(\tilde G)$: the matroid-base form of KT Lemma~4.3(ii) at
$k = 0$.

\begin{lemma}[Matroid-base form of KT Lemma~4.3(ii), forward half; KT Claim 6.11, step 1]
  \label{lem:case-III-claim-6-11-base}
  \lean{Graph.splitOff_exists_base_inter_fiber_lt}
  \leanok
  \uses{def:k-dof, def:matroid-MG, thm:def-eq-corank, lem:forest-surgery-count, lem:splitoff-deficiency}
  Let $G$ be a $k$-dof-graph with a degree-$2$ vertex $v$ with distinct
  neighbours $a, b$, incident to exactly the two edges $e_a \ne e_b$, and let
  $e_0 \notin E(G)$ be a fresh edge. Suppose the splitting-off $G_v^{ab}$ is
  again a $k$-dof-graph. Then $M(\widetilde{G_v^{ab}})$ has a base $B'$ with
  \[
    \bigl|\widetilde{ab} \cap B'\bigr| \;<\; D - 1,
  \]
  where $\widetilde{ab}$ denotes the $D - 1$ copies of the split-off edge $ab$
  in $\widetilde{G_v^{ab}}$; that is, $\widetilde{ab} \not\subseteq B'$, so one
  copy of $ab$ is redundant.
\end{lemma}
\begin{proof}
  \leanok
  Reroute a balanced $D$-forest packing of a base $B$ of $M(\tilde G)$ across
  $v$ (\cref{lem:forest-surgery-count}): this yields an
  $M(\widetilde{G_v^{ab}})$-independent set $I'$ with $|I'| = |B| - D$ and
  $|\widetilde{ab} \cap I'| < D - 1$ (the two conclusions of KT Lemma~4.1). Both
  graphs are $k$-dof, so the $\operatorname{def} = \operatorname{corank}$
  identity (\cref{thm:def-eq-corank}) gives $\operatorname{rank}
  M(\widetilde{G_v^{ab}}) = D(|V \setminus \{v\}| - 1) - k = |B| - D = |I'|$. An
  independent set of full rank is a base, so $I'$ is the required base.
\end{proof}

The second step is KT's nested induction. The vertex-removal $G_v := G - v =
G_v^{ab} - ab$ is itself a \emph{minimal} $k'$-dof-graph for $k' :=
\operatorname{def}(\tilde G_v)$, with $k' \le D - 2$. Its minimality and the
deficiency bound are pure matroid theory in $M(\tilde G)$; the analytic step
they set up --- applying the geometric induction hypothesis to $G_v$ at the
inductively-fixed seed (KT eq.\ (6.22)) --- is
\cref{lem:case-III-nested-rank-lower} below.

\begin{lemma}[$G - v$ is minimal $k'$-dof with $k' \le D - 2$; KT Lemma~4.3(ii)/nested IH, Claim~6.11 prerequisite]
  \label{lem:case-III-gap3-minimalKDof}
  \lean{Graph.splitOff_removeVertex_minimalKDof}
  \leanok
  \uses{def:k-dof, def:graph-operations, lem:subgraph-minimality, thm:def-eq-corank,
        lem:splitoff-deficiency, lem:case-III-claim-6-11-base}
  Let $G$ be a minimal $0$-dof-graph with a degree-2 vertex $v$ (distinct
  neighbours $a, b$, incident edges exactly $e_a \ne e_b$, $e_0 \notin E(G)$
  fresh). Then the vertex-removal $G_v := G - v$ is a minimal
  $k'$-dof-graph for $k' = \operatorname{def}(\tilde G_v)$, with
  \[
    0 \;\le\; k' \;\le\; D - 2.
  \]
\end{lemma}
\begin{proof}
  \leanok
  Minimality transports along $G_v \le G$ by \cref{lem:subgraph-minimality}: a
  subgraph of a minimal $k$-dof-graph is minimal at its own deficiency. For the
  bound, take the base $B'$ of $M(\widetilde{G_v^{ab}})$ with
  $h := |\widetilde{ab} \cap B'| < D - 1$
  (\cref{lem:case-III-claim-6-11-base}). Its surviving part $B' \setminus
  \widetilde{ab}$ lands in $E(\tilde G_v)$ and is independent in $M(\tilde G_v) =
  M(\widetilde{G_v^{ab}}) \restriction E(\tilde G_v)$ (matroid restriction along
  $\tilde G_v \le \widetilde{G_v^{ab}}$), so $\operatorname{rank} M(\tilde G_v)
  \ge |B'| - h$. At $k = 0$ the splitting off is $0$-dof
  (\cref{lem:splitoff-deficiency}), so $|B'| = D(|V \setminus \{v\}| - 1)$; the
  def\,$=$\,corank identity (\cref{thm:def-eq-corank}, same vertex set $V(G)
  \setminus \{v\}$) then gives $\operatorname{def}(\tilde G_v) = D(|V \setminus
  \{v\}| - 1) - \operatorname{rank} M(\tilde G_v) \le h < D - 1$, i.e.\
  $k' \le D - 2$. The lower bound $0 \le k'$ is the nonnegativity of deficiency.
\end{proof}

The third step is the analytic content of KT's footnote~6. The nested induction
fixes the realizing seed with algebraically-independent coordinates, so the
inductively-fixed seed --- not merely \emph{some} generic seed --- attains the
matroid-predicted rank of each subgraph. At the $0$-dof level this is KT
eq.~(6.18): the split-off $G_v^{ab}$ is $0$-dof, so its generic realization is
rigid, and the following lemma transfers that rigidity to the fixed seed.

\begin{lemma}[Seed-rank bridge: rigidity transfers to every algebraically-independent seed; KT footnote 6, eqs.\ (6.18)/(6.22)]
  \label{lem:case-III-seed-rank-bridge}
  \lean{CombinatorialRigidity.Molecular.PanelHingeFramework.isInfinitesimallyRigidOn_ofNormals_of_algebraicIndependent}
  \leanok
  \uses{lem:genericity-device, def:rank-hypothesis}
  Let $(G, q_0)$ be a free-normal panel framework with nondegenerate hinges
  ($C(e) \ne 0$), its endpoint selector recording every link of $G$. If it is
  infinitesimally rigid on $V(G)$ at some seed $q_0$, then the free-normal
  framework $(G, q)$ is infinitesimally rigid on $V(G)$ at every seed $q$ that
  is algebraically independent over $\Q$.
\end{lemma}
\begin{proof}
  \leanok
  Infinitesimal rigidity at $q_0$ gives a nonzero rank polynomial $Q$ in the
  seed coordinates whose non-roots witness a full-size independent subfamily of
  panel rows (the coordinatization of the genericity device,
  \cref{lem:genericity-device}). Its coefficients are the change-of-basis
  constants of the complement isomorphism, hence rational. A seed $q$
  algebraically independent over $\Q$ lies off the zero locus of every nonzero
  rational polynomial (KT footnote~6), so $Q(q) \ne 0$, and the genericity
  device reads infinitesimal rigidity at $q$ off that non-root.
\end{proof}

The seed-rank bridge transfers \emph{full} rigidity (the $0$-dof case,
eq.~(6.18)). KT eq.~(6.22) needs the $\operatorname{def} > 0$ counterpart: at
the same inductively-fixed seed, the nested subgraph $G_v$ (minimal $k'$-dof,
$k' = \operatorname{def}(\tilde G_v)$) attains rank $D(|V_v| - 1) - k'$,
equivalently $\dim Z(G_v, q) = D + k'$. This equality is pinned between two
bounds: the genericity-free lower bound $D + \operatorname{def} \le \dim Z$
(\cref{lem:trivial-motions-rank-bound}, the lower-bound half of KT
Proposition~1.1, \cref{prop:rigidity-matrix-prop11}) and the matching upper
bound proved next.

\begin{lemma}[Eq.~(6.22) rank upper bound at every algebraically-independent seed; KT footnote 6, eq.\ (6.22)]
  \label{lem:case-III-seed-rank-upper}
  \lean{CombinatorialRigidity.Molecular.PanelHingeFramework.finrank_infinitesimalMotions_le_of_rankPolynomial_algebraicIndependent}
  \leanok
  \uses{lem:case-III-seed-rank-bridge, thm:def-eq-corank}
  Let $Q$ be a nonzero rational polynomial (coefficients in the image of $\Q \to
  \R$) whose non-roots witness a linearly independent subfamily $s$ of the panel
  rows of the free-normal panel framework on $G$, the edge of each member of $s$
  linking in $G$. Then at any seed $q$ algebraically independent over $\Q$,
  \[
    \dim Z(G, q) \;\le\; D\,|\alpha| - |s|,
  \]
  where $|\alpha|$ is the number of bodies.
\end{lemma}
\begin{proof}
  \leanok
  A seed algebraically independent over $\Q$ is a non-root of the nonzero
  rational $Q$ (KT footnote~6, as in \cref{lem:case-III-seed-rank-bridge}), so
  the witnessed subfamily $s$ is linearly independent at $q$ itself. Each panel
  row of $s$ lies in the rigidity rows (its edge links in $G$), so $|s| \le \dim
  \operatorname{span}(\text{rigidity rows}) = D\,|\alpha| - \dim Z(G, q)$ by
  rank--nullity (the rigidity-row span and the null space are dual-annihilator
  complements). Rearranging gives the bound.
\end{proof}

The upper bound has a lower-bound companion at the nested subgraph, obtained as
in KT. The induction hypothesis (6.1) ranges over minimal $k$-dof-graphs for
\emph{all} $k$, so it applies to the auxiliary graph $G_v$ (minimal $k'$-dof)
inside the $k = 0$ case. The formalized induction runs on the same all-$k$
statement (\cref{thm:minimal-kdof-reduction-all-k}), so this nested application
is available and the bound follows from the induction hypothesis.

\begin{lemma}[Eq.~(6.22) rank lower bound at the nested $k'$-dof subgraph; KT eq.\ (6.22), nested hypothesis (6.1)]
  \label{lem:case-III-nested-rank-lower}
  \lean{CombinatorialRigidity.Molecular.PanelHingeFramework.case_III_nested_rank_lower_all_k,
        CombinatorialRigidity.Molecular.PanelHingeFramework.case_III_nested_rank_lower}
  \leanok
  \uses{def:rank-hypothesis, lem:case-III-gap3-minimalKDof}
  Let $G$ be a minimal $0$-dof-graph with a degree-$2$ vertex $v$ (neighbours
  $a, b$, fresh edge $e_0$), let $\mathit{ends}$ be an endpoint selector
  recording every link of the splitting-off $G_v^{ab}$, and let $q$ be a seed,
  algebraically independent over $\Q$, whose body normals are pairwise linearly
  independent. Then the rigidity rows of the free-normal panel framework on the
  vertex-removal $G_v = G - v$ at $(\mathit{ends}, q)$ satisfy
  \[
    D\,\bigl(|V(G_v^{ab})| - 1\bigr) - (D - 2)
    \;\le\; \dim \operatorname{span}\bigl(\text{rigidity rows of }(G_v, q)\bigr):
  \]
  the nested subgraph $G_v$ --- minimal $k'$-dof with $k' \le D - 2$
  (\cref{lem:case-III-gap3-minimalKDof}) --- attains at the inductively-fixed
  seed at least the rank $D(|V_v| - 1) - k'$ that KT's hypothesis (6.1)
  predicts for it.
\end{lemma}
\begin{proof}
  \leanok
  The all-$k$ induction hypothesis gives a generic full-rank realization of
  $G_v$ (minimal $k'$-dof, simple, nonempty, and strictly smaller than $G$).
  From it, a rank-polynomial extractor produces a nonzero rational polynomial
  $P$ in the seed coordinates with $N \le \dim \operatorname{span}(\text{rows at
  }q')$ for every $q'$ off its zero locus, where $N = D(|V_v| - 1) - k'$. The
  seed $q$ is algebraically independent over $\Q$ and $P$ is rational, so $q$ is
  not a root of $P$ (KT footnote~6); the bounds $k' \le D - 2$ and
  $|V(G_v^{ab})| = |V_v|$ close the arithmetic.
\end{proof}
\begin{fmlnote}
  The bound is formalized at general grade $k$, with the $d = 3$ case ($k = 2$)
  recorded as a separate specialization.
\end{fmlnote}

The two analytic halves now combine. The nested subgraph $G_v$ attains its
predicted rank at the inductively-fixed seed by pinning the equality between the
genericity-free lower bound and the matroid-predicted upper bound.

\begin{lemma}[Eq.~(6.22) rank attainment at the inductively-fixed seed; KT footnote 6, eq.\ (6.22)]
  \label{lem:case-III-rank-attainment}
  \lean{CombinatorialRigidity.Molecular.PanelHingeFramework.rankHypothesis_ofNormals_of_rankPolynomial_algebraicIndependent}
  \leanok
  \uses{lem:case-III-seed-rank-upper, lem:trivial-motions-rank-bound, thm:def-eq-corank}
  Let $(G, q)$ be the free-normal panel framework on a spanning graph $G$
  ($V(G) = \alpha$) with every hinge nondegenerate ($C(e) \ne 0$), and let the
  inductively-fixed seed $q$ be algebraically independent over $\Q$. Suppose a
  nonzero rational polynomial $Q$ witnesses a linearly independent subfamily $s$
  of the panel rows (edges linking in $G$) of the matroid-predicted full size
  \[
    |s| \;\ge\; D(|V| - 1) - \operatorname{def}(\tilde G).
  \]
  Then $G$ attains its predicted rank at $q$,
  \[
    \dim Z(G, q) \;=\; D + \operatorname{def}(\tilde G),
  \]
  the rank hypothesis (\cref{def:rank-hypothesis}) at $\operatorname{def}(\tilde G)$.
\end{lemma}
\begin{proof}
  \leanok
  The eq.~(6.22) upper bound (\cref{lem:case-III-seed-rank-upper}) gives $\dim
  Z(G, q) \le D|V| - |s|$ at $q$. Substituting the matroid-predicted size $|s|
  \ge D(|V| - 1) - \operatorname{def}$ (the matroid rank $\operatorname{rank}
  M(\tilde G) = D(|V| - 1) - \operatorname{def}$ from the $\operatorname{def} =
  \operatorname{corank}$ identity, \cref{thm:def-eq-corank}) and the split $D|V|
  = D(|V| - 1) + D$ collapses the bound to $\dim Z(G, q) \le D +
  \operatorname{def}$. The matching genericity-free lower bound $D +
  \operatorname{def} \le \dim Z(G, q)$ (\cref{lem:trivial-motions-rank-bound},
  available since every hinge is nondegenerate) pins the equality; deficiency
  nonnegativity rules out an over-count.
\end{proof}

The two rank attainments (eq.~(6.18) for $G_v^{ab}$, eq.~(6.22) for $G_v =
G_v^{ab} - ab$) combine into KT's eq.~(6.23). The $D - 1$ rows of $R(G_v^{ab},
q)$ at the edge $ab$ raise the rank of $R(G_v, q)$ by only $k' \le D - 2 < D -
1$ (the smaller graph $G_v$ already realizes all but $k'$ of the full rank), so
one of those $ab$-rows is a linear combination of the others together with the
$R(G_v, q)$ rows. The pure linear-algebra pigeonhole
(\cref{lem:case-III-claim-6-11-redundant-row}) applies at KT's two graphs
through the row-set identity
\[
  \operatorname{span}(R(G_v^{ab})\text{-rows}) =
  \operatorname{span}(R(G_v)\text{-rows}) + \operatorname{span}(ab\text{-block}):
\]
since $G_v = G_v^{ab} - ab$ and both frameworks share the seed $q$ and endpoint
selector $\mathit{ends}$ (hence the same supporting extensors), the only
rigidity rows of $R(G_v^{ab})$ not already in $R(G_v)$ are the $ab$-rows, which
span the $ab$-block. Combining the two rank attainments through it discharges
the corank gap and yields KT's redundant $ab$-row (\cref{lem:case-III-claim-6-11}).

\begin{lemma}[Redundant $ab$-row from a small corank over the $ab$-block; KT Claim 6.11, eq.\ (6.23)]
  \label{lem:case-III-claim-6-11-redundant-row}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.exists_redundant_panelRow_of_edge_of_finrank_lt}
  \leanok
  \uses{lem:case-III-seed-rank-bridge, lem:case-III-rank-attainment,
        lem:case-III-gap3-minimalKDof, lem:case-II-placement-new-rows}
  Let $F$ be a body-hinge framework with a nondegenerate hinge $e$ (distinct
  endpoints, $C(e) \ne 0$), and let $W$ be any subspace of the dual row space.
  The $e$-block --- the span of the panel rows at $e$ --- has dimension $D - 1$.
  If it raises the dimension of $W$ by strictly fewer than $D - 1$,
  \[
    \dim\bigl(W + \operatorname{span}(e\text{-block})\bigr) \;<\; \dim W + (D - 1),
  \]
  then there is a $(D-1)$-element linearly independent family $r$ spanning the
  $e$-block with one redundant member:
  \[
    r\,i_0 \in W + \operatorname{span}\bigl(r \restriction \{j \ne i_0\}\bigr).
  \]
\end{lemma}
\begin{proof}
  \leanok
  The $e$-block is $(D - 1)$-dimensional and is spanned by the $D - 1$
  independent panel rows a nondegenerate hinge contributes
  (\cref{lem:case-II-placement-new-rows}). The dimension hypothesis is exactly
  the abstract pigeonhole's input at this $(D-1)$-element family: a family
  raising $\dim W$ by fewer than its cardinality has a member lying in $W$ plus
  the span of the rest, and that member is the redundant $ab$-row.
\end{proof}

\begin{lemma}[KT Claim 6.11, eq.\ (6.23): the deleted $ab$-edge has a redundant row]
  \label{lem:case-III-claim-6-11}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.exists_redundant_panelRow_ab_of_finrank_eq}
  \leanok
  \uses{lem:case-III-claim-6-11-redundant-row, lem:case-III-seed-rank-bridge,
        lem:case-III-rank-attainment, lem:case-III-gap3-minimalKDof}
  Let $G_{ab} = G_v^{ab}$ be the split-off graph and $G_v = G_v^{ab} - ab$ its
  single-edge deletion, both realized as free-normal panel frameworks at the
  same inductively-fixed seed $q$ and endpoint selector $\mathit{ends}$, with
  $e_0 = ab$ the deleted nondegenerate hinge ($C(e_0) \ne 0$). Given the two
  rank attainments --- eq.~(6.18)
  \[
    \dim \operatorname{span}(R(G_{ab}, q)\text{-rows}) = D(m - 1)
  \]
  (the $0$-dof full rank, $m = |V(G_{ab})| = |V(G_v)|$) and eq.~(6.22)
  \[
    \dim \operatorname{span}(R(G_v, q)\text{-rows}) = D(m - 1) - k', \qquad k'
    \le D - 2
  \]
  --- one of the $D - 1$ rows of $R(G_{ab}, q)$ at the edge $ab$ is redundant
  modulo $\operatorname{span}(R(G_v, q)\text{-rows})$ and the other $ab$-rows:
  KT's eq.~(6.23). This redundant row is the $+1$ that the candidate-completion
  (\cref{sec:molecular-algebraic-induction-candidate-completion}) converts into
  the missing full-rank row.
\end{lemma}
\begin{proof}
  \leanok
  The two frameworks differ only by the edge $e_0 = ab$ (their links agree off
  $e_0$, and they share the same supporting extensors), so the row-set identity
  \[
    \operatorname{span}(R(G_{ab})\text{-rows}) =
    \operatorname{span}(R(G_v)\text{-rows}) + \operatorname{span}(ab\text{-block})
  \]
  holds. Writing $W = \operatorname{span}(R(G_v)\text{-rows})$, eq.~(6.18) reads
  $\dim\bigl(W + \operatorname{span}(ab\text{-block})\bigr) = D(m - 1)$ while
  eq.~(6.22) gives $\dim W = D(m - 1) - k'$; so the $ab$-block raises $\dim W$ by
  only $k' \le D - 2 < D - 1$. This is the corank gap that the redundant-row
  pigeonhole (\cref{lem:case-III-claim-6-11-redundant-row}) turns into the
  redundant $ab$-row.
\end{proof}

\subsubsection{The candidate-completion: converting the redundant row into the
missing \texorpdfstring{$+1$}{+1} (KT eqs.\ (6.24)--(6.29))}
\label{sec:molecular-algebraic-induction-candidate-completion}

Claim~6.11 (\cref{lem:case-III-claim-6-11}) gives a redundant $ab$-row of
$R(G_v^{ab}, q)$; the candidate-completion turns it into the full-rank $D$-th row
of the stratum-1 brick. KT's mechanism (eqs.~(6.24)--(6.29)) is a single fundamental
row operation \emph{followed by the eqs.~(6.14)--(6.15) column operation}: the redundant
row's linear combination, transported from $R(G_v^{ab}, q)$ up to $R(G, p_1)$ and then
column-reduced (add column $v$ to column $a$), produces a row $w$ of $R(G, p_1)$ whose
$V \setminus \{v\}$ part vanishes (eqs.~(6.27)--(6.28)). The pieces: the
eqs.~(6.24)--(6.25) decomposition of the redundant row
(\cref{lem:case-III-redundant-decomposition}), the per-edge seam identity
(\cref{lem:case-III-seam}), the pure-$v$-column characterization
(\cref{lem:case-III-vanish-off-column}), and the column operation $\Phi$
(\cref{lem:case-III-columnop}, the $\mathrm{col}_a \mathrel{+}= \mathrm{col}_v$
change-of-variables automorphism). The column op --- not the seam alone --- is what
makes the $V \setminus \{v\}$ part vanish: the transported row
$w = \mathrm{hingeRow}\ v\ a\ \rho_g$ is supported on \emph{both} columns $v$ and $a$,
and $w \circ \Phi$ is pure $v$-column.
The per-row eq.~(6.27) transport collapse (\cref{lem:case-III-transport-collapse}) closes the
core algebra: each transported $(vb)j$-row minus the inductive $(ab)j$-row is the pure $(va)$-hinge
row $\mathrm{hingeRow}\ v\ a\ (\mathrm{annihRow}\ C\ \cdot)$ (same supporting extensor by the shear
identity), and it packages the column-op composition certifying $w \circ \Phi$ pure $v$-column.
The combination-level threading
(\cref{lem:case-III-candidate-row-construction}): the redundant row's common $G_v$-row element
$w_{\mathrm{Gv}}$ is an $ab$-block element, so it is $\mathrm{hingeRow}\ a\ b\ \rho$ for some
$\rho \in r(p(e_0))$, its transported $(vb)i^*$-row $\mathrm{hingeRow}\ v\ b\ \rho$ is a genuine
rigidity row of $R(G, p_1)$, and the collapse yields the candidate row $\mathrm{hingeRow}\ v\ a\ \rho$.
The full $D(|V|-1)$-family assembly (\cref{lem:case-III-candidate-row}): the
operated row $w \circ \Phi$ joins the stratum-1 brick's $va$-block and old block in the pin-a-body
augment (\cref{lem:case-III-conditional-block}), and the operated independence transports back
through the automorphism $\Phi$ to the original family of size $((D-1)+1)+D(|V_v|-1) = D(|V|-1)$.
The one \emph{conditional} it passes through --- the eq.~(6.29) top-left $D \times D$
full rank for some line choice --- is discharged by
Claim~6.12 (\cref{lem:case-III-claim612},
\cref{sec:molecular-algebraic-induction-claim612}). The
$a$-column-block vanishing (\cref{lem:case-III-acolumn-zero}) is a Claim~6.12 fact (the $M_3$
case), not the candidate-row input.

\begin{lemma}[A rigidity row vanishing off $v$'s column is a pure $v$-column row; KT eq.\ (6.28)]
  \label{lem:case-III-vanish-off-column}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.dualMap_eq_comp_single_proj_of_vanish_off}
  \leanok
  Let $f$ be a functional on the screw assignments $\alpha \to \mathrm{ScrewSpace}\,k$
  that vanishes on every assignment supported off a fixed body $v$ (i.e.\
  $S v = 0 \Rightarrow f S = 0$). Then $f$ factors through body $v$'s screw column:
  $f = (f \circ \mathrm{single}_v) \circ \mathrm{proj}_v$. This is KT eq.~(6.28)'s
  formal content: the candidate-completion's row operation (eqs.~(6.24)--(6.27))
  reduces the $(vb)i^*$-row of $R(G, p_1)$ to a row whose $V \setminus \{v\}$ part is
  all zero, and such a row is determined by its $v$-column block --- written
  $S = \mathrm{single}_v(S v) + (S - \mathrm{single}_v(S v))$, the second summand is
  supported off $v$, so $f$ kills it.
\end{lemma}
\begin{proof}
  \leanok
  Split $S = \mathrm{single}_v(S v) + (S - \mathrm{single}_v(S v))$. The second
  summand reads $0$ at $v$, so $f$ kills it by hypothesis, leaving
  $f S = f(\mathrm{single}_v(S v)) = (f \circ \mathrm{single}_v)(S v)$, which is
  exactly $((f \circ \mathrm{single}_v) \circ \mathrm{proj}_v)\,S$.
\end{proof}

\begin{lemma}[The column operation $\Phi$: the candidate row is pure $v$-column in the operated frame; KT eqs.\ (6.14)--(6.16)]
  \label{lem:case-III-columnop}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.columnOp,
        CombinatorialRigidity.Molecular.BodyHingeFramework.hingeRow_comp_columnOp_apply,
        CombinatorialRigidity.Molecular.BodyHingeFramework.hingeRow_comp_columnOp_vanish_off}
  \leanok
  \uses{lem:case-III-vanish-off-column}
  KT's column operation $\mathrm{col}_a \mathrel{+}= \mathrm{col}_v$ (eqs.~(6.14)--(6.15)), the
  device that makes the transported candidate row vanish off $v$. Since the rigidity rows read
  relative screws $S_u - S_w$, the operation is modelled by its dual substitution on the screw
  assignment $\alpha \to \mathrm{ScrewSpace}\,k$:
  $\Phi(S) = \mathrm{update}\ S\ v\ (S_v + S_a)$, a linear automorphism for $v \neq a$ (inverse
  $\Phi^{-1}(S) = \mathrm{update}\ S\ v\ (S_v - S_a)$). It is rank-preserving (precomposition with a
  linear equivalence), so it certifies the candidate row's vanishing without changing any rank. The
  transported candidate row $w = \mathrm{hingeRow}\ v\ a\ \rho$ is supported on \emph{both} columns
  $v$ and $a$, with $w(S) = \rho(S_v - S_a)$; precomposing with $\Phi$ reads $\Phi(S)$ as $S_v + S_a$
  at $v$ and $S_a$ at $a$ (using $v \neq a$), so
  \[ w(\Phi(S)) = \rho\bigl((S_v + S_a) - S_a\bigr) = \rho(S_v), \]
  the $a$-column contribution cancels. In particular $w(\Phi(S)) = 0$ whenever $S_v = 0$, which is
  exactly the off-$v$ vanishing hypothesis \cref{lem:case-III-vanish-off-column} consumes: $w \circ
  \Phi$ is a pure $v$-column row. This --- KT's eqs.~(6.14)--(6.16) column reduction --- is the
  vanishing mechanism, \emph{not} the per-edge seam plus eq.~(6.43).
\end{lemma}
\begin{proof}
  \leanok
  $\Phi$ is a linear equivalence by direct verification of additivity, homogeneity, and the two
  inverse identities (each a case split on whether the coordinate is $v$; the off-$v$ coordinates are
  untouched, the $v$-coordinate cancels by $(S_v + S_a) - S_a = S_v$ and $(S_v - S_a) + S_a = S_v$,
  using $v \neq a$ so $\Phi(S)_a = S_a$). The evaluation $w(\Phi(S)) = \rho(S_v)$ is then immediate
  from $\Phi(S)_v = S_v + S_a$, $\Phi(S)_a = S_a$, and the cancellation; the $S_v = 0$ case
  substitutes $\rho(0) = 0$.
\end{proof}

\begin{lemma}[The seam identity: an edge's row depends only on the seed at its endpoints; KT eq.\ (6.26)]
  \label{lem:case-III-seam}
  \lean{CombinatorialRigidity.Molecular.PanelHingeFramework.ofNormals_panelRow_eq_of_ends_seed_eq}
  \leanok
  Two $\mathrm{ofNormals}$ realizations at seeds $q_0$, $q$ that \emph{agree at the
  two endpoints} of an edge $e$ --- $q_0(\mathrm{ends}\,e)_1 = q(\mathrm{ends}\,e)_1$
  and $q_0(\mathrm{ends}\,e)_2 = q(\mathrm{ends}\,e)_2$ as panel normals --- produce
  identical $e$-rows of their rigidity matrices: for every basis pair $(t_1, t_2)$,
  \[ R(G, q_0; e, t_1, t_2) = R(G_v^{ab}, q; e, t_1, t_2), \]
  regardless of the two underlying graphs. This is KT's seam identity
  $R(G, p_1; E \setminus \{vb\}, V \setminus \{v\}) = R(G_v^{ab}, q)$ in per-row form:
  a panel row $\mathrm{hingeRow}\,(\mathrm{ends}\,e)_1\,(\mathrm{ends}\,e)_2\,
  (\mathrm{annihRow}\,C(p(e))\,t_1\,t_2)$ reads only the endpoint selector, the index,
  and the supporting extensor $C(p(e)) = \mathrm{panelSupportExtensor}\,(\mathrm{normal}\,
  (\mathrm{ends}\,e)_1)\,(\mathrm{normal}\,(\mathrm{ends}\,e)_2)$, which depends on the
  seed only at the two endpoints (the underlying graph never enters a panel row).
\end{lemma}
\begin{proof}
  \leanok
  Both rows unfold to $\mathrm{hingeRow}\,(\mathrm{ends}\,e)_1\,(\mathrm{ends}\,e)_2\,
  (\mathrm{annihRow}\,C\,t_1\,t_2)$ with $C = \mathrm{panelSupportExtensor}$ of the two
  endpoint normals; the hypotheses $q_0 = q$ at $(\mathrm{ends}\,e)_1$, $(\mathrm{ends}
  \,e)_2$ make the two normals equal, hence the extensors and the rows.
\end{proof}

\begin{lemma}[The redundant $ab$-row as an explicit vanishing combination; KT eqs.\ (6.24)--(6.25)]
  \label{lem:case-III-redundant-decomposition}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.exists_redundant_panelRow_ab_decomposition}
  \leanok
  \uses{lem:case-III-claim-6-11}
  The functional-identity form of Claim~6.11 (\cref{lem:case-III-claim-6-11}, eq.~(6.23)) that the
  candidate-completion row operation consumes. Where \cref{lem:case-III-claim-6-11} delivers the
  \emph{membership} $r\,i^* \in \operatorname{span}(R(G_v, q)\text{-rows}) \sqcup
  \operatorname{span}(r \mathbin{''} \{j \neq i^*\})$ --- the $D - 1$ independent $ab$-rows $r$
  spanning the $ab$-block, one of them ($i^*$) redundant modulo the $G_v$-rows and the rest --- this
  lemma unwinds it (submodule $\sqcup$-membership) into KT's eq.~(6.24): the redundant $ab$-row
  $r\,i^*$ \emph{equals} a $G_v$-row element $w_{\mathrm{Gv}} \in \operatorname{span}(R(G_v,
  q)\text{-rows})$ plus a combination $w_{\mathrm{Other}} \in \operatorname{span}(r \mathbin{''} \{j
  \neq i^*\})$ of the other $ab$-rows. Rearranged this is the vanishing combination
  $r\,i^* - w_{\mathrm{Other}} - w_{\mathrm{Gv}} = 0$ of eq.~(6.24) with the $(ab)i^*$-coefficient
  $\lambda_{(ab)i^*} = 1$ (eq.~(6.25)): $r\,i^*$ carries coefficient $1$, $w_{\mathrm{Other}}$ the
  other $\lambda_{(ab)j}$ ($j \neq i^*$), $w_{\mathrm{Gv}}$ the $\lambda_{ej}$ over $E_v$.
\end{lemma}
\begin{proof}
  \leanok
  Immediate from \cref{lem:case-III-claim-6-11} by splitting the $\sqcup$-membership of the redundant
  row $r\,i^*$ into its $\operatorname{span}(R(G_v, q)\text{-rows})$ summand $w_{\mathrm{Gv}}$ and its
  $\operatorname{span}(r \mathbin{''} \{j \neq i^*\})$ summand $w_{\mathrm{Other}}$.
\end{proof}

\begin{lemma}[The $a$-column block of the eq.\ (6.24) vanishing combination is $0$; KT eq.\ (6.43)]
  \label{lem:case-III-acolumn-zero}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.exists_redundant_panelRow_ab_decomposition_acolumn_zero}
  \leanok
  \uses{lem:case-III-redundant-decomposition}
  The eqs.~(6.24)--(6.25) decomposition (\cref{lem:case-III-redundant-decomposition}) writes the
  redundant $ab$-row as $r\,i^* = w_{\mathrm{Gv}} + w_{\mathrm{Other}}$, i.e.\ the vanishing
  combination $g := w_{\mathrm{Gv}} + w_{\mathrm{Other}} - r\,i^* = 0$ --- KT's eq.~(6.24)
  $\sum_j \lambda_{(ab)j} R(G_v^{ab}, q; (ab)j) + \sum_{e \in E_v, j} \lambda_{ej} R(G_v^{ab}, q;
  e_j) = 0$ as a functional on the screw assignments $\alpha \to \mathrm{ScrewSpace}\,k$. KT
  eq.~(6.43) is its restriction to any single body $a$'s screw column: precomposing $g$ with the
  column injection $\mathrm{single}_a \colon \mathrm{ScrewSpace}\,k \to (\alpha \to
  \mathrm{ScrewSpace}\,k)$ (a screw on $a$, $0$ elsewhere) is again $0$,
  \[ g \circ \mathrm{single}_a = 0, \qquad\text{i.e.}\qquad
     \sum_{e \in E_v \cup \{ab\},\, j} \lambda_{ej}\, R(G_v^{ab}, q; e_j, a) = 0. \]
  This is the residual that the candidate-completion transport
  (\cref{lem:case-III-candidate-row}) must kill to certify that the transported row $w$'s
  $V \setminus \{v\}$ part vanishes: over $V \setminus \{v\}$ the $(vb)j$-row of $R(G, p_1)$ and the
  $(ab)j$-row of $R(G_v^{ab}, q)$ differ by exactly the $a$-column block $r_j(\cdot\,a)$, so $w$'s
  $V \setminus \{v\}$ part is the eq.~(6.24) sum ($= g(S) = 0$) minus that $a$-block.
\end{lemma}
\begin{proof}
  \leanok
  Since $r\,i^* = w_{\mathrm{Gv}} + w_{\mathrm{Other}}$
  (\cref{lem:case-III-redundant-decomposition}), the combination $g = w_{\mathrm{Gv}} +
  w_{\mathrm{Other}} - r\,i^*$ is the zero functional; precomposing it with $\mathrm{single}_a$ is
  therefore $0$ for every body $a$.
\end{proof}

\begin{lemma}[The conditional $D$-row block: the operated row lifts the $va$-block from $D-1$ to $D$; KT eq.\ (6.29)]
  \label{lem:case-III-conditional-block}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.linearIndependent_sum_pinned_block_augment}
  \leanok
  \uses{lem:case-II-placement-block-independent, lem:case-III-vanish-off-column,
        lem:case-III-columnop}
  KT eq.~(6.29)'s block-triangular $+1$, the assembly that takes the stratum-1 brick
  $D(|V|-1) - 1$ (\cref{lem:case-II-realization-placement}) to full $D(|V|-1)$ \emph{conditional} on
  the top-left $D \times D$ block being full rank. The new block is the $D - 1$ rows $r_n$ of body
  $v$'s $va$-hinge \emph{plus} the candidate-completion's operated extra row $w$ --- the pure
  $v$-column row $\mathrm{hingeRow}\,v\,a\,\rho_g \circ \Phi$ produced by the column op
  (\cref{lem:case-III-columnop}, \cref{lem:case-III-vanish-off-column}); the old block is the
  $D(|V_v|-1)$ rows $r_o$ of the split-off $G_v^{ab} \setminus (ab)i^*$, vanishing at
  $\mathrm{update}\ 0\ v$. The two blocks are jointly independent ---
  $\mathrm{Sum.elim}\,(\mathrm{Sum.elim}\,r_n\,(\lambda \_.\ w))\,r_o$ --- provided the
  \emph{augmented} pinned new block $\mathrm{Sum.elim}\,(r_n \circ \mathrm{single}_v)\,(w \circ
  \mathrm{single}_v)$ of $D$ functionals is independent on body $v$'s $D$-dimensional screw column
  (KT's eq.~(6.29) top-left $D \times D$ full rank: the $va$-block's $D-1$ pinned rows plus the
  operated $(vb)i^*$-row $w$'s $v$-column block). The $+1$ over the stratum-1 brick is the extra
  row; the count becomes $((D-1)+1) + D(|V_v|-1) = D(|V|-1)$.
\end{lemma}
\begin{proof}
  \leanok
  \uses{lem:case-II-placement-block-independent}
  $w$ is a new-block row pinned through $v$'s column (the pure-$v$-column property of the operated
  row, carried into the augmented-pinned-independence hypothesis), so it joins $r_n$ in the
  pin-a-body new block rather than needing the old-block vanishing. The statement is then exactly
  the pin-a-body column split (\cref{lem:case-II-placement-block-independent}) applied to the
  augmented new block indexed by $\iota_n \oplus \{\ast\}$: the augmented new block composed with
  $\mathrm{single}_v$ equals the supplied augmented pinned family (the $\mathrm{Sum.elim}$
  distributes over the per-index $\circ\,\mathrm{single}_v$), discharging $h_{\mathrm{newpin}}$.
\end{proof}

\begin{lemma}[The eq.\ (6.27) transport collapse: the $(vb)$-row minus the $(ab)$-row is a $(va)$-hinge row; KT eqs.\ (6.26)--(6.27)]
  \label{lem:case-III-transport-collapse}
  \lean{CombinatorialRigidity.Molecular.PanelHingeFramework.panelRow_vb_sub_panelRow_ab_eq_hingeRow_va,
        CombinatorialRigidity.Molecular.BodyHingeFramework.hingeRow_sub_hingeRow_eq,
        CombinatorialRigidity.Molecular.BodyHingeFramework.comp_columnOp_eq_comp_single_proj}
  \leanok
  \uses{lem:case-III-seam, lem:case-III-columnop, lem:case-III-vanish-off-column}
  The per-row form of KT eq.~(6.27)'s transport step, plus the column-op packaging that closes the
  eq.~(6.28) vanishing. At the degenerate eq.~(6.12) placement $p_1 = q_0$ ($v$'s normal at
  $n_a + t\,n_b$, $q_0 = q$ at the surviving endpoint $b$), the transported $(vb)j$-row of
  $R(G, p_1)$ and the inductive $(ab)j$-row of $R(G_v^{ab}, q)$ read the \emph{same} supporting
  extensor $C = \mathrm{panelSupportExtensor}\,n_a\,n_b$ --- the shear identity
  $\mathrm{panelSupportExtensor}\,(n_a + t\,n_b)\,n_b = \mathrm{panelSupportExtensor}\,n_a\,n_b$
  (KT eq.~(6.16)) --- so they are $\mathrm{hingeRow}\,v\,b\,(\mathrm{annihRow}\,C\,t_1\,t_2)$ and
  $\mathrm{hingeRow}\,a\,b\,(\mathrm{annihRow}\,C\,t_1\,t_2)$, and their difference is the pure
  $(va)$-hinge row
  \[ R(G, p_1; (vb)j) - R(G_v^{ab}, q; (ab)j) = \mathrm{hingeRow}\,v\,a\,(\mathrm{annihRow}\,C\,t_1\,t_2), \]
  by the hinge-difference collapse $\mathrm{hingeRow}\,v\,b\,\rho - \mathrm{hingeRow}\,a\,b\,\rho =
  \mathrm{hingeRow}\,v\,a\,\rho$ (since $(S_v - S_b) - (S_a - S_b) = S_v - S_a$). This is exactly the
  collapse KT eq.~(6.27) performs: transporting the redundant-$(ab)$-row combination as $(vb)$-rows
  and subtracting the inductive $(ab)$-combination (which the eq.~(6.24) decomposition makes vanish,
  $r\,i^* = w_{\mathrm{Gv}} + w_{\mathrm{Other}}$) collapses the transported row to
  $w = \mathrm{hingeRow}\,v\,a\,\rho_g$. The companion fact composes the column op $\Phi$
  (\cref{lem:case-III-columnop}) with the off-$v$ vanishing (\cref{lem:case-III-vanish-off-column}):
  the operated row $w \circ \Phi$ \emph{is} its own pure $v$-column factorization
  $(w \circ \Phi \circ \mathrm{single}_v) \circ \mathrm{proj}_v$, the eq.~(6.29) $(vb)i^*$-row.
\end{lemma}
\begin{proof}
  \leanok
  Both panel rows unfold to $\mathrm{hingeRow}\,\cdot\,b\,(\mathrm{annihRow}\,C\,t_1\,t_2)$ with the
  same $C$ (the shear identity at $q_0$ vs.\ the $(ab)$-extensor at $q$); the hinge-difference
  identity then gives the $(va)$-hinge row. The column-op packaging is the composition of
  \cref{lem:case-III-columnop} (the operated row vanishes off $v$) with
  \cref{lem:case-III-vanish-off-column} (a row vanishing off $v$ factors through $v$'s column).
\end{proof}

\begin{lemma}[The candidate row $w$ from the redundant row's common element; KT eqs.\ (6.24)--(6.28)]
  \label{lem:case-III-candidate-row-construction}
  \lean{CombinatorialRigidity.Molecular.PanelHingeFramework.exists_candidate_row_eq612}
  \leanok
  \uses{lem:case-III-redundant-decomposition, lem:case-III-transport-collapse}
  The combination-level threading that produces the candidate row $w$ from KT's redundant-$ab$-row.
  The input is the \emph{common} element $w_{\mathrm{Gv}}$ of the eqs.~(6.24)--(6.25) decomposition
  (\cref{lem:case-III-redundant-decomposition}): the $G_v$-row part of the redundant $ab$-row, which
  is itself an element of the $ab$-block $\operatorname{span}\{R(G_v^{ab}, q; (ab)\cdot)\}$
  (it equals $r\,i^* - w_{\mathrm{Other}}$, both terms in the block). By the per-edge block identity
  (\cref{lem:case-III-claim-6-11-redundant-row} infra) the $ab$-block is the
  $\mathrm{hingeRow}\,a\,b$-image of the $(D-1)$-dimensional hinge-row block
  $r(p(e_0)) = (\operatorname{span} C)^\perp$, so $w_{\mathrm{Gv}} = \mathrm{hingeRow}\,a\,b\,\rho$
  for some $\rho \in r(p(e_0))$. At the eq.~(6.12) seed $p_1 = q_0$ the shear identity
  $\mathrm{panelSupportExtensor}\,(n_a + t\,n_b)\,n_b = \mathrm{panelSupportExtensor}\,n_a\,n_b$
  (KT eq.~(6.16)) makes $\rho$ a hinge-row-block functional of $R(G, p_1)$'s $e_b = vb$-hinge too,
  so $\mathrm{hingeRow}\,v\,b\,\rho$ is the transported $(vb)i^*$-row, a genuine rigidity row of
  $R(G, p_1)$. Its eq.~(6.27) collapse against the inductive $(ab)$-part is the pure $(va)$-hinge
  candidate row $w$:
  \[ \mathrm{hingeRow}\,v\,b\,\rho \;-\; w_{\mathrm{Gv}} \;=\; \mathrm{hingeRow}\,v\,a\,\rho \;=\; w, \]
  by the hinge-difference collapse (\cref{lem:case-III-transport-collapse}).
\end{lemma}
\begin{proof}
  \leanok
  \uses{lem:case-III-transport-collapse}
  The $ab$-block element $w_{\mathrm{Gv}}$ is $(\mathrm{screwDiff}\,a\,b)^\vee\,\rho =
  \mathrm{hingeRow}\,a\,b\,\rho$ for some $\rho$ in the hinge-row block at $e_0$, by the block
  identity $\operatorname{span}\{R(G_v^{ab}, q; (ab)\cdot)\} = \mathrm{hingeRow}\,a\,b\,[r(p(e_0))]$.
  The same $\rho$ annihilates $C(p_1(e_b))$ because $C(p_1(e_b)) = C(q(e_0))$ by the shear identity,
  so $\mathrm{hingeRow}\,v\,b\,\rho$ is a rigidity row at the $vb$-link $e_b$. The collapse is the
  hinge-difference identity $(S_v - S_b) - (S_a - S_b) = S_v - S_a$.
\end{proof}

\begin{lemma}[The candidate-completion row operation: the missing $+1$ row $w$; KT eqs.\ (6.24)--(6.29)]
  \label{lem:case-III-candidate-row}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.linearIndependent_sum_augment_candidateRow}
  \leanok
  \uses{lem:case-III-claim-6-11, lem:case-III-redundant-decomposition,
        lem:case-III-candidate-row-construction,
        lem:case-III-vanish-off-column, lem:case-III-columnop, lem:case-III-seam,
        lem:case-III-transport-collapse,
        lem:case-III-conditional-block, lem:case-III-claim612,
        lem:case-II-realization-placement}
  Realize $G$ at the degenerate eq.~(6.12) placement $p_1$ with $p_1(vb) = q(ab)$
  (\cref{lem:case-II-realization-placement}). The redundant-$ab$-row of $R(G_v^{ab}, q)$,
  decomposed by \cref{lem:case-III-redundant-decomposition} (eqs.~(6.24)--(6.25)) into a $G_v$-row
  part and an other-$ab$-row part with $\lambda_{(ab)i^*} = 1$, is transported up to $R(G, p_1)$
  by the eq.~(6.26) row correspondence: the $(ab)j$-rows of $R(G_v^{ab}, q)$ map to the $(vb)j$-rows
  of $R(G, p_1)$ (KT eq.~(6.16)) and the $E_v$-rows correspond directly. After the KT eqs.~(6.14)--(6.15)
  \emph{column operation} (add column $v$ to column $a$), the resulting row
  $w = \sum_{1 \le j \le 5} \lambda_{(ab)j} R(G, p_1; (vb)j) + \sum_{e \in E_v, j}
  \lambda_{e_j} R(G, p_1; e_j)$ (eq.~(6.27)) has its $V \setminus \{v\}$ part all zero. By
  \cref{lem:case-III-vanish-off-column}, $w$ is then the pure $v$-column row
  $w = \bigl(\sum_j \lambda_{(ab)j}\, r_j(q(ab)),\ 0\bigr)$ of eq.~(6.28) (using $p_1(vb) = q(ab)$ and
  $R(G, p_1; E_v, v) = 0$). It is the new $(vb)i^*$-row that, joining the $va$-block, lifts $v$'s
  column contribution from $D - 1$ to $D$: the matrix $R(G, p_1)$ becomes block-triangular as in
  eq.~(6.29), with bottom-right $R(G_v^{ab} \setminus (ab)i^*, q)$ of rank $D(|V \setminus \{v\}| - 1)$.
\end{lemma}
\begin{proof}
  \leanok
  \uses{lem:case-III-candidate-row-construction, lem:case-III-columnop,
        lem:case-III-vanish-off-column, lem:case-III-transport-collapse,
        lem:case-III-conditional-block, lem:case-III-claim612,
        lem:case-II-realization-placement}
  The one conditional --- the eq.~(6.29) top-left $D \times D$ full rank for some
  line choice --- is discharged by Claim~6.12 (\cref{lem:case-III-claim612}). The
  combination-level threading:
  \cref{lem:case-III-candidate-row-construction} takes the common element $w_{\mathrm{Gv}}$ of the
  eqs.~(6.24)--(6.25) decomposition (\cref{lem:case-III-redundant-decomposition}) --- the $G_v$-row
  part of the redundant $ab$-row, which is itself an $ab$-block element --- writes it as
  $w_{\mathrm{Gv}} = \mathrm{hingeRow}\ a\ b\ \rho$ for some $\rho \in r(p(e_0))$, certifies that the
  transported $(vb)i^*$-row $\mathrm{hingeRow}\ v\ b\ \rho$ is a genuine rigidity row of $R(G, p_1)$
  (the shear identity makes $\rho$ a hinge-row-block functional of the $e_b = vb$-hinge too), and
  collapses it to the candidate row $w = \mathrm{hingeRow}\ v\ a\ \rho$
  (\cref{lem:case-III-transport-collapse}). The row $w$ is supported on \emph{both} columns $v$ and
  $a$, with $w(S) = \rho(S_v - S_a)$; the column operation $\Phi(S) = \mathrm{update}\ S\ v\ (S_v +
  S_a)$ of eqs.~(6.14)--(6.15) (\cref{lem:case-III-columnop}) makes $w \circ \Phi$ pure $v$-column
  (\cref{lem:case-III-vanish-off-column}). \emph{The assembly}: precomposing the whole
  family $\mathrm{rn}\sqcup\{w\}\sqcup\mathrm{ro}$ with the automorphism $\Phi$ (the dual equivalence
  $\Phi^\vee$, preserving independence) yields the operated family, in which $w \circ \Phi$ joins the
  $va$-block $\mathrm{rn}$ in the pin-a-body augment (\cref{lem:case-III-conditional-block}) --- the
  old block $\mathrm{ro}$ is unchanged at the pin assignment $\mathrm{update}\ 0\ v\ x$ (since $\Phi$
  fixes it, $v \ne a$), so it keeps the brick's $\mathrm{hold}$ vanishing, and the $va$-rows keep
  their $v$-column pins. The augment is independent provided the eq.~(6.29) top-left $D \times D$
  block --- the $D - 1$ pinned $va$-rows plus the operated $w$'s $v$-column --- is full rank
  (the one \emph{conditional}, discharged by Claim~6.12, \cref{lem:case-III-claim612}); then
  transporting the operated independence back through $\Phi^\vee$ gives the original family
  $\mathrm{rn}\sqcup\{w\}\sqcup\mathrm{ro}$ independent, of size
  $((D-1)+1) + D(|V_v|-1) = D(|V|-1)$ --- the full-rank assembly with bottom-right
  $R(G_v^{ab} \setminus (ab)i^*, q)$ of rank $D(|V_v|-1)$. The off-route eq.~(6.43)
  $a$-column vanishing (\cref{lem:case-III-acolumn-zero}) is a Claim~6.12 fact (the $M_3$ case), not
  used here.
\end{proof}

\subsubsection{The \texorpdfstring{$D$}{D}-candidate disjunction (KT eqs.\ (6.30)--(6.45))}
\label{sec:molecular-algebraic-induction-claim612}

The candidate-completion (\cref{sec:molecular-algebraic-induction-candidate-completion}) lifts a
\emph{single} degenerate placement $p_1$ to full rank $D(|V|-1)$ \emph{provided} its top-left
$D \times D$ block (eq.~(6.29)) is full rank for some line choice (Claim~6.12,
\cref{lem:case-III-claim612}). At $d = 2$ KT
show this holds for $p_1$ or its symmetric partner $p_2$; \emph{at $d = 3$ it may fail for both}, so
the conditional is genuinely an existential over a free choice of lines. KT introduce a third candidate $p_3$ ---
splitting off at the \emph{other} degree-2 vertex $a$ along $vc$, whose split-off graph $G_a^{vc}$ is
isomorphic to $G_v^{ab}$ (eqs.~(6.31)--(6.41)) --- and prove Claim~6.12: at least one of the three
top-left blocks $M_1, M_2, M_3$ (of eqs.~(6.29), (6.30), (6.41)) has full rank, for some choice of
lines $L \subset \Pi(a)$, $L' \subset \Pi(b)$, $L'' \subset \Pi(c)$ (writing $\Pi(u) := \Pi_{G_v^{ab},
q}(u)$). The argument is a clean contrapositive: if all three fail, a single nonzero $r \in \R^D$
$\bigl(r := \sum_j \lambda_{(ab)j}\, r_j(q(ab))\bigr)$ is orthogonal to every supporting extensor in
the union~(6.45), which Lemma~2.1 forces to span all of $\R^D$, so $r = 0$ --- a contradiction.

\begin{lemma}[The $D$ panel-support extensors of $k+2$ linearly-independent homogeneous vectors span $\R^D$; KT eq.\ (6.45) via Lemma 2.1]
  \label{lem:case-III-claim612-extensor-span}
  \lean{CombinatorialRigidity.Molecular.span_omitTwoExtensor_eq_top}
  \leanok
  \uses{lem:extensor-independence, def:rigidity-matrix}
  Let $\bar p_1, \dots, \bar p_{k+2} \in \R^{k+2}$ be linearly independent homogeneous vectors. The
  $\binom{k+2}{2} = D$ panel-support $k$-extensors $C(\bar p_i \bar p_j) = \bar p_i \vee \bar p_j$ ---
  one per pair, the codim-2 panel through each pair --- span $\bigwedge^k \R^{k+2} =
  \mathrm{ScrewSpace}\,k$:
  \[
    \operatorname{span}\{\, \bar p_i \vee \bar p_j \mid 1 \le i < j \le k+2 \,\} = \bigwedge^k \R^{k+2}
    \quad (\operatorname{finrank} D).
  \]
  At $d = 3$ this is the $k = 2$ case ($D = 6$, $\bigwedge^2 \R^4$), the form the Case-III $d = 3$
  argument consumes.
\end{lemma}
\begin{proof}
  \leanok
  By the affine-free Lemma~2.1 (\cref{lem:extensor-independence},
  \texttt{omitTwoExtensor\_linearIndependent\_of\_li} at $e = k$: any $k+2$ linearly-independent
  vectors in $\R^{k+2}$) the $D$ omit-two extensors $\bar p_i \vee \bar p_j$ are linearly independent
  in $\bigwedge^k \R^{k+2}$; a linearly independent family of $D = \operatorname{finrank} \bigwedge^k
  \R^{k+2}$ vectors is a basis, so it spans. (The count of strictly-increasing index pairs is
  $\binom{k+2}{2} = D$, \texttt{Fintype.card\_subtype\_fst\_lt\_snd}.)
\end{proof}

\begin{lemma}[A functional annihilating a spanning set is zero]
  \label{lem:case-III-claim612-orthseq-vanish}
  \lean{CombinatorialRigidity.Molecular.eq_zero_of_annihilates_span_top}
  \leanok
  \uses{lem:case-III-claim612-extensor-span}
  If a screw-space functional $r \in (\R^D)^* = \mathrm{Module.Dual}\,\R\,(\mathrm{ScrewSpace}\,k)$
  vanishes on every element of a set $S$ whose span is all of $\R^D$, then $r = 0$. (The
  non-degeneracy of the dual pairing on $\R^D$: a functional vanishing on a spanning set vanishes
  everywhere.)
\end{lemma}
\begin{proof}
  \leanok
  Two linear maps agreeing on a spanning set are equal (\texttt{LinearMap.ext\_on}); $r$ agrees with
  the zero functional on $S$, and $\operatorname{span} S = \R^D$, so $r = 0$.
\end{proof}

The contrapositive needs the union~(6.45) to span $\R^D$, and this in turn rests on
\emph{two} facts that the single old node bundled but that the capstone consumes separately:
(N3a) four \emph{affinely-independent} points $p_1, \dots, p_4$ realizing the
$\Pi(a)/\Pi(b)/\Pi(c)$ incidence pattern --- so that Lemma~2.1
(\cref{lem:case-III-claim612-extensor-span}) forces their six joins to span; and (N3b) the
\emph{point-join $\leftrightarrow$ panel-meet duality}, which turns the capstone's hypothesis
``$r \perp C(L)$ for every line $L$ in $\Pi(a) \cup \Pi(b) \cup \Pi(c)$'' into ``$r$
annihilates each spanning join $p_i \vee p_j$''. The first is an \emph{existence}
(one good seed realizes the projective construction) fact, dispatched by the same Zariski device
the Case-I/II realization sites use; the second --- the genuinely new content --- is
Grassmann--Cayley projective duality.

\begin{lemma}[The intersection of fewer than $d+1$ panels has a homogeneous representative; KT eq.\ (6.45), existence half]
  \label{lem:case-III-claim612-panel-point-exists}
  \lean{CombinatorialRigidity.Molecular.exists_ne_zero_dotProduct_eq_zero}
  \leanok
  \uses{def:panel-hinge-framework}
  Given $m < d + 1$ panel normals $n \colon \mathrm{Fin}\,m \to \R^{d+1}$, the homogeneous incidence
  system $\langle \bar p, n_i \rangle = 0$ ($1 \le i \le m$) has a \emph{nonzero} solution
  $\bar p \in \R^{d+1}$: the $m \times (d+1)$ coefficient matrix has strictly more columns than rows,
  so its null space is nontrivial. This is the existence half of KT's "for any $j$ of the
  hyperplanes their intersection forms a $(d - j)$-dimensional affine space" (p.~698, eq.~(6.67)):
  with $m = j$ chosen panels, the projective solution set is a genuine $(d + 1 - j)$-dimensional
  subspace, hence carries a homogeneous representative incident to every chosen panel ($p_1$ on all
  three at $j = 3$; $p_2, p_3, p_4$ on each pair at $j = 2$).
\end{lemma}
\begin{proof}
  \leanok
  The incidence system is the linear map $\R^{d+1} \to \R^m$, $\bar p \mapsto (\langle \bar p, n_i
  \rangle)_i$, whose source dimension $d + 1$ strictly exceeds its target dimension $m$. By
  rank--nullity its kernel has dimension $\ge (d + 1) - m > 0$, so the kernel is nontrivial and
  contains a nonzero vector --- a homogeneous point orthogonal to every $n_i$.
\end{proof}

\begin{lemma}[A nonzero determinant polynomial in the seed yields an affinely-independent realization; KT eq.\ (6.45), existence-route closure half]
  \label{lem:case-III-claim612-points-affineIndep-producer}
  \lean{CombinatorialRigidity.Molecular.exists_affineIndependent_of_det_polynomial_ne_zero}
  \leanok
  \uses{lem:affine-indep-iff}
  Suppose the four candidate points are built as functions $p \colon (\sigma \to \R) \to
  \mathrm{Fin}\,4 \to \R^3$ of a panel-coordinate seed $q \colon \sigma \to \R$, and their
  affine-independence determinant --- the homogenization determinant
  $\det(\overline{p(q)}) = \det(\mathrm{homogenize} \circ p(q))$ of \cref{lem:affine-indep-iff} ---
  is the evaluation at $q$ of a \emph{nonzero} multivariate polynomial $P \in \R[\sigma]$. Then there
  is a seed $q$ at which the four points $p(q)$ are affinely independent. This is the \emph{closure}
  half of \cref{lem:case-III-claim612-points-affineIndep}: over the infinite field $\R$ a nonzero
  polynomial has a non-root, so it isolates the existence-bearing fact ($P \ne 0$, equivalently
  ``one seed works'') from the surrounding linear algebra, exactly as the existence
  half \cref{lem:case-III-claim612-panel-point-exists} isolates the dimension count.
\end{lemma}
\begin{proof}
  \leanok
  Over the infinite field $\R$ a nonzero polynomial does not vanish identically (the device's
  foundational non-root brick $\mathrm{MvPolynomial.exists\_eval\_ne\_zero}$, the same one Case~I
  uses to pick a shared seed), so $P$ has a non-root $q$; there $\det(\overline{p(q)}) = P(q) \ne 0$,
  and the determinant characterization \cref{lem:affine-indep-iff} turns this into affine
  independence of $p(q)$.
\end{proof}

\begin{lemma}[The affine-independence determinant of polynomial-valued candidate points is a polynomial in the seed; KT eq.\ (6.45), determinant-polynomial bridge]
  \label{lem:case-III-claim612-detpoly}
  \lean{CombinatorialRigidity.Molecular.exists_detPolynomial_of_pointPolynomial}
  \leanok
  \uses{lem:affine-indep-iff}
  Suppose the four candidate points are built coordinate-by-coordinate as multivariate polynomials
  in the panel-coordinate seed: a family $pp \colon \mathrm{Fin}\,4 \to \R^3[\sigma]$ of polynomial
  entries, with the realized point $p(q)_i = (q \mapsto \mathrm{eval}_q(pp_i))$. Then their
  affine-independence determinant --- the homogenization determinant $\det(\mathrm{homogenize} \circ
  p(q))$ of \cref{lem:affine-indep-iff} --- is the evaluation at $q$ of a \emph{single} polynomial
  $P \in \R[\sigma]$: the determinant of the $(d+1) \times (d+1)$ polynomial matrix whose rows are
  the homogenized polynomial points $\mathrm{Fin.snoc}(pp_i, 1)$. This discharges the determinant
  hypothesis ($\det(\overline{p(q)}) = P(q)$) of the closure producer
  \cref{lem:case-III-claim612-points-affineIndep-producer}, isolating the
  \emph{polynomial-in-the-seed} structure of the determinant from the genuinely genericity-bearing
  fact that this polynomial is nonzero ($P \ne 0$).
\end{lemma}
\begin{proof}
  \leanok
  Take $P$ to be the determinant of the polynomial matrix with rows $\mathrm{Fin.snoc}(pp_i, 1)$.
  Evaluation $\mathrm{eval}_q$ is a ring homomorphism, so it commutes with the determinant
  ($\mathrm{RingHom.map\_det}$); and it commutes with $\mathrm{Fin.snoc}(\cdot, 1)$ entrywise (it
  fixes the constant final coordinate $1$). Hence $\mathrm{eval}_q(P)$ equals the determinant of the
  matrix whose rows are the homogenized evaluated points $\mathrm{homogenize}(p(q)_i)$, which is
  exactly $\det(\overline{p(q)})$.
\end{proof}

\begin{lemma}[The explicit good seed: four affinely-independent points realizing the $\Pi(a)/\Pi(b)/\Pi(c)$ incidence pattern; KT eq.\ (6.45), the $P \ne 0$ witness]
  \label{lem:case-III-claim612-points-affineIndep-witness}
  \lean{CombinatorialRigidity.Molecular.exists_affineIndependent_panel_incidence}
  \leanok
  \uses{lem:affine-indep-iff}
  There exist three panel normals $n \colon \mathrm{Fin}\,3 \to \R^4$ in nonparallel position
  (linearly independent) and four \emph{affinely-independent} points $p \colon \mathrm{Fin}\,4 \to
  \R^3$ realizing the triple-intersection incidence pattern of KT eq.~(6.45):
  $p_1 \in \Pi(a) \cap \Pi(b) \cap \Pi(c)$, $p_2 \in \Pi(a) \cap \Pi(b) \setminus \Pi(c)$,
  $p_3 \in \Pi(b) \cap \Pi(c) \setminus \Pi(a)$, $p_4 \in \Pi(c) \cap \Pi(a) \setminus \Pi(b)$, where
  panel incidence is $p_i \in \Pi(u) \Leftrightarrow \langle \bar{p_i}, n_u \rangle = 0$ on the
  homogenization $\bar{p_i}$.

  This is the existence-route residual ($P \ne 0$) of
  \cref{lem:case-III-claim612-points-affineIndep}: by the converse of the non-root brick
  $\mathrm{MvPolynomial.exists\_eval\_ne\_zero}$ together with the determinant-polynomial bridge
  \cref{lem:case-III-claim612-detpoly}, $P \ne 0$ for the cross-product construction is
  \emph{logically equivalent} to exhibiting \emph{one} seed at which the constructed points are
  affinely independent --- no algebraic independence of the seed is needed.
\end{lemma}
\begin{proof}
  \leanok
  Take the coordinate-aligned seed: panel normals $n_a = e_0$, $n_b = e_1$, $n_c = e_2$ (the first
  three standard covectors of $\R^4$, hence linearly independent --- the nonparallel hypothesis the
  framework supplies), and the standard affine $3$-simplex $p = (0, e_3, e_1, e_2)$ of $\R^3$ (the
  origin plus three axis points). The incidence pattern is immediate from the coordinates: each
  $\bar{p_i}$ vanishes against exactly the panel covectors whose coordinate it is zero at. The
  $4 \times 4$ homogenization determinant is $\pm 1 \ne 0$, so by \cref{lem:affine-indep-iff} the four
  points are affinely independent (an explicit $\mathrm{Matrix.det\_succ\_row\_zero}$ cofactor
  expansion on the numeric matrix).
\end{proof}

\begin{lemma}[Four affinely-independent points with the $\Pi(a)/\Pi(b)/\Pi(c)$ incidence pattern; KT eq.\ (6.45) point choice]
  \label{lem:case-III-claim612-points-affineIndep}
  \lean{CombinatorialRigidity.Molecular.exists_affineIndependent_panel_incidence}
  \leanok
  \uses{def:rigidity-matrix, def:panel-hinge-framework,
        lem:case-III-claim612-panel-point-exists, lem:case-III-claim612-points-affineIndep-producer,
        lem:case-III-claim612-detpoly, lem:case-III-claim612-points-affineIndep-witness}
  At $d = 3$ there is a nonparallel arrangement of three panels --- three normals
  $n \colon \mathrm{Fin}\,3 \to \R^4$ in nonparallel position (linearly independent), spanning
  panels $\Pi(a), \Pi(b), \Pi(c) \subset \R^3$ --- together with four \emph{affinely-independent}
  points $p \colon \mathrm{Fin}\,4 \to \R^3$ realizing the triple-intersection incidence pattern
  \[
    p_1 \in \Pi(a) \cap \Pi(b) \cap \Pi(c), \quad
    p_2 \in \Pi(a) \cap \Pi(b) \setminus \Pi(c), \quad
    p_3 \in \Pi(b) \cap \Pi(c) \setminus \Pi(a), \quad
    p_4 \in \Pi(c) \cap \Pi(a) \setminus \Pi(b),
  \]
  where membership is the panel-incidence relation $p_i \in \Pi(u) \Leftrightarrow
  \langle \bar{p_i}, n_u \rangle = 0$ on the homogenization $\bar{p_i}$. The four points
  span an affine $3$-simplex; this is the affine independence \cref{lem:case-III-claim612-extensor-span}
  consumes (its six $2$-extensors $\bar{p_i} \vee \bar{p_j}$ then span $\R^D$).

  The panel normals are existentially quantified because the Case~III realization is itself being
  \emph{constructed}: the candidate placements $p_1, p_2, p_3$
  (\cref{lem:case-II-realization-placement,lem:case-III-claim612-p2-placement,%
  lem:case-III-claim612-p3-placement}) realize $G_v^{ab}$ at a panel-coordinate seed $q$ that the
  capstone (\cref{lem:case-III-claim612}) is free to choose, so the points and the panels they are
  incident to come from one chosen realization --- not from a fixed external normal arrangement.
\end{lemma}
\begin{proof}
  \leanok
  This is an \emph{existence} (Zariski) statement, like the realization sites of Cases~I and~II: it
  does not require the inductive seed to be generic, only that \emph{one} good seed exist. By the
  existence-route equivalence --- the converse of the non-root brick
  $\mathrm{MvPolynomial.exists\_eval\_ne\_zero}$ (run forward by the closure half
  \cref{lem:case-III-claim612-points-affineIndep-producer}, green) together with the
  determinant-polynomial bridge \cref{lem:case-III-claim612-detpoly} (green: the homogenization
  determinant of the polynomial-valued cross-product points is the evaluation of a single polynomial
  $P$ in the seed) --- the nonvanishing $P \ne 0$ that the closure route would need is
  \emph{logically equivalent} to exhibiting \emph{one} seed at which the constructed points are
  affinely independent. So no separate cross-product family, and no algebraic independence of the
  seed, is required: the explicit witness \cref{lem:case-III-claim612-points-affineIndep-witness}
  (green) supplies that one seed outright. It takes the coordinate-aligned arrangement $n_a = e_0$,
  $n_b = e_1$, $n_c = e_2$ (linearly independent in $\R^4$, hence nonparallel) and the standard
  affine $3$-simplex $p = (0, e_3, e_1, e_2)$ of $\R^3$: the $4 \times 4$ homogenization determinant
  is $\pm 1 \ne 0$ (\cref{lem:affine-indep-iff}, via the
  $\mathrm{Matrix.det\_succ\_row\_zero}$ cofactor expansion), and the incidence pattern is read
  straight off the coordinates. The pure dimension count
  \cref{lem:case-III-claim612-panel-point-exists} (a system of $< d + 1$ homogeneous incidence
  equations always has a nonzero solution) is the projective-existence content the cross-product
  construction would invoke for an arbitrary normal arrangement; the witness route makes it
  unnecessary by fixing the arrangement.
\end{proof}

\begin{lemma}[The witness points, parameterized by the \emph{real} split-leg panel normals; the (R1) reconciliation]
  \label{lem:case-III-claim612-points-affineIndep-real-normals}
  \lean{CombinatorialRigidity.Molecular.exists_homogeneousIncidence_of_normals}
  \leanok
  \uses{lem:case-III-claim612-panel-point-exists, lem:affine-indep-iff}
  The witness-point arrangement of \cref{lem:case-III-claim612-points-affineIndep} fixes its three
  panel normals to the coordinate frame $e_0, e_1, e_2$. The $d = 3$ Case~III producer
  (\cref{lem:case-II-realization}), however, reads its three panel normals $n_a, n_b, n_c$ off the
  \emph{inductive} split-leg realization $q$ it is handed --- nonparallel (linearly independent),
  but otherwise arbitrary --- and the candidate placement's supporting extensors must be incident to
  \emph{those} real normals. So the producer needs the witness points
  \emph{parameterized by given nonparallel $n \colon \mathrm{Fin}\,3 \to \R^4$}, not by a fixed
  frame. Given any linearly independent $n$, there exist four homogeneous coordinate vectors
  $\bar p \colon \mathrm{Fin}\,4 \to \R^4$ that are linearly independent and realize the same
  triple-intersection incidence pattern relative to $n$: $\langle \bar p_1, n_u\rangle = 0$ for all
  $u$, and each $\bar p_{i+1}$ orthogonal to two of the $n_u$ and strictly off the third.

  This homogeneous-coordinate construction is the \emph{whole} of the reconciliation: no
  de-homogenization to affine $\R^3$ points is needed. The two consumers --- the spanning
  \cref{lem:case-III-claim612-extensor-span} and its contrapositive \cref{lem:case-III-claim612} ---
  bottom out on Lemma~2.1 (\cref{lem:extensor-independence}), whose conclusion that the six omit-two
  joins span $\bigwedge^2 \R^4$ holds for \emph{any} linearly independent family of four vectors in
  $\R^4$ (the homogenization is only used to derive that linear independence from affine
  independence); and the seed-from-line transfer
  \cref{lem:case-III-claim612-line-in-panel-union} consumes the join points as bare vectors with
  $\langle \cdot, n_u\rangle = 0$ incidence, not as homogenizations of finite points. So the
  consumers take a linearly independent $\bar p \colon \mathrm{Fin}\,4 \to \R^4$ directly, fed by
  this lemma; the at-infinity worry (an orthogonal vector with vanishing last coordinate) lives
  only in a de-homogenization map the producer never calls, and the reconciliation carries no
  genericity residual.
\end{lemma}
\begin{proof}
  \leanok
  The pairing map $x \mapsto (u \mapsto \langle x, n_u\rangle)$ is the $\mathrm{mulVecLin}$ of the
  row matrix whose rows are the $n_u$; its rows are linearly independent, so the matrix has full row
  rank $3$ and the map is \emph{surjective}. Hence for any prescribed pairing target $t \in \R^3$
  there is a preimage $x$ with $\langle x, n_u\rangle = t_u$. Take $\bar p_2, \bar p_3, \bar p_4$ as
  the preimages of $(0,0,1), (1,0,0), (0,1,0)$ (giving exactly the off-one-panel incidence), and
  $\bar p_1$ as the nonzero common-perp vector of all three normals
  (\cref{lem:case-III-claim612-panel-point-exists}, $3 < 4$). Linear independence of the four
  $\bar p_i$ is the triangular argument on the incidence matrix: pairing
  $\sum_i g_i \bar p_i = 0$ against $n_0$ kills all but $\bar p_3 \Rightarrow g_3 = 0$; against $n_1
  \Rightarrow g_4 = 0$; against $n_2 \Rightarrow g_2 = 0$; then $g_1 \bar p_1 = 0$ with
  $\bar p_1 \ne 0 \Rightarrow g_1 = 0$. No genericity of $n$ beyond linear independence is used.
\end{proof}

\begin{lemma}[Point-join $\leftrightarrow$ panel-meet duality: a line in $\Pi(u)$ is annihilated by $r$ once $r \perp C(L)$; KT eq.\ (6.45) via Grassmann--Cayley]
  \label{lem:case-III-claim612-line-in-panel-union}
  \lean{CombinatorialRigidity.Molecular.complementIso_smul_eq_extensor_join,
        CombinatorialRigidity.Molecular.extensor_join_eq_zero_of_complementIso_eq_zero}
  \leanok
  \uses{def:join, def:meet, def:meet-complement-iso, def:panel-support-extensor,
        lem:extensor-independence, lem:case-III-claim612-extensor-span,
        lem:complement-iso-toDual, lem:complement-iso-decomposable-wedge-perp,
        lem:complement-iso-line-one-dim}
  At $d = 3$ ($D = 6$, $k = 2$, $\mathrm{ScrewSpace}\,2 = \bigwedge^2 \R^4$), let $p_i, p_j$ be two
  of the points of \cref{lem:case-III-claim612-points-affineIndep} whose connecting line
  $L = p_i p_j$ lies in a panel $\Pi(u)$ ($u \in \{a, b, c\}$). Then the point-join
  $\bar{p_i} \vee \bar{p_j} \in \bigwedge^2 \R^4$ and the panel-meet supporting extensor
  $C(L) = \mathrm{panelSupportExtensor}\ n_u\ (\cdot) = \mathrm{complementIso}(n_u \wedge \cdot)
  \in \bigwedge^2 \R^4$ of the same line are \emph{scalar multiples} of one another (both are the
  Pl\"ucker coordinate vector of the line $L$, up to the projective scale). Consequently a screw
  functional $r \in \mathrm{Module.Dual}\ \R\ (\mathrm{ScrewSpace}\,2)$ that annihilates
  \emph{every} panel-meet extensor $C(L)$ of lines $L \subset \Pi(a) \cup \Pi(b) \cup \Pi(c)$
  also annihilates each spanning join: $r(C(L)) = 0 \Rightarrow r(\bar{p_i} \vee \bar{p_j}) = 0$.
\end{lemma}
\begin{proof}
  \leanok
  Both $\bar{p_i} \vee \bar{p_j}$ and $C(L)$ are $2$-extensors in $\bigwedge^2 \R^4$
  representing the same projective line $L$. The join $\bar{p_i} \vee \bar{p_j}$
  (\cref{def:join}) is the line's supporting extensor written from \emph{two points on it}; the
  meet $C(L)$ (\cref{def:meet}, the complement iso \cref{def:meet-complement-iso} of the
  grade-$2$ join of two panel normals, \cref{def:panel-support-extensor}) is the same line's
  supporting extensor written from \emph{two hyperplanes through it}, and the Grassmann--Cayley
  duality of \cref{sec:molecular-meet} makes these two presentations of a $1$-dimensional
  projective subspace agree up to a nonzero scalar.

  We realize that common line concretely as the $1$-dimensional space
  \[
    \Omega := \{\, Z \in \textstyle\bigwedge^2 \R^4 \;:\; \langle Z, \phi\rangle_b = 0
      \text{ for all } \phi \in \widetilde\Phi \,\},
    \qquad
    \widetilde\Phi := (n_u \wedge \R^4) + (n' \wedge \R^4),
  \]
  the $\langle\cdot,\cdot\rangle_b$-orthogonal complement of the shared-direction span, where
  $\langle\cdot,\cdot\rangle_b$ is the coordinate (perfect) pairing of $\bigwedge^2 \R^4$ induced by
  the standard basis. The span $\widetilde\Phi$ is $5$-dimensional --- $\dim(n_u \wedge \R^4) +
  \dim(n' \wedge \R^4) - \dim(n_u \wedge \R^4 \cap n' \wedge \R^4) = 3 + 3 - 1$ by
  inclusion--exclusion, the intersection being the line $\mathrm{span}\{n_u \wedge n'\}$
  (\cref{lem:complement-iso-line-one-dim} supplies the exterior-square dimensions; the decomposable
  intersection is the genuinely new count). As $\langle\cdot,\cdot\rangle_b$ is a perfect pairing on
  the $6$-dimensional $\bigwedge^2 \R^4$, $\dim \Omega = 6 - 5 = 1$.

  Both extensors lie in $\Omega$. The panel-meet $C(L) = \mathrm{complementIso}(n_u \wedge n')$ does
  by step (i) (\cref{lem:complement-iso-decomposable-wedge-perp}): through the dictionary entry
  \cref{lem:complement-iso-toDual}, a wedge with a shared factor has vanishing complement pairing, so
  $\mathrm{complementIso}(n_u \wedge n')$ is annihilated by every $2$-extensor sharing a vector with
  $n_u \wedge n'$ --- in particular by each $n_u \wedge v$ (sharing $n_u$ directly) and each
  $n' \wedge v$ (the appended family repeats $n'$), which together span $\widetilde\Phi$. The
  point-join $\bar{p_i} \vee \bar{p_j}$ does too: under $\langle\cdot,\cdot\rangle_b$ the pairing
  $\langle \bar{p_i} \vee \bar{p_j},\, n_u \wedge v\rangle_b$ is the Gram determinant of the dot
  products $\langle \bar{p}, n_u\rangle$, $\langle \bar{p}, v\rangle$, whose row of incidences
  $\langle \bar{p_i}, n_u\rangle = \langle \bar{p_j}, n_u\rangle = 0$ vanishes (likewise for $n'$).
  Since $C(L) \ne 0$ precisely because $L$ is a genuine line ($\{n_u, n'\}$ independent), the two
  members of the line $\Omega$ are scalar multiples: $\bar{p_i} \vee \bar{p_j} = \lambda\, C(L)$
  (\cref{lem:complement-iso-line-one-dim}). The annihilation transfer is then immediate:
  $r(C(L)) = 0$ gives $r(\bar{p_i} \vee \bar{p_j}) = \lambda\, r(C(L)) = 0$. This is the
  duality KT use implicitly when reading eq.~(6.45): the spanning family of
  \cref{lem:case-III-claim612-extensor-span} (point-joins) and the annihilated family of the
  contrapositive (panel-meets) are the \emph{same} extensors of the lines in the panel union.
\end{proof}

\begin{lemma}[Block full rank $\Leftrightarrow$ $\hat r \notin (\operatorname{span} C)^\perp$; KT (6.42) row-space criterion]
  \label{lem:case-III-claim612-block-iff-perp}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.linearIndependent_sumElim_candidateRow_iff,
        CombinatorialRigidity.Molecular.BodyHingeFramework.mem_hingeRowBlock_iff}
  \leanok
  \uses{def:rigidity-matrix, def:hinge-row-block, lem:case-II-placement-new-rows}
  Let the $D - 1$ rows of a candidate's hinge block span $(\operatorname{span} C)^\perp = r(p(e))$
  (the $D-1$-dimensional orthogonal complement of the line $\operatorname{span} C$ supporting the
  candidate's $(d-2)$-affine subspace $L$; \cref{lem:case-II-placement-new-rows}), and let
  $\hat r := \sum_j \lambda_{(ab)j}\, r_j(q(ab))$ be the candidate row. The augmented $D$-functional
  family ($D - 1$ block rows plus $\hat r$) is linearly independent iff
  \[
    \hat r \notin (\operatorname{span} C)^\perp \quad\Longleftrightarrow\quad \hat r(C) \ne 0,
  \]
  i.e.\ the candidate's top-left $D \times D$ block has full rank iff $\hat r$ is not orthogonal to
  the supporting extensor $C$.
\end{lemma}
\begin{proof}
  \leanok
  The $D - 1$ block rows already span the hyperplane $(\operatorname{span} C)^\perp$; appending
  $\hat r$ raises the dimension to $D$ iff $\hat r$ is fresh, i.e.\ $\hat r \notin
  (\operatorname{span} C)^\perp$ --- the row-space criterion for extending an independent family by
  one vector ($\mathrm{linearIndependent\_sumElim\_unit\_iff}$, else the family stays inside the
  $(D-1)$-dimensional span). Freshness off the hinge-row block is exactly $\hat r(C) \ne 0$ by the
  dual-annihilator $\perp\!\perp$ correspondence $\hat r \in (\operatorname{span} C)^\perp = r(p(e))
  \Leftrightarrow \hat r(C) = 0$ ($\mathrm{mem\_hingeRowBlock\_iff}$, the annihilator evaluated at
  the spanning $C$).
\end{proof}

\begin{lemma}[The common vector $r$ is nonzero; KT (6.42)]
  \label{lem:case-III-claim612-r-nonzero}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.candidateRow_ne_zero}
  \leanok
  \uses{lem:case-III-redundant-decomposition, lem:case-II-placement-new-rows}
  The common vector of the disjunction, $r := \sum_{1 \le j \le 5} \lambda_{(ab)j}\, r_j(q(ab))
  \in \R^D$, is nonzero. (The $r_j(q(ab))$ are the $D - 1$ linearly independent rows of the $ab$-hinge
  block, and the coefficient $\lambda_{(ab)i^*} = 1$ from the redundant-row decomposition of
  eq.~(6.25); a combination with a unit coefficient on a member of a linearly independent family is
  nonzero.)
\end{lemma}
\begin{proof}
  \leanok
  The $r_j(q(ab))$ ($1 \le j \le 5$, i.e.\ $D - 1$ of them) are linearly independent
  (\cref{lem:case-II-placement-new-rows}); by \cref{lem:case-III-redundant-decomposition} the
  coefficient $\lambda_{(ab)i^*} = 1$ (eq.~(6.25)). A linear combination of a linearly independent
  family with a nonzero coefficient is nonzero
  ($\mathrm{linearIndependent\_sum\_smul\_ne\_zero}$).
\end{proof}

\begin{lemma}[The $p_2$ placement: the symmetric $va \leftrightarrow vb$ candidate; KT eqs.\ (6.19)/(6.30)]
  \label{lem:case-III-claim612-p2-placement}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.linearIndependent_sum_p2_candidateRow}
  \leanok
  \uses{lem:case-II-realization-placement, lem:case-III-columnop,
        lem:case-III-claim612-block-iff-perp}
  The second candidate $p_2$: split off at $v$ along $vb$ (rather than $va$), the symmetric image of
  $p_1$ under $a \leftrightarrow b$. It reuses the eq.~(6.12) degenerate placement
  (\cref{lem:case-II-realization-placement}) with the roles of $a, b$ exchanged, and the column
  operation at $(v, b)$ (\cref{lem:case-III-columnop}) in place of $(v, a)$. By the same
  candidate-completion assembly (\cref{lem:case-III-candidate-row}, applied at the swapped roles),
  \emph{if} its top-left $D \times D$ block $M_2$ (eq.~(6.30)) is full rank --- equivalently
  $r \notin (\operatorname{span} C(L'))^\perp$, i.e.\ $r(C(L')) \ne 0$ for the line $L' \subset
  \Pi(b)$ (the row-space criterion \cref{lem:case-III-claim612-block-iff-perp}) --- then $R(G, p_2)$
  attains the full $D(|V|-1)$-size independent family. The same $\lambda_{(ab)j}$ and index $i^*$
  appear in $M_1$ and $M_2$ (they are determined by $(G_v^{ab}, q)$, independent of $p_1, p_2$).
\end{lemma}
\begin{proof}
  \leanok
  Verbatim symmetric to the $p_1$ construction under $a \leftrightarrow b$. The candidate-completion
  assembly (\cref{lem:case-III-candidate-row}) takes the splitting body $v$, its two endpoints, the
  candidate functional $\rho$, and the two row blocks abstractly, so applying it at the column op
  $\Phi = \mathrm{columnOp}\ (v \ne b)$ for the edge $vb$ in place of $va$ is the same lemma. Its one
  hypothesis --- the operated, $v$-pinned top-left block being full rank --- is supplied by the
  row-space criterion (\cref{lem:case-III-claim612-block-iff-perp}) at the $vb$-hinge once the
  operated, pinned candidate row $(\mathrm{hingeRow}\ v\ b\ \rho) \circ \Phi \circ \mathrm{single}_v$
  is identified with $\rho$ itself (the column op makes it pure $v$-column with value $\rho(S_v)$),
  which holds iff $\rho(C) \ne 0$. The redundant-row combination is unchanged: it lives in
  $R(G_v^{ab}, q)$, common to both candidates.
\end{proof}

\begin{lemma}[$\mathrm{IsLink}$ under the $\rho = (a\,v)$ relabel; KT eq.\ (6.31)]
  \label{lem:splitOff-isLink-relabel}
  \lean{Graph.splitOff_isLink_relabel}
  \leanok
  \uses{}
  Given chain data $G.\mathrm{IsLink}\,e_a\,v\,a$, $G.\mathrm{IsLink}\,e_b\,v\,b$,
  $G.\mathrm{IsLink}\,e_c\,a\,c$ with degree-2 closures at $v$ and $a$, and fresh
  edges $e_0, e_1 \notin E(G)$ with $e_1 \ne e_0$, the two split-offs are
  related by
  \[
    (G.\mathrm{splitOff}\,a\,v\,c\,e_1).\mathrm{IsLink}\,e\,x\,y \;\iff\;
    (G.\mathrm{splitOff}\,v\,a\,b\,e_0).\mathrm{IsLink}\,(\sigma e)\,(\rho x)\,(\rho y),
  \]
  where $\rho = \mathrm{Equiv.swap}\,a\,v$ and
  $\sigma = (\mathrm{Equiv.swap}\,e_b\,e_0) \cdot (\mathrm{Equiv.swap}\,e_1\,e_c)$.
  This is KT eq.~(6.31): the graph-level isomorphism $G_a^{vc} \cong G_v^{ab}$ sending
  $e_b \mapsto e_0$ (the surviving $v\text{-}b$ edge becomes the fresh $a\text{-}b$ edge)
  and $e_1 \mapsto e_c$ (the fresh $v\text{-}c$ edge becomes the surviving $a\text{-}c$ edge).
\end{lemma}
\begin{proof}
  \leanok
  Expand $\mathrm{splitOff\_isLink}$ on both sides and match cases using the degree-2
  closures. Surviving edges $e \notin \{e_b, e_1\}$ have both endpoints away from $v$
  and $a$ (by closure at each), so $\sigma$ and $\rho$ fix them. The $e_b$-case maps to
  the fresh $e_0$-edge (endpoints $v \mapsto a$, $b \mapsto b$); the $e_1$-case maps to
  the surviving $e_c$-edge (endpoints $v \mapsto a$, $c \mapsto c$).
\end{proof}

\begin{lemma}[Fixed-seed realization transport across the $\rho = (a\,v)$ relabel; KT eq.\ (6.31)]
  \label{lem:splitOff-ofNormals-relabel}
  \lean{CombinatorialRigidity.Molecular.PanelHingeFramework.ofNormals_relabel,
        CombinatorialRigidity.Molecular.PanelHingeFramework.hasGenericFullRankRealization_of_splitOff_relabel}
  \leanok
  \uses{lem:splitOff-isLink-relabel}
  Under the same chain data and degree-2 hypotheses as \cref{lem:splitOff-isLink-relabel}, the
  transport runs in the \emph{producer's direction} and is stated at the $\mathrm{ofNormals}$ level
  --- \emph{not} as an existential. The induction hypothesis realizes the $v$-split
  $G.\mathrm{splitOff}\,v\,a\,b\,e_0$ ($G_v^{ab}$) concretely as
  $\mathrm{ofNormals}\,(G.\mathrm{splitOff}\,v\,a\,b\,e_0)\,\mathrm{ends}_0\,q_0$ with the four
  generic-realization conjuncts (general position; rigidity on $V(G)\setminus\{v\}$;
  link-recording; algebraic independence over $\Q$ of the seed). The lemma produces the same four
  conjuncts for the $a$-split $G.\mathrm{splitOff}\,a\,v\,c\,e_1$ ($G_a^{vc}$) at the
  \emph{same} seed transported by $\rho = \mathrm{Equiv.swap}\,a\,v$: the named framework
  $\mathrm{ofNormals}\,(G.\mathrm{splitOff}\,a\,v\,c\,e_1)\,\mathrm{ends}_{\sigma\rho}\,q_\rho$ with
  seed $q_\rho(x, i) := q_0(\rho\,x, i)$, selector
  $\mathrm{ends}_{\sigma\rho}(e) := (\rho\,(\mathrm{ends}_0(\sigma e))_1,\,\rho\,(\mathrm{ends}_0(\sigma e))_2)$,
  and $\sigma = (\mathrm{Equiv.swap}\,e_b\,e_0) \cdot (\mathrm{Equiv.swap}\,e_1\,e_c)$.
  Stating it at the fixed seed (rather than transporting the bare existential
  $\mathrm{HasGenericFullRankRealization}$) preserves the seed identity that eq.~(6.44) requires:
  the $M_3$ arm of the producer tests the \emph{same} $\hat r$ against all three panels, which is
  only meaningful when $(G_a^{vc}, q_\rho)$ shares the $v$-split's seed via $\rho$ --- an
  independent realization would carry a different seed and collapse the trichotomy. The bare
  existential form ($G_v^{ab}$-realization $\Rightarrow$ $G_a^{vc}$-realization at $q_0 \circ \rho$)
  is recorded as a short corollary.
\end{lemma}
\begin{proof}
  \leanok
  \uses{lem:splitOff-isLink-relabel}
  Each of the four conjuncts is transported by the vertex-swap $\rho$ and edge-permutation $\sigma$
  (an involution, since its two transpositions have disjoint supports). General position: the
  $\rho$-reindexed seed $q_\rho$ is GP because $\rho$ is injective and $q_0$ is GP. Rigidity: any
  motion $S$ of the $a$-split framework pulls back to a motion $S \circ \rho$ of the $v$-split
  framework, using \cref{lem:splitOff-isLink-relabel} to move each $a$-split link to a $v$-split
  link and the support-extensor equality across the two $\mathrm{ofNormals}$ terms; the $v$-split's
  rigidity on $V(G)\setminus\{v\}$ then forces $S \circ \rho$ constant, hence $S$ constant on
  $V(G)\setminus\{a\}$ since $\rho$ maps $V(G)\setminus\{a\} \to V(G)\setminus\{v\}$ bijectively
  (using $\rho\rho = \mathrm{id}$). Link-recording: \cref{lem:splitOff-isLink-relabel} forward
  translates each $a$-split link to a $v$-split link whose endpoints $\mathrm{ends}_0$ records,
  transported back through $\rho$ to $\mathrm{ends}_{\sigma\rho}$. Algebraic independence: $q_\rho$
  is an injective $\rho$-reindex of $q_0$.
\end{proof}

\begin{lemma}[Row-space correspondence under the $\rho = (a\,v)$ relabel; KT eqs.\ (6.31)/(6.44)]
  \label{lem:splitOff-rigidityRows-relabel}
  \lean{CombinatorialRigidity.Molecular.PanelHingeFramework.rigidityRows_ofNormals_relabel}
  \leanok
  \uses{lem:splitOff-isLink-relabel, lem:splitOff-ofNormals-relabel}
  The deliverable the $M_3$ candidate-row step (\cref{lem:case-III-claim612-p3-placement}) consumes.
  The coordinate-relabel map $\mathrm{funLeft}\,\rho \colon (\alpha \to \mathrm{ScrewSpace}\,k) \to
  (\alpha \to \mathrm{ScrewSpace}\,k)$, $S \mapsto S \circ \rho$, has dual
  $(\mathrm{funLeft}\,\rho)^* \colon \phi \mapsto \phi \circ (\cdot \circ \rho)$. Under it the
  rigidity rows of the relabelled $a$-split framework
  $\mathrm{ofNormals}\,(G.\mathrm{splitOff}\,a\,v\,c\,e_1)\,\mathrm{ends}_{\sigma\rho}\,q_\rho$ are
  exactly the image of the $v$-split framework's rigidity rows: each $a$-split row
  $\mathrm{hingeRow}\,u\,w\,r$ is $(\mathrm{funLeft}\,\rho)^*$ of the $v$-split row
  $\mathrm{hingeRow}\,(\rho u)\,(\rho w)\,r$.
\end{lemma}
\begin{proof}
  \leanok
  \uses{lem:splitOff-isLink-relabel}
  Since $\rho\rho = \mathrm{id}$, a direct computation gives
  $(\mathrm{funLeft}\,\rho)^*(\mathrm{hingeRow}\,(\rho u)\,(\rho w)\,r) = \mathrm{hingeRow}\,u\,w\,r$.
  \Cref{lem:splitOff-isLink-relabel} puts the two graphs' links in $\rho$-correspondence, and the
  hinge-row blocks at corresponding edges coincide (the same supporting extensor, by the same $q_0$
  reindex as in \cref{lem:splitOff-ofNormals-relabel}, since $\sigma$ is an involution). Hence the
  two rigidity-row \emph{sets} correspond exactly under $(\mathrm{funLeft}\,\rho)^*$, by mutual
  inclusion.
\end{proof}

\begin{lemma}[The $p_3$ placement: the third candidate via $G_v^{ab} \cong G_a^{vc}$; KT eqs.\ (6.31)--(6.41)]
  \label{lem:case-III-claim612-p3-placement}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.linearIndependent_sum_p3_candidateRow}
  \leanok
  \uses{lem:case-II-realization-placement, lem:case-III-columnop,
        lem:case-III-claim612-block-iff-perp, lem:case-III-claim612-eq644}
  The third candidate $p_3$, available because $a$ too is a degree-2 vertex: split off at $a$ along
  $vc$. The resulting graph $G_a^{vc}$ is isomorphic to $G_v^{ab}$ by $\rho \colon V \setminus \{a\}
  \to V \setminus \{v\}$, $\rho(v) = a$, $\rho(u) = u$ otherwise; the induced placement
  $q_\rho$ satisfies $p_3(va) = q(ac)$, $p_3(vb) = q(ab)$, $p_3(ac) = L''$, $p_3(e) = q(e)$ otherwise
  (eq.~(6.33)). The column operation at $(a, c)$ ($\mathrm{col}_a \mathrel{+}= \mathrm{col}_c$,
  eq.~(6.35)) block-triangularizes $R(G, p_3)$ to eq.~(6.41), whose top-left $D \times D$ block is
  $M_3$ with rows $r(p_3(ac)) = r(L'')$ and $\sum_j \lambda_{(ac)j}\, r_j(q(ac))$, full rank iff
  $r \notin (\operatorname{span} C(L''))^\perp$ for $L'' \subset \Pi(c)$ --- using eq.~(6.44)
  (\cref{lem:case-III-claim612-eq644}) to route the $M_3$ row onto the same $r$.
\end{lemma}
\begin{proof}
  \leanok
  The isomorphism $\rho$ identifies $(G_a^{vc}, q_\rho)$ with $(G_v^{ab}, q)$ as panel-hinge
  frameworks (eq.~(6.31)), so $\Pi_{G_a^{vc}, q_\rho}(u) = \Pi(u)$ for $u \ne v, a$ and
  $\Pi_{G_a^{vc}, q_\rho}(v) = \Pi(a)$. Formally this is invisible: the candidate-completion assembly
  (\cref{lem:case-III-candidate-row}) takes the splitting body, its two endpoints, the candidate
  functional $\rho$, and the two row blocks abstractly, so applying it at the column op $\Phi =
  \mathrm{columnOp}\ (a \ne c)$ for the edge $ac$ --- the split body now $a$, its operated endpoint
  $c$ --- is the same lemma; no graph swap is needed, and the iso is handled functionally through
  the candidate row (verbatim symmetric to the $p_2$ construction under the role change
  $v, b \mapsto a, c$). Its one hypothesis --- the operated, $a$-pinned top-left block being full
  rank --- is supplied by the row-space criterion (\cref{lem:case-III-claim612-block-iff-perp}) at
  the $ac$-hinge once the operated, pinned candidate row
  $(\mathrm{hingeRow}\ a\ c\ \rho) \circ \Phi \circ \mathrm{single}_a$ is identified with $\rho$
  itself (the column op makes it pure $a$-column with value $\rho(S_a)$), which holds iff
  $\rho(C) \ne 0$. The column op at $(a,c)$ (eq.~(6.35)$\to$(6.36)) and the substitution
  $p_3(va) = q(ac)$, $p_3(vb) = q(ab)$ (eq.~(6.34)) reduce $R(G, p_3)$ to eq.~(6.41) by the same row
  operations as for $R(G, p_1)$ (the eq.~(6.40) row reduction). The eq.~(6.44) identity
  (\cref{lem:case-III-claim612-eq644}) gives $M_3$'s row as $-\sum_j \lambda_{(ac)j}\, r_j(q(ac)) =
  r$ (up to sign), so the capstone reads its full-rank criterion off the same common vector $r$.
\end{proof}

\begin{lemma}[Eq.\ (6.44): the $M_3$ row equals $-r$, routing $M_3$ onto the same $r$; KT eq.\ (6.44)]
  \label{lem:case-III-claim612-eq644}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.candidateRow_ac_eq_neg,
        CombinatorialRigidity.Molecular.BodyHingeFramework.hingeRow_comp_single_tail,
        CombinatorialRigidity.Molecular.BodyHingeFramework.hingeRow_comp_single_off,
        CombinatorialRigidity.Molecular.BodyHingeFramework.acolumn_mem_hingeRowBlock_of_span_rigidityRows,
        CombinatorialRigidity.Molecular.BodyHingeFramework.hingeRow_acolumn_mem_span_rigidityRows}
  \leanok
  \uses{lem:case-III-acolumn-zero, lem:case-III-redundant-decomposition}
  In $G_v^{ab}$ only $ab$ and $ac$ are incident to the body $a$, so the $a$-column block of the
  eq.~(6.24) vanishing combination (\cref{lem:case-III-acolumn-zero}, eq.~(6.43)) reads
  \[
    0 = \sum_{1 \le j \le 5} \lambda_{(ab)j}\, r_j(q(ab)) + \sum_{1 \le j \le 5} \lambda_{(ac)j}\,
        r_j(q(ac)),
  \]
  which, with $r = \sum_j \lambda_{(ab)j}\, r_j(q(ab))$, gives KT's eq.~(6.44):
  \[
    r = -\sum_{1 \le j \le 5} \lambda_{(ac)j}\, r_j(q(ac)).
  \]
  Hence the $M_3$ candidate row $\sum_j \lambda_{(ac)j}\, r_j(q(ac))$ equals $-r$, routing the
  third candidate's full-rank criterion onto the \emph{same} common vector $r$ as $M_1, M_2$.
\end{lemma}
\begin{proof}
  \leanok
  Eq.~(6.43) (\cref{lem:case-III-acolumn-zero}) is the $a$-column restriction of the eq.~(6.24)
  vanishing combination; since only the $ab$- and $ac$-edges are incident to $a$ in $G_v^{ab}$
  (every other $R(G_v^{ab}, q; e_j, a) = 0$ by the rigidity-matrix definition --- the
  column-restriction leaves $\mathrm{hingeRow\_comp\_single\_tail}$ for the $ab$/$ac$-rows and
  $\mathrm{hingeRow\_comp\_single\_off}$ for every other edge), the surviving terms are exactly
  the $ab$- and $ac$-sums. Rearranging $\sum_j \lambda_{(ab)j}\, r_j(q(ab)) + \sum_j
  \lambda_{(ac)j}\, r_j(q(ac)) = 0$ for $r$ is eq.~(6.44).
\end{proof}

\begin{lemma}[Claim 6.12: for some line choice the candidate block has full rank; KT Claim 6.12, eqs.\ (6.30)--(6.45)]
  \label{lem:case-III-claim612}
  \lean{CombinatorialRigidity.Molecular.BodyHingeFramework.case_III_claim612}
  \leanok
  \uses{lem:case-III-claim612-points-affineIndep-real-normals, lem:case-III-claim612-r-nonzero}
  At $d = 3$, let $r \ne 0$ be the common candidate row and $\bar p_1, \bar p_2, \bar p_3, \bar p_4
  \in \R^4$ four linearly-independent homogeneous vectors
  (\cref{lem:case-III-claim612-points-affineIndep-real-normals}). Then \emph{for some
  pair} $(i, j)$ the join $r(\bar p_i \vee \bar p_j) \ne 0$. Since a candidate's hinge support is the
  panel-meet of its (freely chosen) line and the panel-meet of a line through two of the four
  points is exactly their join $p_i \vee p_j$, this is KT's Claim~6.12: for \emph{some} choice of
  $(d-2)$-affine line, the corresponding candidate's top-left $D \times D$ block (one of $M_1, M_2,
  M_3$ of eqs.~(6.29), (6.30), (6.41)) has full rank, equivalently $R(G, p_i)$ attains the desired
  rank $D(|V| - 1)$ --- discharging the candidate-completion conditional the candidate-completion
  assembly (\cref{lem:case-III-candidate-row}) passes through.
\end{lemma}
\begin{proof}
  \leanok
  \uses{lem:case-III-claim612-extensor-span, lem:case-III-claim612-orthseq-vanish}
  Contrapositive of the existential. Suppose $r$ annihilated \emph{every} one of the six joins
  $\bar p_i \vee \bar p_j$. (This negated existential is precisely KT's eq.~(6.45) hypothesis ``for
  \emph{every} choice of lines $L \subset \Pi(a)$, $L' \subset \Pi(b)$, $L'' \subset \Pi(c)$'' once
  each join is identified with the panel-meet $C(L)$ of the line through that pair via the
  point-join $\leftrightarrow$ panel-meet duality (\cref{lem:case-III-claim612-line-in-panel-union},
  with eq.~(6.44) routing all three failed blocks onto the same $r$); the formal argument works
  directly with the joins.) But the six joins of four linearly-independent homogeneous vectors span
  $\R^D$ (\cref{lem:case-III-claim612-extensor-span}, Lemma~2.1). So $r$ would annihilate a spanning
  set
  and hence be $0$ (\cref{lem:case-III-claim612-orthseq-vanish}), contradicting $r \ne 0$
  (\cref{lem:case-III-claim612-r-nonzero}). Therefore $r(p_i \vee p_j) \ne 0$ for some pair.

  The claim is genuinely an existential over a \emph{free} choice of lines, not a disjunction over
  three fixed supports: three $2$-extensors $C_1, C_2, C_3$ span at most $3$ of the $6$ dimensions
  of $\bigwedge^2 \R^4$, so $r \perp C_1, C_2, C_3$ alone cannot force $r = 0$; only the full
  six-join span~(6.45) (the free line sweep) does. The realization producer consumes the
  existential by building its candidate placement so its hinge line \emph{is} the witness join's
  line $L = \bar p_i \bar p_j$.
\end{proof}

\medskip\noindent\emph{The triangle floor.} The Case-III recursion needs a floor
of its own, and finding it is a point where the formalization must do visible
work that the paper covers in a phrase. KT open Lemma~6.13 by Lemma~4.6: $G$ is
either a short cycle (handled by the Lemma~5.4 cycle family,
\cref{lem:cycle-realization}) or carries a length-$d$ chain. The formalized
reduction runs this dichotomy directly (\cref{lem:chain-data-extract}); at
$d = 3$ its cycle disjunct is mathematically vacuous (a cycle witness forces
$|V(G)| = cy.m \le d = 3$, contradicting the $|V(G)| \ge 4$ floor the chain arm
runs under), so in practice the split recursion always descends via the chain
disjunct until $|V(G)| = 3$. There it genuinely cannot recurse: splitting off a degree-2 vertex
of a triangle leaves the \emph{double edge} on two bodies --- a non-simple graph
--- so the conditioned induction hypothesis (``general position \emph{if
simple}'') supplies only the bare conjunct for it, and with it the candidate
placement loses its transversality input (the linear independence of the two
split-leg normals). A triangle is a minimal $0$-dof-graph, and the realization
must be produced \emph{directly}: pin the vertex set and find the third edge by
edge counting (\cref{lem:triangle-third-edge}), realize three mutually
transversal panels by a cyclically consistent standard-basis seed
(\cref{lem:triangle-normals}), conclude rigidity from the three-extensor
telescoping independence (\cref{lem:theorem-55-triangle}, the three-body sibling
of the two-body base case \cref{lem:theorem-55-base}), and upgrade the attained
rank to a general-position, algebraically-independent seed by the genericity
device. This is the one floor of the $d = 3$ recursion, in the spirit of KT's
Lemma~5.4 short-cycle family, whose general-$m$ node \cref{lem:cycle-realization}
the dichotomy above now consumes directly (vacuously at $d = 3$).

\begin{lemma}[$|V|=3$ triangle base: vertex and edge pin; KT \S6.4]
  \label{lem:triangle-third-edge}
  \lean{Graph.exists_isLink_of_isMinimalKDof_card_three}
  \leanok
  \uses{}
  Let $G$ be a simple minimal $0$-dof graph on exactly three vertices, with edges $e_a : v$--$a$
  and $e_b : v$--$b$ and $e_a \ne e_b$. Then $a \ne b$, $V(G) = \{v,a,b\}$, and there exists an
  edge $f : a$--$b$.
\end{lemma}
\begin{proof}
  \leanok
  (1) $a \ne b$: if $a = b$ then $e_a, e_b$ are parallel, contradicting simplicity.
  (2) Vertex pin: $\{v,a,b\} \subseteq V(G)$ from the edge hypotheses; $|\{v,a,b\}| = 3 = |V(G)|$
  forces equality.
  (3) Third edge: the rank formula $\operatorname{rank} M(\tilde G) = D(|V|-1) = 2D$ and the rank
  bound $\operatorname{rank} M(\tilde G) \le (D-1)|E(G)|$ force $|E(G)| \ge 3$, so a third edge
  $f \notin \{e_a, e_b\}$ exists. Its endpoints lie in $\{v,a,b\}$; simplicity and
  \texttt{unique\_edge} eliminate all cases except $f : a$--$b$.
\end{proof}

\begin{lemma}[$|V|=3$ triangle base: generic realization; KT \S6.4]
  \label{lem:triangle-realization}
  \lean{CombinatorialRigidity.Molecular.PanelHingeFramework.hasGenericFullRankRealization_of_triangle}
  \leanok
  \uses{lem:triangle-third-edge, lem:theorem-55-triangle, lem:triangle-normals}
  Let $G$ be a simple minimal $0$-dof graph on exactly three vertices, with two
  edges $e_a, e_b$ from a common vertex $v$. Then $G$ has a generic full-rank
  panel-hinge realization.
\end{lemma}
\begin{proof}
  \leanok
  \uses{lem:triangle-third-edge, lem:theorem-55-triangle, lem:triangle-normals}
  T1 (\cref{lem:triangle-third-edge}) gives $V(G)=\{v,a,b\}$ and a completing edge
  $f:a$--$b$. T3 (\cref{lem:triangle-normals}) produces normals $n_0,n_1,n_2$ with
  pairwise nonvanishing joins and a linearly independent cyclic extensor family
  $C_0=C(n_0,n_1)$, $C_1=C(n_1,n_2)$, $C_2=C(n_2,n_0)$. Assigning $v\mapsto n_0$,
  $a\mapsto n_1$, $b\mapsto n_2$ yields a seed $q_0$; each triangle edge's support
  extensor equals $\pm C_i$ (sign from the canonical orientation), so T2
  (\cref{lem:theorem-55-triangle}) gives rigidity on $\{v,a,b\}=V(G)$.
  Upgrade to generic motive by
  \texttt{hasGenericFullRankRealization\_of\_rigidOn\_ofNormals}.
\end{proof}

\medskip\noindent\emph{The general-$d$ chain dispatch (KT \S6.4.2,
Lemma~6.13).} KT generalize the $d = 3$ Lemma~6.10 to arbitrary dimension
with the remark that ``the proof strategy is exactly the same as $d = 3$''
(p.~692) --- a phrase that compresses the two genuinely new general-$d$
moves, which we spell out here in full. The setting: $G$ simple, minimal
$0$-dof, no proper rigid subgraph, $|V| \ge 4$; the dichotomy
(\cref{lem:chain-data-extract}) yields a length-$d$ chain
$v_0 v_1 \cdots v_d$ whose interior vertices all have degree $2$ (the
short-cycle disjunct is closed directly by \cref{lem:cycle-realization}).

\emph{(1) One base, $d$ candidate views (KT eqs.\ (6.46)--(6.56)).} All
$d$ candidate frameworks are re-views of a \emph{single} splitting-off:
split off once at $v_1$ along $(v_0, v_2)$ with a fresh short-circuit edge
$e_0$, giving the base graph $G_1 = G_{v_1}^{v_0 v_2}$, which the (all-$k$)
induction hypothesis realizes as a generic framework $(G_1, q_1)$. The
$i$-th candidate $(G, p_i)$ re-attaches $v_1$ with its new hinge drawn on a
line $L_i$ inside the panel of the $i$-th chain vertex; KT's isomorphisms
$\rho_i$ (eqs.\ (6.54)--(6.56), ``exactly the same framework'') identify
candidate $i$ with candidate $1$ by an index shift along the chain --- there
is no second split and no second application of the IH. (The formalization
absorbs the $\rho_i$ into a single uniform relabel arm, so the isomorphisms
are invisible in the Lean; they are, however, how the \emph{paper} transfers
what it knows at the base panel to an arbitrary chain panel, and this
account keeps them explicit.)

\emph{(2) The single redundancy $r$ and its $\pm$ carry (KT eqs.\
(6.57)--(6.66)).} Claim~6.11 (\cref{lem:case-III-claim-6-11}) applies once,
at the base: $R(G_1, q_1)$ has a redundant $(v_0 v_2)$-row --- a functional
$\rho_0 \ne 0$ supported in the $e_0$-block whose hinge row lies in the span
of the genuine $G - v_1$ rows and which annihilates the base panel-support
extensor $C(v_0 v_2)$. (Its input, the eq.-(6.22) nested-IH rank bound at
the $k'$-dof subgraph $G - v_1$, is derived from the all-$k$ IH,
\cref{lem:case-III-nested-rank-lower}.) KT then carry this \emph{one}
redundancy across all $d$ panels of the chain, picking up only a sign ---
the ``$\pm r$'' of eq.\ (6.66) --- and this carry is where the paper's
compression is deepest. The mechanism matters: it is \emph{not} a transport
of a fixed row functional across the $d$ candidate frameworks. KT's own
eq.\ (6.62) moves the \emph{member}: the redundant row of $R(G_1, q_1)$
corresponds to a \emph{different} row of $R(G, p_i)$ for each $i$, so no
fixed-member transport exists. The faithful shape --- and the formalized one
--- is whole-matrix bookkeeping on the one matrix $R(G, p_i)$:
unit-determinant column operations (eqs.\ (6.60)--(6.61)) expose
$R(G_1, q_1)$ as a submatrix, and the $\pm r$ read at each panel is the
degree-2 column-vanishing fact of eq.\ (6.44) iterated along the chain ---
at an interior chain vertex only the two chain edges are incident, so that
vertex's column block of the vanishing combination has exactly two surviving
groups, and reading it converts the redundancy at one panel into the
sign-flipped redundancy at the next. Two structural points the formalization
forced explicit: the panel block $r(p(e))$ read at each step depends only on
the seed $q_1$ (it is graph-independent), so the candidate's own re-attached
panel is no harder than any other chain panel; and each candidate's target
count ``$\operatorname{rank} M_i + \operatorname{rank}(\text{base minus the
redundant row})$'' (eqs.\ (6.64)--(6.65)) is certified inline as one jointly
independent row family --- the $D - 1$ panel rows at $L_i$, the single
$\pm r$-row, and the trimmed base block --- via a block-form rank
certificate whose $\pm r$-row is sourced by a row operation from a bottom
$G - v_1$ row.

\emph{(3) The discriminator (KT eq.\ (6.67)).} Candidate $i$ closes exactly
when its top-left $D \times D$ block $M_i$ has full rank, which the
row-space criterion converts to $\rho_0(C(L_i)) \ne 0$ for its (freely
chosen) hinge line $L_i$. As at $d = 3$, the free choice is what saves the
argument: the $d + 1 = k + 2$ chain vertices' panel normals at the
algebraically independent seed are linearly independent homogeneous
vectors, so Lemma~2.1 at general grade
(\cref{lem:case-III-claim612-extensor-span}: their $\binom{k+2}{2} = D$
joins span the screw space) forces $\rho_0 \ne 0$ to pair nonzero with
\emph{some} join (\cref{lem:case-III-claim612-orthseq-vanish}) --- a matched
candidate index and a witness direction for its line. (KT invoke the
algebraic independence of the coefficients here, via ``any $j$ of the
hyperplanes meet in a $(d-j)$-dimensional affine space'', p.~698; the
formalization works at the homogeneous-vector layer, where linear
independence of the $d+1$ normals suffices and no new
algebraic-independence obligation arises.) The dispatch routes on the
matched index $i$: at the two base-adjacent panels ($i \le 1$) the $d = 3$
template arms fire at the base split's own roles ($M_1$/$M_2$); at an
interior panel ($2 \le i$) the relabel arm runs the $\pm r$ carry of
step~(2). Either way the matched candidate attains rank $D(|V|-1)$ at a
degenerate placement, and the genericity device re-realizes that rank at a
general-position, algebraically independent seed.

\begin{lemma}[Case III: $k = 0$, no proper rigid subgraph; KT Lemmas 6.10/6.13]
  \label{lem:case-III}
  \lean{CombinatorialRigidity.Molecular.PanelHingeFramework.case_III_realization_all_k,
        CombinatorialRigidity.Molecular.PanelHingeFramework.case_III_realization}
  \leanok
  \uses{def:panel-hinge-framework, def:rank-hypothesis,
        lem:simple-minimal-noRigid, lem:reduction-step}
  Let $G$ be a simple minimal $0$-dof-graph on $|V| \ge 3$ bodies with no
  proper rigid subgraph, at general dimension ($1 \le k$,
  $6 \le D = \binom{n+1}{2} = \binom{k+2}{2}$), and suppose the all-$k$
  induction hypothesis at every strictly smaller minimal $k'$-dof-graph
  ($k' \ge 0$ arbitrary): the graph is nonempty, it carries a
  general-position, link-recording realization at an algebraically-independent
  seed when it is simple, and a bare full-rank realization
  (\cref{def:rank-hypothesis}) in all cases.
  Then $G$ carries a general-position full-rank realization at rank
  $D(|V|-1)$ --- the splitting-off premise ($\mathtt{hsplitGP}$) of the
  all-$k$ Theorem~5.5 spine. One hypothesis rides explicitly: a fresh edge
  name for $G$ itself (an unused label outside $E(G)$, needed for the
  splitting-off's short-circuit edge $e_0$). The eq.-(6.22) nested-IH rank
  bound is discharged internally from the all-$k$ IH
  (\cref{lem:case-III-nested-rank-lower}), not carried; the $d = 3$ form is
  the $k = 2$ specialization.
\end{lemma}
\begin{proof}
  \leanok
  \uses{lem:triangle-realization, lem:chain-data-extract,
        lem:adjacent-degree-two-pair, lem:cycle-realization,
        lem:simple-minimal-noRigid,
        lem:reduction-step, lem:case-II-realization-placement,
        lem:case-III-claim-6-11, lem:case-III-candidate-row,
        lem:case-III-claim612, lem:case-III-claim612-p2-placement,
        lem:case-III-claim612-p3-placement, lem:case-III-claim612-eq644,
        lem:case-III-claim612-extensor-span,
        lem:case-III-claim612-orthseq-vanish,
        lem:splitOff-ofNormals-relabel, lem:splitOff-rigidityRows-relabel,
        lem:case-III-rank-attainment, lem:case-III-nested-rank-lower,
        lem:genericity-device}
  Induction on $|V(G)|$ inside the conditioned motive. At $|V| = 3$ the
  recursion bottoms on the triangle floor: an adjacent degree-2 vertex
  (\cref{lem:adjacent-degree-two-pair}) supplies the vertex-and-edge data
  \cref{lem:triangle-realization} needs, realized directly (see the
  discussion preceding \cref{lem:triangle-third-edge}). At $|V| \ge 4$, the
  general-$d$ chain extractor (\cref{lem:chain-data-extract}) runs the KT
  Lemma-4.6/4.8 chain-vs-short-cycle dichotomy. On the short-cycle disjunct,
  the Lemma-5.4 cycle family (\cref{lem:cycle-realization}) closes $G$
  directly --- a genuine branch at $d \ge 4$, mathematically vacuous at
  $d = 3$ (a cycle witness forces $|V(G)| = m \le d = 3$, contradicting the
  $|V| \ge 4$ floor this arm runs under). On the chain disjunct it returns
  the length-$d$ chain $v_0 v_1 \cdots v_d$ together with the base
  $v_1$-split's minimality/simplicity/measure data; splitting off at $v_1$
  along $(v_0, v_2)$ gives a smaller simple minimal $0$-dof-graph
  (\cref{lem:reduction-step}; simplicity of the split-off holds at
  $|V| \ge 4$), so the induction hypothesis hands a general-position,
  link-recording realization $(G_1, q_1)$ at an algebraically-independent
  seed. The eq.-(6.12) placement
  (\cref{lem:case-II-realization-placement}) re-attaches $v_1$ one row short
  of full rank; Claim~6.11 (\cref{lem:case-III-claim-6-11}, its eq.-(6.22)
  input discharged from the all-$k$ IH via
  \cref{lem:case-III-nested-rank-lower}) produces the redundant base row
  $\rho_0$, and the candidate-completion
  (\cref{lem:case-III-candidate-row}) converts it into the missing $+1$
  provided the matched candidate's top-left block is full rank. The
  candidate selection is the chain dispatch of the narrative above: the
  discriminator (\cref{lem:case-III-claim612-extensor-span} $+$
  \cref{lem:case-III-claim612-orthseq-vanish}, KT eq.\ (6.67)) picks a
  matched chain index; the base-adjacent arms fire the $M_1$/$M_2$ template
  at the base split's roles, and an interior index runs the uniform relabel
  arm --- the $\pm r$ carry of eqs.\ (6.60)--(6.66), the general-$d$ face of
  the $d = 3$ $M_3$ mechanism (\cref{lem:splitOff-ofNormals-relabel},
  \cref{lem:splitOff-rigidityRows-relabel},
  \cref{lem:case-III-claim612-eq644}); at $d = 3$ the dispatch specializes
  to the three-candidate Claim-6.12 disjunction
  (\cref{lem:case-III-claim612}, with \cref{lem:case-III-claim612-p2-placement}
  and \cref{lem:case-III-claim612-p3-placement} the $M_2$/$M_3$ placements).
  Each arm certifies its KT-(6.29)/(6.64) row count at the $t = 0$
  hinge-level family and transfers the attained rank along the
  one-parameter shear (the KT Lemma-5.2 rank-transfer, packaged with the
  seed-rank attainment \cref{lem:case-III-rank-attainment}); the genericity
  device (\cref{lem:genericity-device}) then re-realizes the attained rank
  at a general-position, algebraically-independent, link-recording seed,
  which is the conditioned motive's conclusion.

  The $d = 3$-only extraction pair \cref{lem:adjacent-degree-two-pair}
  remains consumed on the triangle floor; the dichotomy itself is run by
  the general extractor at every grade, its cycle disjunct closed by
  \cref{lem:cycle-realization} (vacuously at $d = 3$), and the cycle family
  stays --- as it always was --- load-bearing for the $k > 0$ Case~II
  (\cref{lem:case-II}).
\end{proof}
