\section{The $(k, \ell)$-count matroid}
\label{sec:count-matroid}

The proof-of-Lov\'asz--Yemini machinery in
\cref{sec:rigidity-matroid} characterises planar row-independence by
$(2, 3)$-sparsity. Before we package the result as a
\texttt{Matroid} in \cref{sec:rigidity-matroid-instance}, we record
the abstract object the characterisation eventually maps into:
the \emph{$(k, \ell)$-count matroid} on the edge set of $K_{V}$,
defined whenever $\ell < 2k$ and finite $V$. The matroid has no
linear-algebra content; it is determined entirely by the
combinatorial $(k, \ell)$-sparsity predicate of
\cref{def:isSparse}. Its independent sets are exactly the
$(k, \ell)$-sparse subgraphs; the planar rigidity matroid is the
$(k, \ell) = (2, 3)$ specialisation, identified with the linear
matroid of the rigidity matrix via \cref{thm:edgeSet-rowIndependent-iff-isSparse}.

This is the standard $(k, \ell)$-count matroid construction
introduced by Whiteley~\cite{whiteley1996} and developed in the
``pebble game'' framework of Lee--Streinu~\cite{leeStreinu2008}; the
range $\ell < 2k$ is the maximal regime in which the
$(k, \ell)$-sparse subsets are exactly the independent sets of a
matroid on $E(K_{V})$. Outside this regime the bases need not
agree in cardinality (e.g.\ $(3, 6)$-tight graphs in $3$-dimensional
rigidity).

\subsection{Augmentation in the matroidal regime}
\label{sec:count-matroid-aug}

The combinatorial heart of the matroid construction. The proof
follows Lee--Streinu~\cite{leeStreinu2008} (also Whiteley's appendix
of~\cite{whiteley1996}), via the \emph{$I$-component} decomposition
of a sparse graph: the maximal $I$-tight subsets partition the
non-free $I$-edges, are pairwise edge-disjoint, and (in the upper
range $\ell \in [k, 2k)$) overlap in at most one vertex. Two
$I$-tight sets sharing $\ge \lceil \ell / k \rceil$ vertices have
$I$-tight intersection (specialisation of
\cref{thm:isTightOn-union-inter}), so the family of $I$-tight sets
containing any fixed pair $\{u, v\}$ has a unique minimum, the
\emph{minimum $I$-block} containing $\{u, v\}$.

\begin{lemma}[$(k, \ell)$-sparsity augmentation, matroidal regime]
  \label{lem:isSparse-aug}
  \uses{def:isSparse, def:isTightOn, thm:isTightOn-union-inter}
  Let $V$ be finite, $k \ge 1$ and $0 \le \ell < 2k$, and let
  $I, J \subseteq \mathrm{Sym}_{2}V$ be off-diagonal edge sets such
  that both $\mathrm{fromEdgeSet}\,I$ and $\mathrm{fromEdgeSet}\,J$
  are $(k, \ell)$-sparse and $|I| < |J|$. Then there exists
  $e \in J \setminus I$ such that
  $\mathrm{fromEdgeSet}\,(I \cup \{e\})$ is $(k, \ell)$-sparse.
\end{lemma}
\begin{proof}
  \uses{def:isSparse, def:isTightOn, thm:isTightOn-union-inter}
  Suppose for contradiction that for every $e = uv \in J \setminus I$
  the set $\mathrm{fromEdgeSet}\,(I \cup \{e\})$ is not
  $(k, \ell)$-sparse. Since
  $|\edgesIn[\mathrm{fromEdgeSet}\,(I\cup\{e\})]{S}| =
   |\edgesIn[\mathrm{fromEdgeSet}\,I]{S}| + [u, v \in S]$, the
  sparsity violation at some $S \ni u, v$ pins $S$ to an
  $I$-block (\cref{def:isTightOn}) with $\{u, v\} \subseteq S$ and
  $|S| \ge 2$.

  The family of $I$-blocks containing $\{u, v\}$ is closed under
  pairwise intersection: two $I$-blocks $S, T \supseteq \{u, v\}$
  satisfy $|S \cap T| \ge 2 \ge \lceil \ell / k \rceil$ for
  $\ell < 2k$, so \cref{thm:isTightOn-union-inter} applies and
  $S \cap T$ is itself $I$-tight. Iterating gives a unique minimum
  $I$-block $T_{e} \supseteq \{u, v\}$. Define the
  \emph{$I$-component of $e$}, $C_{e} \supseteq T_{e}$, as the
  maximal $I$-tight superset of $T_{e}$ obtained by iterating
  \cref{thm:isTightOn-union-inter}'s union conclusion against
  every $I$-tight set sharing $\ge \lceil \ell / k \rceil$
  vertices with the current iterate.

  Two $I$-components $C_{1}, C_{2}$ share at most
  $\lceil \ell / k \rceil - 1$ vertices (else union closure
  would merge them, contradicting maximality), so each
  off-diagonal edge of $K_{V}$ with both endpoints in
  $C_{1} \cap C_{2}$ would have both endpoints in a shared
  $<\lceil \ell / k \rceil$-vertex set, which contains no
  off-diagonal edges in matroidal regime. Hence $I$-components
  are \emph{edge-disjoint}: each non-free $I$- (and $J$-) edge
  lies in a unique $I$-component.

  Now compute. Let $\mathrm{Comps}$ be the family of $I$-components
  and let $I_{\mathrm{free}} \subseteq I$ (resp.\
  $J_{\mathrm{free}} \subseteq J$) be the $I$- (resp.\ $J$-) edges
  with at least one endpoint outside every $I$-component. By the
  contradiction hypothesis, every $e \in J \setminus I$ has both
  endpoints in $C_{e} \in \mathrm{Comps}$, so $J_{\mathrm{free}}
  \subseteq J \cap I \subseteq I$, hence
  $|J_{\mathrm{free}}| \le |I_{\mathrm{free}}|$. At each
  $C \in \mathrm{Comps}$, $J$-sparsity at $C$ gives
  $|\edgesIn[\mathrm{fromEdgeSet}\,J]{C}| \le k|C| - \ell =
  |\edgesIn[\mathrm{fromEdgeSet}\,I]{C}|$. Summing over the
  edge-disjoint $I$-component partition:
  \[
    |J| = \sum_{C \in \mathrm{Comps}}
    |\edgesIn[\mathrm{fromEdgeSet}\,J]{C}| + |J_{\mathrm{free}}|
    \le \sum_{C \in \mathrm{Comps}}
    |\edgesIn[\mathrm{fromEdgeSet}\,I]{C}| + |I_{\mathrm{free}}|
    = |I|,
  \]
  contradicting $|I| < |J|$.

  In the upper range $\ell \in [k, 2k)$ the $I_{\mathrm{free}}$,
  $J_{\mathrm{free}}$ accounting vanishes since every edge is
  in a component (\cite[Theorem 5(4)]{leeStreinu2008}); the
  argument is uniform.
\end{proof}

The component-based proof above closes the matroid axiom directly.
Lee--Streinu also give an algorithmic proof via the
\emph{$(k, \ell)$-pebble game}: a greedy procedure that maintains
$k$ pebbles per vertex and a directed orientation, accepting edges
when $\mathrm{peb}(u) + \mathrm{peb}(v) \ge \ell + 1$. Formalising
the pebble game would yield a decision algorithm and an alternative
augmentation proof (rejected edges $\leftrightarrow$ tight
components); this project does not pursue it.

\subsection{The matroid}
\label{sec:count-matroid-def}

Now we package the augmentation as a \texttt{Matroid} via mathlib's
\texttt{IndepMatroid.ofFinite}. The three trivial axioms
($\mathrm{indep\_empty}$, $\mathrm{indep\_subset}$,
$\mathrm{subset\_ground}$) are immediate from the sparsity
predicate; the $\mathrm{indep\_aug}$ axiom is \cref{lem:isSparse-aug}.

\begin{definition}[$(k, \ell)$-count matroid]
  \label{def:countMatroid}
  \uses{def:isSparse, lem:isSparse-aug}
  For finite $V$ and integers $k \ge 1$ with $\ell < 2k$, the
  \emph{$(k, \ell)$-count matroid on $V$} is the matroid on the
  ground set $E(K_{V})$ (the off-diagonal symmetric pairs over $V$)
  whose independent sets are exactly the
  $(k, \ell)$-sparse edge sets:
  \[
    \mathrm{Indep}(I) \;\Longleftrightarrow\; I \subseteq E(K_{V})
    \wedge (\mathrm{fromEdgeSet}\,I).\mathrm{IsSparse}\,k\,\ell.
  \]
  Existence as a matroid is by mathlib's
  \texttt{IndepMatroid.ofFinite} applied to
  \cref{lem:isSparse-aug}.
\end{definition}

\begin{theorem}[Independent sets of the count matroid]
  \label{thm:countMatroid-indep-iff}
  \uses{def:countMatroid, def:isSparse}
  For finite $V$, $k \ge 1$, $\ell < 2k$ and
  $I \subseteq \mathrm{Sym}_{2}V$:
  $I$ is independent in the $(k, \ell)$-count matroid on $V$ if and
  only if $I \subseteq E(K_{V})$ and $\mathrm{fromEdgeSet}\,I$ is
  $(k, \ell)$-sparse.
\end{theorem}
\begin{proof}
  \uses{def:countMatroid}
  Immediate from the definition (definitional in Lean via
  \texttt{IndepMatroid.ofFinite\_indep}).
\end{proof}
