\section{Verified DFS interlude}
\label{sec:dfs}

The basic pebble game of \cref{sec:pebble-game} repeatedly searches a
partial orientation for a vertex with a free pebble reachable from a
target endpoint, returning the witness path so it can be reversed.
The inner primitive is a depth-first search on a directed graph with
predicate-based early termination; we formalize it once here, as a
short interlude, and consume it as a black box in
\cref{def:tryReachPebble}.

Mathlib's \texttt{Combinatorics.SimpleGraph}
$\mathrm{Walk}$ / $\mathrm{Reachable}$ infrastructure is declarative
(``a walk exists'', ``a reverse exists''), not algorithmic. We model
termination on \texttt{Batteries.UnionFind}, whose
\texttt{termination\_by self.rankMax - self.rank x} is a
strictly-decreasing numeric measure on a finite data structure;
here the measure is specialised to a $\mathrm{Finset}$-cardinality
on unvisited vertices.

\begin{definition}[Directed walk]
  \label{def:directed-walk}
  Let $V$ be a finite type with decidable equality and $R : V \to V
  \to \mathrm{Prop}$ a relation. A \emph{directed $R$-walk} from
  $u$ to $w$ is a non-empty list $u = v_0, v_1, \ldots, v_m = w$ of
  vertices satisfying $R\,v_i\,v_{i+1}$ for $0 \le i < m$. The walk
  is \emph{simple} if all $v_i$ are distinct.
\end{definition}

\begin{definition}[Reachable-finding DFS]
  \label{def:reachable-finding}
  \uses{def:directed-walk}
  Let $V$ be finite with decidable equality and $R$ a relation on $V$
  (with $R$ decidable in each argument, e.g.\ presented as an
  out-neighbour function $V \to \mathrm{Finset}\,V$). Given a source
  vertex $v$ and a predicate $P : V \to \mathrm{Bool}$, the
  procedure $\mathrm{reachableFinding}\,R\,v\,P$ performs a
  depth-first search along $R$-arcs from $v$, terminating at one of:
  \begin{itemize}
    \item (success) the first reachable vertex $w$ with $P(w)$,
      returning $\mathrm{some}\,(p, w)$ where $p$ is a simple
      $R$-walk from $v$ to $w$; or
    \item (failure) when the $R$-reachability closure of $v$
      contains no vertex satisfying $P$, returning $\mathrm{none}$.
  \end{itemize}
  Termination measure: $(\mathrm{Finset.univ} \setminus
  \mathrm{visited}).\mathrm{card}$, where $\mathrm{visited}$ is the
  finite set of vertices the DFS has already explored. Each
  recursive call expands $\mathrm{visited}$ by exactly one vertex,
  so the measure strictly decreases. (Compare
  \texttt{Batteries.UnionFind}'s
  \texttt{termination\_by self.rankMax - self.rank x}: the shape is
  the same --- a strictly-decreasing finite-domain measure --- but
  specialised to a $\mathrm{Finset}$-cardinality on a state set.)
\end{definition}

\begin{theorem}[Correctness of \texorpdfstring{$\mathrm{reachableFinding}$}{reachableFinding}]
  \label{thm:reachable-finding-correct}
  \uses{def:reachable-finding, def:directed-walk}
  Let $R$, $v$, $P$ be as in \cref{def:reachable-finding}. Then:
  \begin{enumerate}
    \item (soundness) If $\mathrm{reachableFinding}\,R\,v\,P =
      \mathrm{some}\,(p, w)$, then $p$ is a simple $R$-walk from $v$
      to $w$ and $P(w)$ holds.
    \item (completeness) If some vertex $w$ in the $R$-reachability
      closure of $v$ satisfies $P(w)$, then
      $\mathrm{reachableFinding}\,R\,v\,P = \mathrm{some}\,(p, w')$
      for some such $w'$ (possibly $w' \ne w$, since the DFS returns
      the first match in its traversal order, not necessarily the
      one the existence proof points at).
  \end{enumerate}
\end{theorem}
\begin{proof}
  \uses{def:reachable-finding}
  Both parts are by well-founded induction on the termination
  measure of \cref{def:reachable-finding}. Soundness: at each
  recursive step the walk extends by one arc (the DFS advances
  along an $R$-arc); the predicate is checked at each newly visited
  vertex before recursion; simplicity follows from the visited-set
  invariant (the DFS does not recurse into already-visited
  vertices). Completeness: assume some reachable $w$ satisfies $P$
  and the DFS returns $\mathrm{none}$. Then the visited set at
  termination contains the entire $R$-reachability closure of $v$
  (the recursion only stops when the frontier is empty), so it
  contains $w$; but a $P$-satisfying vertex would have returned
  $\mathrm{some}$ on first visit, contradicting termination by
  $\mathrm{none}$.
\end{proof}

\emph{Reuse pattern.} The pebble game's \cref{def:tryReachPebble}
specialises this primitive: it instantiates $R$ to be the outgoing-arc
relation of the partial orientation, $P$ to be ``has a free pebble'',
and on success post-composes a path-reversal step that flips every
arc of the returned walk. The DFS itself is fully reusable across
this and any future application that needs reachability with a
witness path; the path-reversal is pebble-game-specific.
