\section{Executable pebble game}
\label{sec:executable}

Phase~9 (\cref{sec:pebble-game}) formalises the basic
$(k, \ell)$-pebble game and proves its certificate-form correctness
theorem \cref{thm:pebble-game-correct}: in the matroidal regime
$\ell < 2k$, the finite simple graph $G$ is $(k, \ell)$-sparse if
and only if $\mathrm{runPebbleGame}\,k\,\ell\,G$ returns
$\mathrm{some}\,D$. But the Phase~9 \texttt{runPebbleGame} is
\texttt{noncomputable}: its body invokes \texttt{Finset.toList} (to
project a partial orientation's out-neighbour finset to a list for
the verified DFS) and \texttt{Quot.out} (to project
$G.\mathrm{edgeFinset} : \mathrm{Finset}\,(\mathrm{Sym}_2 V)$ to an
ordered-pair enumeration), both of which lift through
mathlib's \texttt{Classical}-flavored
\texttt{Quotient.lift}. Neither \texttt{\#eval} nor
\texttt{native\_decide} fires on it.

This chapter closes the gap. It surfaces the existing computable
workhorses
(\texttt{tryReachPebbleWith} / \texttt{tryAddEdgeWith} /
\texttt{runPebbleGameWith}, all already computable in Phase~9
since the math-layer / exec-layer split was anticipated) through a
\emph{computable} wrapper \texttt{runPebbleGameExec} that plugs the
two noncomputable primitives with \texttt{Finset.sort}-based
analogues under $[\mathrm{LinearOrder}\,V]$. The certificate-form
correctness theorem lifts to \texttt{runPebbleGameExec} with no
substantive new proof; on top of it we register $\mathrm{Decidable}$
instances for the project predicates $\mathrm{IsSparse}\,k\,\ell$,
$\mathrm{IsTight}\,k\,\ell$, and $\mathrm{IsLaman}$. The instances
have polynomial-time reduction bodies; they decide the same
propositions a brute-force $\forall$-finset iteration would
decide, but pick a fundamentally faster reduction.

A small \texttt{lake exe pebble-game} CLI built on the same
\texttt{Decidable} instance reads an edge-list file and prints
$\mathrm{LAMAN}$ / $\mathrm{SPARSE}$ / $\mathrm{NOT\_SPARSE}$.

\subsection{Runtime and backends}
\label{sec:executable-backends}

A reader who is new to Lean's evaluation model may be surprised
that the same \texttt{def} can be exercised through three
syntactically-different invocations that share the same function
body but route through different runtimes. Phase~10 leans on this
explicitly, so we spell it out.

Three Lean-level invocations land on
\texttt{runPebbleGameExec}'s compiled body:

\begin{itemize}
  \item \texttt{\#eval (decide (G.IsSparse 2 3))} (top-level
    elaborator command). Reduces $\mathrm{decide}\,P$ through the
    registered $\mathrm{Decidable}$ instance, whose body is the
    pebble game; runs in the bytecode interpreter. Fast in practice
    (polynomial in $|V| + |E|$, no kernel small-step overhead).
  \item \texttt{lake exe pebble-game input.txt} (built CLI binary).
    Same compiled body, but executed by Lean's native-code backend
    rather than the bytecode interpreter. Strictly faster constant
    factor; identical big-$O$.
  \item \texttt{by decide} (tactic, inside a proof of the form
    \texttt{theorem foo : G.IsLaman := by decide}). Reduces the
    \emph{same} $\mathrm{Decidable}$ instance, but through the
    Lean kernel's small-step reducer, not the compiled backend.
    Polynomially many small-step reductions per pebble-game
    operation $\Rightarrow$ impractically slow on any
    realistic graph. The tactic
    \texttt{native\_decide} compiles and runs the same body
    natively, paying a trust-the-compiled-binary axiom (Lean's
    \texttt{nativeDecide}); its trust footprint is identical to
    the CLI's.
\end{itemize}

The point of separating the chapter into these three rows is that
Phase~10 commits to making \texttt{\#eval} and
\texttt{lake exe pebble-game} fast on real graphs (polynomial,
end-to-end); \texttt{by decide} stays formally correct but its
runtime is academic. Worked examples in Lean (\cref{sec:executable-examples})
use \texttt{\#eval} or \texttt{native\_decide} accordingly.

\emph{Practical consequence: zero \texttt{noncomputable} in the
instance body.} For \texttt{\#eval} / \texttt{native\_decide} to
fire, the registered $\mathrm{Decidable}$ instance must reduce
without invoking \texttt{Classical.choice}, \texttt{Quot.out},
\texttt{Finset.toList}, or any other \texttt{noncomputable}
primitive. This is the operational reason Phase~10 builds
\texttt{runPebbleGameExec} from scratch rather than reusing
Phase~9's \texttt{runPebbleGame}: the latter is the same algorithm
mathematically, but its body has unreachable \texttt{noncomputable}
sub-terms that break native compilation.

\subsection{Computable list views}
\label{sec:executable-list-views}

The two noncomputable primitives in Phase~9's
\texttt{runPebbleGame} are localised: an out-neighbour list of a
partial orientation $D$ at a vertex $v$ (the input to the verified
DFS \texttt{reachableFinding}), and an enumeration of the input
graph's edges as ordered pairs (the input to the
\texttt{runPebbleGameWith} fold). Both are repaired by
$\mathrm{Finset.sort}$ under a $\mathrm{LinearOrder}$ on $V$.

\begin{definition}[Sorted out-neighbour list]
  \label{def:outListSorted}
  \lean{CombinatorialRigidity.PebbleGame.PartialOrientation.outListSorted}
  \leanok
  \uses{def:partial-orientation, def:pebble-counts}
  Let $V$ be a finite type with $[\mathrm{LinearOrder}\,V]$ and
  $[\mathrm{DecidableEq}\,V]$, and let $D$ be a partial orientation
  on $V$ (\cref{def:partial-orientation}). For $v \in V$, define
  \[
    \mathrm{outListSorted}_D(v)
    \;:=\; (D.\mathrm{outNbhd}\,v).\mathrm{sort}\,(\cdot \le \cdot)
    \quad : \quad \mathrm{List}\,V.
  \]
  Computable: \texttt{Finset.sort} requires only
  $\mathrm{LinearOrder}$, not \texttt{Classical} machinery. Replaces
  Phase~9's noncomputable \texttt{D.outList}.
\end{definition}

\begin{lemma}[Membership in the sorted out-neighbour list]
  \label{lem:mem-outListSorted}
  \lean{CombinatorialRigidity.PebbleGame.PartialOrientation.mem_outListSorted}
  \leanok
  \uses{def:outListSorted}
  Under the hypotheses of \cref{def:outListSorted}, for all $v, u
  \in V$,
  \[
    u \in \mathrm{outListSorted}_D(v) \;\iff\; (v, u) \in D.\mathrm{arcs}.
  \]
\end{lemma}
\begin{proof}
  \leanok
  $\mathrm{Finset.sort}$ preserves membership; the result follows
  from $D.\mathrm{mem\_outNbhd}$.
\end{proof}

\begin{definition}[Sorted edge enumeration]
  \label{def:edgeListSorted}
  \lean{SimpleGraph.edgeListSorted}
  \leanok
  \uses{def:partial-orientation}
  Let $V$ have $[\mathrm{LinearOrder}\,V]$ and
  $[\mathrm{DecidableEq}\,V]$, and let $G$ be a finite simple graph
  on $V$. Define
  \[
    \mathrm{edgeListSorted}\,G
    \;:=\; \mathrm{ofLex}\,\circ\,
      \big((G.\mathrm{edgeFinset}.\mathrm{image}\,(e \mapsto
        \mathrm{toLex}\,(e.\inf, e.\sup))).\mathrm{sort}\,(\cdot \le \cdot)\big)
    \quad : \quad \mathrm{List}\,(V \times V),
  \]
  where each $e : \mathrm{Sym}_2 V$ is first imaged to its
  lex-smaller-first ordered pair $(e.\inf, e.\sup)$ viewed as
  $\mathrm{Lex}\,(V \times V)$, then sorted under
  $\mathrm{Prod.Lex}$'s linear order, and finally untagged via
  $\mathrm{ofLex}$ to recover $V \times V$. Replaces Phase~9's
  noncomputable \texttt{G.edgeFinset.toList.map Quot.out}.

  \emph{Layer~0 audit~\#1 (revised outcome).} Sorting
  $\mathrm{Finset}\,(\mathrm{Sym}_2 V)$ directly via
  $(\cdot \le \cdot)$ is structurally blocked: mathlib's
  \texttt{instance : PartialOrder (Sym2 \(\alpha\))} from
  \texttt{SetLike} already occupies the slot for an order on
  $\mathrm{Sym}_2 V$ itself with the non-total subset order, so a
  competing linear order on $\mathrm{Sym}_2 V$ cannot be registered.
  Sorting via $\mathrm{Lex}\,(V \times V)$ on the
  $\mathrm{Sym}_2.\mathrm{sortEquiv}$ projection $(e.\inf, e.\sup)$
  is the mathlib-idiomatic alternative and requires no new
  $\mathrm{LinearOrder}$ instance to be mirrored.
\end{definition}

\begin{lemma}[Membership in the sorted edge list]
  \label{lem:mem-edgeListSorted}
  \lean{SimpleGraph.mem_edgeListSorted}
  \leanok
  \uses{def:edgeListSorted}
  Under the hypotheses of \cref{def:edgeListSorted}, for all
  $u, v \in V$,
  \[
    (u, v) \in \mathrm{edgeListSorted}\,G \;\iff\;
      u \le v \;\wedge\; s(u, v) \in G.\mathrm{edgeFinset}.
  \]
  Loop-freeness ($u \ne v$ for any list entry, via
  $\mathrm{not\_isDiag\_of\_mem\_edgeSet}$) and the
  $\mathrm{Sym}_2$-image round-trip
  $\big((\mathrm{edgeListSorted}\,G).\mathrm{map}\,s(\cdot,\cdot)\big).\mathrm{toFinset}
  = G.\mathrm{edgeFinset}$ are packaged as separate blueprint nodes
  with \cref{thm:runPebbleGameWith-correct} (the three discharges
  the workhorse-level statement consumes).
\end{lemma}
\begin{proof}
  \leanok
  Forward: an entry $(u, v) \in \mathrm{edgeListSorted}\,G$ is
  $\mathrm{ofLex}(\mathrm{toLex}\,(e.\inf, e.\sup))$ for some
  $e \in G.\mathrm{edgeFinset}$, so $(u, v) = (e.\inf, e.\sup)$;
  by case-split on $\mathrm{le\_total}\,a\,b$ for $e = s(a, b)$, this
  identifies $s(u, v) = e$ and so $s(u, v) \in G.\mathrm{edgeFinset}$.
  Backward: given $u \le v$ and $s(u, v) \in G.\mathrm{edgeFinset}$,
  take $e := s(u, v)$ and use $\mathrm{inf\_of\_le\_left}$ /
  $\mathrm{sup\_of\_le\_right}$ to compute $e.\inf = u$,
  $e.\sup = v$.
\end{proof}

\subsection{Computable wrapper}
\label{sec:executable-wrapper}

Phase~9 already factored the algorithm into a computable workhorse
\texttt{runPebbleGameWith} parametrised over a caller-supplied
\texttt{toSucc} and edge enumeration; the math-layer wrapper
\texttt{runPebbleGame} plugs in $\mathrm{outList}$ and
\texttt{G.edgeFinset.toList.map Quot.out}. Phase~10 adds a sibling
wrapper plugging in \cref{def:outListSorted,def:edgeListSorted}
instead.

\begin{definition}[Computable pebble-game wrapper]
  \label{def:runPebbleGameExec}
  \lean{CombinatorialRigidity.PebbleGame.PartialOrientation.runPebbleGameExec}
  \leanok
  \uses{def:runPebbleGame, def:outListSorted, def:edgeListSorted}
  Let $V$ have $[\mathrm{LinearOrder}\,V]$ and
  $[\mathrm{DecidableEq}\,V]$, and let $G$ be a finite simple graph
  on $V$. Define
  \[
    \mathrm{runPebbleGameExec}\,G\,k\,\ell
    \;:=\; \mathrm{runPebbleGameWith}\,\mathrm{empty}\,k\,\ell\,
      (\lambda D'. \mathrm{outListSorted}_{D'}\,\cdot)\,
      (\mathrm{mem\_outListSorted})\,
      (\mathrm{edgeListSorted}\,G).
  \]
  Computable end-to-end; no \texttt{Classical} dependencies.
\end{definition}

The Phase~9 wrapper-level correctness theorem
\cref{thm:pebble-game-correct} is stated against the math-layer
$\mathrm{runPebbleGame}$. Phase~10 lifts it to a workhorse-level
statement first, then re-derives both wrappers as one-line
corollaries.

\begin{theorem}[Pebble-game correctness, workhorse form]
  \label{thm:runPebbleGameWith-correct}
  \lean{CombinatorialRigidity.PebbleGame.PartialOrientation.runPebbleGameWith_correct}
  \leanok
  \uses{def:runPebbleGame, thm:pebble-game-correct,
    lem:runPebbleGame-underline-eq, lem:pebble-game-failure-witness}
  Let $V$ be finite with $[\mathrm{DecidableEq}\,V]$, let
  $\mathrm{toSucc} : \mathrm{PartialOrientation}\,V \to V \to
  \mathrm{List}\,V$ agree with the orientation's out-arc relation
  uniformly, and let $\mathrm{edges} : \mathrm{List}\,(V \times V)$
  enumerate $G.\mathrm{edgeFinset}$ via $\mathrm{Sym}_2.\mathrm{mk}$
  (no loops, pairwise distinct, $\mathrm{Sym}_2$-image equals
  $G.\mathrm{edgeFinset}$). Then in the matroidal regime
  $\ell < 2k$, $G$ is $(k, \ell)$-sparse if and only if
  $\mathrm{runPebbleGameWith}\,\mathrm{empty}\,k\,\ell\,
  \mathrm{toSucc}\,h_{\mathrm{toSucc}}\,\mathrm{edges}$ returns
  $\mathrm{some}\,D'$ for some partial orientation $D'$.
\end{theorem}
\begin{proof}
  \leanok
  \uses{thm:pebble-game-correct}
  The Phase~9 correctness chain already factors through workhorse-level
  helpers, all parametrised over a caller-supplied
  $\mathrm{toSucc}$ / $h_{\mathrm{toSucc}}$ and edge list:
  $\mathrm{runPebbleGameWith\_reachable}$ (Invariant $(4)$ transports
  through the fold), $\mathrm{runPebbleGameWith\_underline\_subset}$
  / $\mathrm{runPebbleGameWith\_mem\_underline}$ (underline tracking
  on monotone and no-skip-fires halves), and the fold-level failure-witness
  extraction $\mathrm{runPebbleGameWith\_eq\_none\_imp\_exists\_witness}$.
  Layer~2 packages these into three workhorse-level statements:
  $\mathrm{runPebbleGameWith\_underline\_eq}$ (underline equation from
  empty, given the three discharges \emph{hloops} / \emph{hpairwise} /
  \emph{himg}), $\mathrm{runPebbleGameWith\_sound}$ (soundness from
  empty, same three discharges), and
  $\mathrm{runPebbleGameWith\_eq\_none\_imp\_exists\_witness\_empty}$
  (contrapositive from empty, requiring only \emph{himg}). The iff
  then assembles in two lines: ($\Leftarrow$) the soundness wrapper;
  ($\Rightarrow$) contrapositive via the failure-witness lemma plus
  $\mathrm{IsSparse}$ at the produced subset. The math-layer
  \cref{thm:pebble-game-correct} re-derives as a one-line corollary
  plugging the math-layer enumeration $\mathrm{outList}$ /
  $G.\mathrm{edgeFinset.toList.map}\,\mathrm{Quot.out}$ back in (the
  three discharges follow from $\mathrm{not\_isDiag\_of\_mem\_edgeSet}$,
  $\mathrm{Finset.nodup\_toList}$, and the $\mathrm{Quot.out\_eq}$
  round-trip).
\end{proof}

\begin{theorem}[Computable wrapper correctness]
  \label{thm:runPebbleGameExec-correct}
  \lean{CombinatorialRigidity.PebbleGame.PartialOrientation.runPebbleGameExec_correct}
  \leanok
  \uses{def:runPebbleGameExec, thm:runPebbleGameWith-correct,
    lem:mem-outListSorted, lem:mem-edgeListSorted}
  Let $V$ be finite with $[\mathrm{LinearOrder}\,V]$
  and $[\mathrm{DecidableEq}\,V]$, and let $G$ be a finite simple
  graph on $V$. In the matroidal regime $\ell < 2k$, $G$ is
  $(k, \ell)$-sparse if and only if
  $\mathrm{runPebbleGameExec}\,G\,k\,\ell$ returns
  $\mathrm{some}\,D'$ for some partial orientation $D'$.
\end{theorem}
\begin{proof}
  \leanok
  \uses{thm:runPebbleGameWith-correct, lem:mem-outListSorted,
    lem:mem-edgeListSorted}
  Apply \cref{thm:runPebbleGameWith-correct} with the Phase~10
  list views \cref{def:outListSorted,def:edgeListSorted}; the
  agreement and round-trip hypotheses are
  \cref{lem:mem-outListSorted,lem:mem-edgeListSorted}. The
  formalisation additionally requires three small edge-list
  discharges (no loops, pairwise Sym$_2$-distinct, Sym$_2$-image
  round-trip), all packaged as one-liners against
  \cref{lem:mem-edgeListSorted} in
  \texttt{PebbleGame/Exec.lean}'s \texttt{SimpleGraph} section.
\end{proof}

\subsection{Decidability of sparsity, tightness, and Laman}
\label{sec:executable-decidable}

\begin{definition}[\texttt{Decidable} instance for sparsity]
  \label{def:isSparse-decidable}
  \lean{SimpleGraph.instDecidableIsSparse}
  \leanok
  \uses{def:runPebbleGameExec, def:isSparse,
    thm:runPebbleGameExec-correct}
  Let $V$ have $[\mathrm{Fintype}\,V]$,
  $[\mathrm{LinearOrder}\,V]$, and $[\mathrm{DecidableEq}\,V]$, and
  let $G$ be a finite simple graph on $V$. In the matroidal regime
  $\ell < 2k$ — packaged in the formalisation as
  $[\mathrm{Fact}\,(\ell < 2 \cdot k)]$ for typeclass synthesis — the
  proposition $G.\mathrm{IsSparse}\,k\,\ell$ is decidable via the rule
  \[
    \mathrm{decide}\,(G.\mathrm{IsSparse}\,k\,\ell)
    \;=\;
    (\mathrm{runPebbleGameExec}\,G\,k\,\ell).\mathrm{isSome}.
  \]
  The instance is registered as the canonical $\mathrm{Decidable}$
  for $\mathrm{IsSparse}\,k\,\ell$ in
  \texttt{PebbleGame/Exec.lean}.
\end{definition}

\begin{remark}[Polynomial vs.\ exponential reduction]
  \label{rem:decidable-runtime}
  $\mathrm{IsSparse}\,k\,\ell$ is $\forall s : \mathrm{Finset}\,V,
  \ldots$, which would also admit a brute-force decision procedure
  iterating all $2^{|V|}$ subsets (under appropriate decidability
  of the body) via $\mathrm{Fintype.decidableForallFintype}$. Both
  procedures decide the \emph{same} proposition and produce a valid
  $\mathrm{Decidable}$ instance; they differ only in the body of
  the instance, which is what $\mathrm{\#eval}$ and $\mathrm{decide}$
  actually reduce. The pebble-game instance reduces in polynomial
  time; the brute-force instance reduces in exponential time. To
  avoid typeclass search picking the slow path, the project
  registers \emph{exactly one} $\mathrm{Decidable}$ instance per
  predicate (this one), and the source comment forbids introducing
  competitors.
\end{remark}

\begin{definition}[\texttt{Decidable} instance for tightness]
  \label{def:isTight-decidable}
  \lean{SimpleGraph.instDecidableIsTight}
  \leanok
  \uses{def:isSparse-decidable, def:isTight}
  Under the hypotheses of \cref{def:isSparse-decidable},
  $G.\mathrm{IsTight}\,k\,\ell$ is decidable: it is the conjunction
  of $G.\mathrm{IsSparse}\,k\,\ell$ (\cref{def:isSparse-decidable})
  with the edge-count equation $|E(G)| + \ell = k \cdot |V|$, the
  latter decidable as an equality on $\mathbb{N}$. The
  formalisation bridges the $\mathrm{Set.ncard}$ /
  $\mathrm{Nat.card}\,V$ form of $\mathrm{IsTight}$ to the
  $\mathrm{Finset.card}$ / $\mathrm{Fintype.card}\,V$ form via the
  project mirror $\mathrm{ncard\_edgeSet\_eq\_card\_edgeFinset}$ and
  the mathlib lemma $\mathrm{Nat.card\_eq\_fintype\_card}$ — both
  pure rewrites with no algorithmic content.
\end{definition}

\begin{definition}[\texttt{Decidable} instance for Laman]
  \label{def:isLaman-decidable}
  \lean{SimpleGraph.instDecidableIsLaman}
  \leanok
  \uses{def:isTight-decidable, def:isLaman}
  Under the hypotheses of \cref{def:isSparse-decidable} at
  $(k, \ell) = (2, 3)$, $G.\mathrm{IsLaman}$ is decidable: it is
  $G.\mathrm{IsTight}\,2\,3$ (\cref{def:isTight-decidable})
  unfolded. The formalisation registers the matroidal-regime witness
  $[\mathrm{Fact}\,(3 < 2 \cdot 2)]$ as a top-level instance, so the
  Laman case is zero-hypothesis at the call site —
  $\mathrm{\#eval}\,(\mathrm{decide}\,G.\mathrm{IsLaman})$ fires
  without any caller-supplied $[\mathrm{Fact}\,\ldots]$.
\end{definition}

\subsection{Worked examples}
\label{sec:executable-examples}

The end-to-end-executability target is exercised by a small set of
$\mathrm{Fin}\,n$ examples in
\texttt{CombinatorialRigidity/PebbleGame/Examples.lean}. Each is a
$\mathrm{\#eval}$ of $\mathrm{decide}\,G.\mathrm{IsLaman}$ on a
concrete graph; the output should match the well-known
classification.

\begin{enumerate}
  \item \textbf{$K_4 \setminus e$} on $\mathrm{Fin}\,4$. Laman: 4
    vertices, 5 edges, $2 \cdot 4 - 3 = 5$. The Phase~3 worked
    example (\cref{thm:top-fin-four-minus-edge-isLaman}); verifies
    the minimal non-trivial accept path.
  \item \textbf{Moser spindle.} A 7-vertex 11-edge Laman graph
    ($2 \cdot 7 - 3 = 11$). Verifies the accept path on a graph
    large enough for the DFS to fire non-trivially.
  \item \textbf{$K_4$ on $\mathrm{Fin}\,4$.} Not Laman, globally
    over-constrained: 4 vertices, 6 edges, $6 > 2 \cdot 4 - 3 = 5$.
    The pebble game rejects on the full-vertex subset at the
    sparsity step; the blocking witness $V' = V$ violates
    Invariant~(4).
  \item \textbf{A sparse-but-not-tight example.} A path on 5
    vertices: $(2, 3)$-sparse but $|E| = 4 < 2 \cdot 5 - 3 = 7$,
    so not tight, hence not Laman. The pebble game accepts; the
    \cref{def:isTight-decidable} edge-count check rejects.
\end{enumerate}

These examples additionally demonstrate the
$\mathrm{\#eval}$ / $\mathrm{native\_decide}$ runtime claim of
\cref{sec:executable-backends}: the reduction completes in
sub-second wall time on a developer laptop.

\subsection{CLI binary}
\label{sec:executable-cli}

A \texttt{lake exe pebble-game} entry point reads a simple
edge-list file and prints the classification. Input format: a
leading line \texttt{n m} giving vertex count and edge count,
followed by $m$ lines of two whitespace-separated integers
$0 \le u, v < n$. Output: one of $\mathrm{LAMAN}$ /
$\mathrm{SPARSE\_NOT\_TIGHT}$ / $\mathrm{NOT\_SPARSE}$.

The binary is a thin \texttt{IO} wrapper:
\begin{enumerate}
  \item Parse the input file; reject malformed input with a
    diagnostic.
  \item Construct $G : \mathrm{SimpleGraph}\,(\mathrm{Fin}\,n)$
    from the parsed edge list via $\mathrm{fromEdgeSet}$.
  \item Invoke
    $\mathrm{runPebbleGameExec}\,G\,2\,3$ via the
    \cref{def:isLaman-decidable} instance.
  \item Print the classification.
\end{enumerate}

No blueprint nodes for the binary itself — it is not Lean-formal
content but the natural delivery vehicle for the runtime claim of
\cref{sec:executable-backends}. \emph{Optional refinement:} extract
a blocking-witness subset on the $\mathrm{NOT\_SPARSE}$ branch via
the wrapper-level failure-witness
\cref{lem:pebble-game-failure-witness}; whether this can be done
without falling back to \texttt{Classical} machinery (the
$\mathrm{Reach}$-closure uses \texttt{Classical.decPred} in
Phase~9) is part of the Phase~10 Layer~5 work.

\emph{Special case: planar rigidity (continued).} The CLI binary
at the default parameters $(k, \ell) = (2, 3)$ is, by
\cref{thm:pebble-game-correct} composed with
\cref{thm:rigidityMatroid-indep-iff-rowIndependent}, an
algorithmic decision procedure for membership in the planar
rigidity matroid $\mathcal{R}(V)$
(\cref{def:rigidityMatroid}), specialised to spanning
subgraphs of edge count $2|V| - 3$ to decide the Laman property
(\cref{def:isLaman}). This is the Jacobs--Hendrickson
$(2, 3)$-pebble game~\cite{jacobsHendrickson1997} as a
formally-verified, end-user-executable program.
