\section{Introduction}
\label{sec:intro}

This blueprint accompanies the
\href{https://github.com/bryangingechen/CombinatorialRigidity}{\texttt{CombinatorialRigidity}}
Lean~4 / mathlib4 formalization. The goal is a proof of \textbf{Laman's
theorem}~\cite{laman1970}: for $n \ge 2$, a graph $G$ on $n$ vertices is generically
rigid in the plane if and only if it contains a spanning subgraph $H$ with
$|E(H)| = 2n - 3$ such that every subgraph $H[S]$ on $k \ge 2$ vertices
satisfies $|E(H[S])| \le 2k - 3$. Such an $H$ is called a \emph{Laman
graph}, or \emph{minimally rigid graph in the plane}.

The Lean development follows the \emph{Henneberg-construction} route: we
characterize Laman graphs inductively via two combinatorial moves (Type~I
adds a vertex of degree~2; Type~II splits an edge), prove the moves
preserve generic rigidity, and combine these facts to deduce the ``Laman
implies generically rigid'' direction. This $\Leftarrow$ direction is the
classical strategy of Tay and Whiteley~\cite{tayWhiteley1985}, in the
modern presentation of Jord\'an~\cite{jordan2016}~\S 2.2. The converse
direction goes through the rigidity matroid. For background on
combinatorial rigidity, see
\cite{graverServatiusServatius1993,jordan2016}.

\subsection*{Phase plan}

The Lean code is organized into phases, tracked in
\href{https://github.com/bryangingechen/CombinatorialRigidity/blob/master/ROADMAP.md}{\texttt{ROADMAP.md}}.
Phases~1--22b and the Grassmann--Cayley meet sub-phase Phase~21a are
complete (see the phase bullets below); \textbf{Phase~21} is
GREEN-modulo-21b (all genericity-free content green, with the shared
genericity device Claim~6.4/6.9 scoped out into sub-phase Phase~21b),
and \textbf{Phase~21b} delivered that device. A math-first feasibility
pass then re-scoped the algebraic induction's realization \emph{producers}
to Phases~22--23, so the Theorem~5.5 capstone \texttt{thm:theorem-55} and
\texttt{prop:rigidity-matrix-prop11} remain green-modulo-22--23 (they await
the producers and the genericity-free \texttt{hub} partition brick).
\textbf{Phase~22} (the realization layer) is sub-lettered: \textbf{Phase~22a}
(Case~I realization) is complete, built via the block-triangular reframe and
shipped green-modulo KT Claim~6.4 (\cref{lem:claim-6-4}); \textbf{Phase~22b}
\emph{fully discharged} that obligation (\cref{lem:claim-6-4} now green,
KT~\S6.2/\S5.1: the exterior-column projection loses zero rank on a rigid block, a
collapse-relabel row reproduction carries the surviving independence back at a
degenerate placement, and a motive strengthening discharges the selector alignment),
so \cref{lem:case-I-realization} is now fully green; \textbf{Phase~22c} (Case~III at
$d=3$, \emph{stratum~1} --- KT~\S6.4.1, Lemma~6.10, the eq.~(6.12) $+(D-1)$
placement; the conjecture's crux, which is itself multi-phase) landed its stratum-1
brick, opened \emph{design-pass-first}; \textbf{Phase~22d} (the \emph{D-candidate
crux} --- KT~\S6.4.1, Lemma~6.10 strata~2--3, Claims~6.11/6.12, the conjecture's
hardest single argument) is in progress, opened \emph{design-pass-first}; and the
$d=3$ assembly is the deferred, unlettered further cut.
\textbf{Phase~17} opened the longer-horizon
molecular-conjecture program (Phases~17--26;
Katoh--Tanigawa~\cite{katohTanigawa2011}), \textbf{Phase~18}
builds the genuine panel-hinge rigidity matrix on top of it,
\textbf{Phase~19} builds the matroid $M(\tilde G)$,
the $D$-deficiency, and the $k$-degree-of-freedom combinatorics,
\textbf{Phase~20} develops the combinatorial induction (graph
operations and Theorem~4.9), and \textbf{Phase~21} lands the
algebraic induction (Theorem~5.5 base, Cases~I \& II, and the analytic
half of Proposition~1.1).
The first six
phases formalize Laman's theorem along the Henneberg route described
above. Phases~7--8 package the planar rigidity matroid as a mathlib
\texttt{Matroid}: first combinatorially, via the Lov\'asz--Yemini
identification with the $(2, 3)$-count matroid, and then as a linear
matroid at a generic placement. Phases~9--11 develop the
$(k, \ell)$-pebble game of Lee--Streinu~\cite{leeStreinu2008} into an
executable, machine-checked decision procedure for
$(k, \ell)$-sparsity in the matroidal regime $\ell < 2k$, returning a
verdict that carries an inline correctness certificate on both the
accept and reject branches. Phases~12--15 build, bottom-up, toward
Tay's body-bar theorem: the abstract matroid-union machinery
(Phase~12), Tutte--Nash-Williams tree-packing (Phase~13), the
$k$-frame matroid as a cycle-matroid union (Phase~14), and Tay's
theorem itself (Phase~15). Phase~16 extends the body-bar theorem to
the body-hinge / panel-hinge Tay--Whiteley theorem via a
hinge-to-body-bar reduction. Phase~17 opens the
molecular-conjecture program with the Grassmann--Cayley extensor-algebra
layer, Phase~18 builds the genuine panel-hinge rigidity matrix
$R(G,p)$ on it, Phase~19 builds its matroidal
substrate $M(\tilde G)$, the $D$-deficiency, and the $k$-dof
combinatorics, Phase~20 develops the combinatorial induction ---
the graph operations and Katoh--Tanigawa's Theorem~4.9 --- and
Phase~21 opens the algebraic induction, realizing that reduction at
the rigidity-matrix rank (Theorem~5.5 base and Cases~I \& II). The
per-phase breakdown:

\begin{enumerate}
  \item \textbf{Sparsity.} Basic $(k,\ell)$-sparsity and $(k,\ell)$-tightness
        for arbitrary $k, \ell$. (Files: \texttt{EdgesIn.lean},
        \texttt{Sparsity.lean}.)
  \item \textbf{Laman graphs.} The $(2,3)$-tight specialization, the
        $K_2$ base case, and degree facts feeding the Henneberg induction.
        (File: \texttt{Laman.lean}.)
  \item \textbf{Henneberg moves.} Type~I and Type~II moves, Laman-preservation
        in both directions, and the existence of an inverse move at every
        non-base Laman graph. (File: \texttt{Henneberg.lean}.)
  \item \textbf{Frameworks.} Frameworks, the rigidity map, infinitesimal
        rigidity, generic rigidity. (File: \texttt{Framework.lean}.)
  \item \textbf{Laman's theorem ($\Leftarrow$).} Per-move rigidity
        preservation and the induction. (Files: \texttt{HennebergRigidity.lean},
        \texttt{LamanTheorem.lean}.)
  \item \textbf{Laman's theorem ($\Rightarrow$).} The companion direction
        via the planar rigidity matroid (Lov\'asz--Yemini's easy
        identification of row-independence with $(2,3)$-sparsity).
        (Files: \texttt{RigidityMatroid.lean}, \texttt{LamanTheorem.lean}.)
  \item \textbf{Lov\'asz--Yemini matroid identification.} The harder
        converse: $(2,3)$-sparse $\Rightarrow$ row-independent at some
        placement; together with Phase~6 this gives the
        Lov\'asz--Yemini~\cite{lovaszYemini1982} row-LI $\Leftrightarrow$
        $(2,3)$-sparse identification. Packaged as a mathlib
        \texttt{Matroid} via the general $(k, \ell)$-count matroid in
        the matroidal regime $\ell < 2k$
        (Whiteley~\cite{whiteley1996}, Lee--Streinu~\cite{leeStreinu2008}).
        (Files: \texttt{CountMatroid.lean}, \texttt{MatroidIdentification.lean}.)
  \item \textbf{Linear-matroid framing.} The planar rigidity matroid
        as a linear matroid via \texttt{Matroid.ofFun} (from the
        \texttt{apnelson1/Matroid} library) of the rigidity-row
        function at a uniformly generic placement, identified with the
        combinatorial $(2,3)$-count matroid of Phase~7
        (Lov\'asz--Yemini, linear-matroid form).
        (File: \texttt{LinearRigidityMatroid.lean}.)
  \item \textbf{Pebble game.} The basic $(k, \ell)$-pebble game of
        Lee--Streinu~\cite{leeStreinu2008} (generalising the original
        $(2, 3)$ algorithm of Jacobs--Hendrickson~\cite{jacobsHendrickson1997})
        and its correctness theorem in certificate form, in the
        matroidal regime $\ell < 2k$ matching Phase~7. Preceded by a
        verified-DFS warmup. (Files:
        \texttt{Search/DFS.lean}, \texttt{PebbleGame/\{Basic,Algorithm,Correctness\}.lean}.)
  \item \textbf{Executable pebble game.} A computable wrapper around
        the Phase~9 algorithm, yielding \texttt{Decidable} instances
        for $\mathrm{IsSparse}$, $\mathrm{IsTight}$, and
        $\mathrm{IsLaman}$ that run in polynomial time;
        \texttt{\#eval}-able worked examples on $\mathrm{Fin}\,n$
        graphs; and a \texttt{lake exe pebble-game} command-line tool.
        (Files: \texttt{PebbleGame/\{Exec,Examples\}.lean},
        \texttt{Main.lean}.)
  \item \textbf{Witness extraction.} The pebble game returns a
        verdict carrying an inline, machine-checked certificate on
        every branch: the accepting partial orientation $D$ when the
        input is $(k, \ell)$-sparse, and a blocking vertex subset
        $V'$ with $\bigl|\edgesIn{V'}\bigr| > k|V'| - \ell$ when it
        is not. The command-line tool prints this witness data
        alongside the classification, making its verdict externally
        checkable. (Files: \texttt{Search/DFS.lean},
        \texttt{PebbleGame/\{Basic,Algorithm,Correctness,Exec\}.lean},
        \texttt{Main.lean}.)
  \item \textbf{Matroid foundations: submodular functions and
        matroid union.} The abstract-matroid prerequisites of the
        body-bar route, formalized \emph{locally}: the
        matroid-from-submodular-function construction and polymatroid
        rank formula (Edmonds~\cite{edmonds1970}), the matroid-union
        theorem (Nash-Williams~\cite{nashWilliams1966} / Edmonds),
        Rado's theorem (Rado~1942,~\cite[Thm 11.2.2]{oxley2011}), and
        Edmonds' matroid-partition rank formula~\cite{edmonds1965}. The
        Lean is ported from Peter Nelson's \texttt{apnelson1/Matroid}
        package (Apache-2.0; the same package supplying
        \texttt{Matroid.ofFun} and \texttt{Graph.cycleMatroid}), whose
        shelved \texttt{WIP/\{Submodular,Union\}.lean} carry complete
        proofs, rebased onto the package's current
        \texttt{FiniteCircuitMatroid} constructor. (Chapter:
        \cref{sec:matroid-union}; files:
        \texttt{CombinatorialRigidity/Matroid/}.)
  \item \textbf{Tutte--Nash-Williams tree-packing.} Specializing the
        matroid-partition theorem to $k$ copies of the cycle matroid
        recovers the tree-packing theorem~\cite{tutte1961,nashWilliams1961}:
        a multigraph is the edge-disjoint union of $k$ forests iff it
        is $(k, k)$-sparse. Introduces a \texttt{Graph}-native
        $(k, \ell)$-sparsity predicate. (File:
        \texttt{BodyBar/TreePacking.lean}.)
  \item \textbf{The $k$-frame matroid as a cycle-matroid union.}
        Whiteley's Theorem~1~\cite{whiteley1988}: the generic
        $k$-frame matroid on a multigraph (a linear matroid via
        \texttt{Matroid.ofFun} over indeterminate coefficients)
        equals the $k$-fold union of the cycle matroid, restricted to
        the bar set $E(G)$. (File: \texttt{BodyBar/KFrame.lean}.)
  \item \textbf{Body-bar Tay theorem.} For $n \ge 2$ and
        $d = n(n+1)/2$, a multigraph $G$ carries an independent body-bar
        framework in $\mathbb{R}^n$ iff $G$ is $(d,d)$-sparse, and an
        isostatic one iff $(d,d)$-tight --- equivalently the
        edge-disjoint union of $d$ forests (resp.\ spanning trees)
        (Tay~1984~\cite{tay1984}, \cref{thm:tay-witness}). Existence is
        witnessed by specializing two-extensor row coefficients to
        standard-basis Pl\"ucker coordinates on a tree packing; the
        converse is the body-bar Lov\'asz--Yemini rank-upper-bound. The
        phase ships the existence-of-realization form, deferring
        Whiteley's full lift via irreducible-variety machinery
        (Proposition~6 in~\cite{whiteley1988}). (Chapter:
        \cref{sec:body-bar}; files:
        \texttt{BodyBar/\{Framework,TayTheorem\}.lean}.)
  \item \textbf{Body-hinge Tay--Whiteley theorem.} The body-hinge
        (panel-hinge) theorem in $n$-space (Tay~1989~\cite{tay1989},
        Whiteley~\cite{whiteley1988}), existence-of-realization form,
        \emph{via the matroid-union reduction to Phase~15}. A hinge ---
        an $(n-2)$-dimensional affine subspace --- constrains all but
        one of the $\delta = n(n+1)/2$ relative screw freedoms of the
        two bodies it joins, so it behaves like a bundle of
        $\delta - 1$ coincident body-bars; a body-hinge framework on
        $G$ is the induced body-bar framework on $(\delta-1)\cdot G$
        (each hinge replaced by $\delta - 1$ parallel bars), and the
        characterization --- $G$ carries an independent (resp.\
        isostatic) body-hinge framework iff $(\delta-1)\cdot G$ is
        $(\delta,\delta)$-sparse (resp.\ tight) --- reduces node-for-node
        to \cref{thm:tay-witness} on $(\delta-1)\cdot G$. The
        $(\delta-1)\cdot G$ device is the multiplied graph of the
        molecular conjecture~\cite{katohTanigawa2011}. (Chapter:
        \cref{sec:body-hinge}; file: \texttt{BodyBar/BodyHinge.lean}.)
  \item \textbf{Grassmann--Cayley extensor algebra.}
        Opens the longer-horizon molecular-conjecture program
        (Phases~17--26, Katoh--Tanigawa~\cite{katohTanigawa2011}) with
        its first stratum: the extensor algebra of
        \cite[\S2.1]{katohTanigawa2011}. Homogeneous coordinatization,
        extensors as decomposable elements of $\bigwedge^{j} \R^{d+1}$,
        the join $\vee$, Plücker coordinates, and the affine-subspace
        extensor $C(\cdot)$, culminating in Lemma~2.1: the
        $D = \binom{d+1}{2}$ many $(d-1)$-extensors of $d+1$ affinely
        independent points are linearly independent --- the
        linear-algebra foundation the conjecture's hardest case bottoms
        out on. (Chapter: \cref{sec:molecular}; file:
        \texttt{Molecular/Extensor.lean} --- all \S2.1 nodes green.)
  \item \textbf{Panel-hinge rigidity matrix.} Stratum~2 of the
        molecular program (Katoh--Tanigawa~\cite[\S2.2--2.4]{katohTanigawa2011}):
        the \emph{genuine} panel-hinge rigidity matrix $R(G,p)$. A
        body-hinge framework assigns a $(d-2)$-dimensional affine hinge
        to each edge; its supporting $(d-1)$-extensor $C(p(e))$ (Phase~17)
        constrains the $D$-dimensional screw centers by
        $S(u) - S(v) \in \operatorname{span} C(p(e))$. The phase builds
        the null space $Z(G,p)$ of the block matrix (basis-free, on the
        degree-$k$ graded screw space $\bigwedge^k \R^{k+2}$ of dimension
        $D$), the $D$ trivial motions and the codimension form of
        $\operatorname{rank} \le D(|V|-1)$, and the rank Lemmas~5.1--5.3
        --- all green: the full rank-form substrate the algebraic
        induction runs on. The reconciliation with the body-hinge
        existence form of Phase~16 (Tay--Whiteley's Proposition~1.1,
        \cref{thm:body-hinge-tay}) is a top-of-DAG corollary depending on
        the analytic generic-rank theorem; it is presented with the
        algebraic induction (Phase~21+). (Chapter:
        \cref{sec:molecular-rigidity-matrix}; file:
        \texttt{Molecular/RigidityMatrix.lean}.)
  \item \textbf{The matroid $M(\tilde G)$, deficiency, and $k$-dof
        graphs.} Stratum~3 of the molecular program
        (Katoh--Tanigawa~\cite[\S2.5, \S3]{katohTanigawa2011}): the
        matroidal substrate of the conjecture's algebraic induction. The
        matroid $M(\tilde G)$ is the $(D,D)$-count matroid of the
        multiplied graph $\tilde G = (D-1)\cdot G$ at the boundary regime
        $\ell = 2k = D$ --- the $D$-fold graphic-matroid union of
        Phases~13--14 with Tutte--Nash-Williams, not the matroidal-regime
        count matroid of Phase~7. The phase builds the $D$-deficiency
        $\operatorname{def}(\tilde G)$, the $k$-degree-of-freedom and
        minimal-$k$-dof hierarchy, rigid subgraphs, the structural lemmas
        (KT 3.1, 3.3, 3.4), and the bridge
        $\operatorname{def}(\tilde G) = \operatorname{corank} M(\tilde G)$
        (Jackson--Jord\'an~\cite{jacksonJordan2009}) --- the matroidal
        half of the Phase~16 reconciliation --- all green. (Chapter:
        \cref{sec:molecular-deficiency}; file:
        \texttt{Molecular/Deficiency.lean}.)
  \item \textbf{Combinatorial induction.} Stratum~4 of the molecular
        program (Katoh--Tanigawa~\cite[\S3.4--3.5, \S4]{katohTanigawa2011}):
        the combinatorial half of the proof. The graph operations on
        \texttt{Graph $\alpha$ $\beta$} --- vertex removal, splitting-off
        $G_v^{ab}$ at a degree-2 vertex, its inverse edge-splitting, and
        rigid-subgraph contraction --- the degree-of-freedom tracking
        lemmas (4.3--4.8), and the capstone \textbf{Theorem~4.9}: every
        minimal $0$-dof-graph reduces to the two-vertex double edge by
        splitting-off and rigid-subgraph contraction, the combinatorial
        skeleton the algebraic induction (Phases~21--23) realizes at the
        rigidity-matrix rank. The formalization surfaced that KT Lemma~4.1
        (the forest surgery) is over-quantified with a balanced-packing
        gloss in its proof (routed around via a deficiency-count
        argument), and that KT's iterated fundamental-circuit swaps are
        bypassed by partition-count / rank-count comparisons through
        \cref{thm:def-eq-corank}; the forest-surgery core (KT 4.1/4.2) is
        off the Theorem-4.9 critical path. \emph{Complete.} (Chapter:
        \cref{sec:molecular-induction}; file:
        \texttt{Molecular/Induction/}.)
  \item \textbf{Grassmann--Cayley meet / projective-duality foundations
        (Phase~21a).} A prerequisite sub-phase of the algebraic induction,
        inserted by the panel-coplanarity re-scope
        (Katoh--Tanigawa~\cite[\S2.1]{katohTanigawa2011}, dual half):
        the \emph{meet} (regressive product) and the projective-duality
        dictionary on the concrete carrier $\R^{k+2}$, the dual of the
        Phase-17 join. Builds the top-power volume-form iso, the
        exterior-power duality iso $\bigwedge^{j} V^{*} \cong (\bigwedge^{j}
        V)^{*}$, the complement iso $\bigwedge^{j} V \cong \bigwedge^{N-j} V$
        (from the nondegenerate perfect wedge pairing), and the meet itself
        as the Grassmann--Cayley dual of the join. Metric-free (orientation
        only, no Hodge star). \emph{Complete.} (Chapter:
        \cref{sec:molecular-meet}; file: \texttt{Molecular/Meet.lean}.)
  \item \textbf{Algebraic induction: Theorem~5.5 base and Cases~I \& II.}
        Stratum~5 of the molecular program
        (Katoh--Tanigawa~\cite[\S5, \S6.1--6.3]{katohTanigawa2011}): the
        \emph{algebraic} half of the proof, which realizes the Phase~20
        combinatorial reduction at the rigidity-matrix rank. The phase
        lands Theorem~5.5 --- every minimal $k$-dof-graph $G$ with
        $|V| \ge 2$ has a panel-hinge realization with
        $\operatorname{rank} R(G,p) = D(|V|-1) - k$ --- and discharges
        its base case ($|V| = 2$), Case~I (a proper rigid subgraph, by
        rigid-subgraph contraction and block-triangular gluing through the
        pin-a-body Lemma~5.1), Case~II ($k > 0$, splitting off a
        reducible degree-2 vertex --- the panel-hinge analogue of
        Whiteley's bar-joint $1$-extension), and the analytic half of
        Proposition~1.1 ($\operatorname{rank} R(G,p) = D(|V|-1) -
        \operatorname{def}(\tilde G)$, the rank/deficiency reconciliation;
        matroidal half green from Phase~19). The induction is driven by
        the same reduction dichotomy as Theorem~4.9
        (\cref{thm:minimal-kdof-reduction}). The shared analytic
        device --- the genericity argument (Claim~6.4/6.9) --- is scoped
        out into its own focused sub-phase (Phase~21b) and enters each
        consuming node as an explicit hypothesis, so the surrounding
        reductions are fully formal modulo that one device. The
        cycle-realization Lemma~5.4~\cite{crapoWhiteley1982} has its four
        Lean pieces green with the same genericity assembly cited. The
        crux Case~III ($k = 0$, no proper rigid subgraph) is deferred to
        Phases~22--23. \emph{Complete (GREEN-modulo-21b).} (Chapter:
        \cref{sec:molecular-algebraic-induction}; file:
        \texttt{Molecular/AlgebraicInduction/}.)
  \item \textbf{Genericity device (Claim~6.4/6.9) (Phase~21b).} The
        analytic sibling of the Phase-21a meet sub-phase, scoped out of
        Phase~21: the panel-coordinate parametrization of $R(G,p)$
        (entries as polynomials in the per-vertex normals) and the
        generic-max-rank argument
        (Katoh--Tanigawa~\cite[\S6.1, \S6.3]{katohTanigawa2011}), the
        shared analytic crux of Cases~I \& II, Theorem~5.5,
        Proposition~1.1, and the cycle assembly. Delivered green, applied
        to the varying panel family (the B0 row coordinatization), with the
        genericity-free accounting iffs and the $V(G)$-relative count
        bridges. A math-first feasibility pass then re-scoped the
        realization \emph{producers} to Phases~22--23 (the reducible-vertex
        split is Katoh--Tanigawa's Case~III, one row short for $k=0$; the
        rigid-subgraph splice is full-rank but research-shaped geometry).
        \emph{Complete.} (Node:
        \cref{lem:genericity-device}; file:
        \texttt{Molecular/AlgebraicInduction/}.)
  \item \textbf{Realization layer (Phase~22, sub-lettered).}
        The Theorem~5.5 case \emph{producers} the genericity device feeds,
        re-scoped out of Phase~21b and split into sub-phases (sub-lettering
        22a, 22b, \dots\ keeps the integer phase numbers 23--26 stable). A
        structural-edit phase extending
        \cref{sec:molecular-algebraic-induction} rather than opening a new
        chapter.
        \begin{itemize}
          \item \textbf{Phase~22a --- Case~I realization} (\emph{complete}):
                the full-rank rigid-subgraph splice
                (Katoh--Tanigawa~\cite[\S6.2]{katohTanigawa2011}: the
                graph/matroid contraction-minimality bridge
                \cref{lem:rigidContract-isMinimalKDof} and the composer
                \cref{lem:case-I-realization}). The composer is built via the
                \emph{block-triangular reframe} --- KT eq.~(6.3)'s rank-addition
                over one common framework, routed through the genericity device's
                independent-row count (the $H$-block edges $\sqcup$ the surviving
                edges made independent by the exterior-column projection), not a
                common-seed splice --- and shipped green-modulo a single dischargeable
                hypothesis, KT Claim~6.4 (\cref{lem:claim-6-4}), discharged in
                Phase~22b.
          \item \textbf{Phase~22b --- KT Claim~6.4} (\emph{complete}): \emph{fully
                discharges} the single obligation Phase~22a left green-modulo
                (\cref{lem:claim-6-4} now green, the surviving block of
                $G \setminus E(H)$ projected to the surviving columns
                $V(G) \setminus V(H)$ keeping its independent rank at the generic
                locus; KT~\S6.2/\S5.1, eqs.~(6.5)/(6.9)), so
                \cref{lem:case-I-realization} is now fully green. The exterior-column
                projection loses zero rank on a rigid block (the genuine crux: $Z
                \sqcup \mathrm{range}(\mathrm{extProj}) = \top$, whose one
                real-content input is the rigid-block pin-count), a collapse-relabel
                row reproduction carries the surviving independence back at the
                degenerate placement (KT eq.~(6.7)), and a motive strengthening
                discharges the selector alignment; an assembly producer composes these
                into the rank-transport witness from the contraction's generic
                induction hypothesis.
          \item \textbf{Phase~22c --- Case~III at $d=3$, stratum~1}
                (\emph{stratum-1 complete};
                Katoh--Tanigawa~\cite[\S6.4.1]{katohTanigawa2011}, Lemma~6.10 --- the
                conjecture's crux and the single largest proof in KT): the
                reducible-vertex producer at $d=3$ ($k=0$). Case~III at $d=3$ is
                \emph{multi-phase}; Phase~22c claimed only the first of three
                difficulty strata --- the degenerate eq.~(6.12) placement on the
                green row infrastructure, which is one row short of full rank
                ($\mathrm{rank}\ge D(|V|-1)-1$, green + axiom-clean). Opened
                \emph{design-pass-first}: on a layer-level design recon (five passes),
                not a build. \cref{lem:case-II-realization,lem:case-III} stay red ---
                Phase~22c lands the $+(D-1)$ brick toward them.
          \item \textbf{Phase~22d --- the D-candidate crux}
                (\emph{in progress}; KT~\cite[\S6.4.1]{katohTanigawa2011}, Lemma~6.10
                strata~2--3 --- the conjecture's hardest single argument): the
                \emph{missing $+1$ row} that lifts Phase~22c's $D(|V|-1)-1$ brick to
                full $D(|V|-1)$. Stratum~2 is Claim~6.11's redundant $(ab)i^*$-row (the
                combinatorial$\leftrightarrow$linear conversion wiring
                $M(\tilde G_v^{ab})$ to the row matroid of $R$, the single
                highest-risk node in Phases~22--23); stratum~3 is the
                candidate-normal-form $+$ Claim~6.12 extensor-span contradiction via
                the extensor-independence Lemma~2.1
                (\cref{lem:extensor-independence}). Opened \emph{design-pass-first},
                with the verdict that Claim~6.11 takes the axiomatize-as-hypothesis
                fallback (not buildable from the green $M(\tilde G)$ machinery) while
                Claim~6.12 is de-risked on the green Lemma~2.1.
                \cref{lem:case-II-realization,lem:case-III} stay red.
          \item \textbf{The $d=3$ assembly} (\emph{planning}, deferred and
                \emph{unlettered}): the genericity-free \texttt{hub} partition brick of
                \cref{prop:rigidity-matrix-prop11} (Track-independent) plus the
                \cref{thm:theorem-55} flip and the Case-I wiring; it gets a letter only
                when its turn comes (the crux may itself split).
        \end{itemize}
        (File: \texttt{Molecular/\{Induction,AlgebraicInduction\}.lean}.)
\end{enumerate}

\subsection*{Reading this blueprint}

Statements that are formalized in Lean carry a green tick and a
\texttt{\textbackslash lean\{\}} pointer linking back to the API
documentation. The \emph{dependency graph view} (linked from the
navigation bar) is the best place to see at a glance how the chapters
fit together; all Phase~1--16 nodes are green: the main Laman line
(\cref{sec:laman-theorem}), the matroid identifications
(\cref{sec:rigidity-matroid,sec:count-matroid}), the verified-DFS
warmup (\cref{sec:dfs}), the pebble-game chapter
(\cref{sec:pebble-game}), the executable pebble-game chapter
(\cref{sec:executable}), the matroid-union / Edmonds-partition
chapter (\cref{sec:matroid-union}), the full body-bar chapter
(\cref{sec:body-bar}) --- Tutte--Nash-Williams tree-packing, the
$k$-frame matroid as a cycle-matroid union, and Tay's body-bar theorem
itself (\cref{thm:tay-witness}) --- and the body-hinge chapter
(\cref{sec:body-hinge}): the hinge-to-body-bar reduction and the
body-hinge Tay--Whiteley theorem (\cref{thm:body-hinge-tay}). With
Phases~17--19 complete the Phase~1--19 dependency graph is green. The
molecular-conjecture program (\cref{sec:molecular}, opened as Phase~17)
runs forward-mode: its extensor-algebra section is fully green,
culminating in \cref{lem:extensor-independence}
(Katoh--Tanigawa~\cite{katohTanigawa2011} Lemma~2.1). The
panel-hinge rigidity-matrix dep-graph
(\cref{sec:molecular-rigidity-matrix}, Phase~18) --- its rank-matrix
skeleton, trivial-motion layer, and three rank lemmas --- is fully
green. The deficiency chapter (\cref{sec:molecular-deficiency},
Phase~19) is also fully green: the matroid $M(\tilde G)$, the
$D$-deficiency, the $k$-dof hierarchy, the structural lemmas, and the
$\operatorname{def} = \operatorname{corank}$ bridge
(\cref{thm:def-eq-corank}) --- the matroidal half of the Phase~16
reconciliation. The whole Phase~1--19 dependency graph is therefore
green. \textbf{Phase~20} (\cref{sec:molecular-induction}) is also
complete: the combinatorial-induction chapter --- the graph operations,
the degree-of-freedom tracking, and the capstone Theorem~4.9
(\cref{thm:minimal-kdof-reduction}) --- is green on its critical path
(the forest-surgery nodes, KT Lemmas 4.1/4.2, stay red as a deferred
off-critical-path TODO). The whole Phase~1--20 dependency graph is
therefore green on the Theorem-4.9 critical path. \textbf{Phase~21a}
(\cref{sec:molecular-meet}) --- the Grassmann--Cayley meet /
projective-duality foundations sub-phase inserted by the
panel-coplanarity re-scope --- is complete: the top-power volume-form
iso, the exterior-power duality iso, the complement iso, and the meet
(\cref{def:meet}) are all green. \textbf{Phase~21}
(\cref{sec:molecular-algebraic-induction}) settled the algebraic
induction's accounting, and \textbf{Phase~21b} delivered the shared
\emph{genericity device}. Green: the panel layer
(\cref{def:panel-support-extensor,def:panel-hinge-framework}), the
regime-agnostic rank/structure nodes, all four Lean pieces of the
cycle-realization Lemma~5.4 (\cref{lem:cycle-realization-base,%
lem:cycle-realization-dim-bound,lem:cycle-realization-rigid,%
lem:panel-support-extensor-independence,%
lem:exists-independent-panel-extensor}), the genericity device
(\cref{lem:genericity-device}) applied to the varying panel family via the
B0 row coordinatization (\cref{lem:rows-polynomial-in-normals}), the
$\alpha$-independent \emph{$V(G)$-relative rank} accounting iffs
(\cref{lem:case-I,lem:case-II}, against the motive
$\operatorname{rank} R(G,p) = D(|V(G)|-1)$, \cref{def:rank-hypothesis}) and
base case (\cref{lem:theorem-55-base}), the $V(G)$-relative count bridges,
and the Case-I splice glue (\cref{lem:case-I-splice-seed}). A math-first
feasibility pass (against Katoh--Tanigawa~\S6.2--6.3) then re-scoped the
realization \emph{producers} to Phases~22--23: the reducible-vertex split
producer (\cref{lem:case-II-realization}) is KT's Case~III --- the
degenerate eq.~(6.12) placement is one row short of full rank for $k=0$ and
needs the Lemma~6.10/6.13 redundant-edge row --- and the rigid-subgraph
splice producer (\cref{lem:case-I-realization}, via
\cref{lem:claim-6-4}) reaches full rank (KT~\S6.2) but is
research-shaped boundary-panel geometry. \emph{Green-modulo-22--23
(awaiting the producers and the genericity-free \texttt{hub} partition
brick):} the capstone \cref{thm:theorem-55} and the generic-rank
reconciliation \cref{prop:rigidity-matrix-prop11}; plus Case~III
(\cref{lem:case-III}), all in Phases~22--23. \textbf{Phase~22}
(\cref{sec:molecular-algebraic-induction}) is the realization layer,
sub-lettered: \textbf{Phase~22a} (complete) builds the Case~I splice producer
(\cref{lem:rigidContract-isMinimalKDof,lem:case-I-realization}), the composer
\cref{lem:case-I-realization} via the block-triangular reframe, shipped
\emph{green-modulo} KT Claim~6.4 (\cref{lem:claim-6-4}). \textbf{Phase~22b} (complete)
\emph{fully discharges} \cref{lem:claim-6-4} --- the exterior-column projection loses
zero rank on a rigid block, a collapse-relabel row reproduction carries the surviving
independence back at a degenerate placement (KT eq.~(6.7)), and a motive strengthening
discharges the selector alignment, all composed into a rank-transport witness from the
contraction's generic induction hypothesis --- so \cref{lem:case-I-realization} is now
fully green. \textbf{Phase~22c} (stratum-1 complete, opened design-pass-first) is
Case~III at $d=3$, stratum~1 (KT~\S6.4.1, Lemma~6.10, the eq.~(6.12) $+(D-1)$ placement,
the conjecture's crux). Case~III at $d=3$ is multi-phase; Phase~22c claimed only the
first of three difficulty strata (the eq.~(6.12) placement, green + axiom-clean from the
green row infra). \textbf{Phase~22d} (in progress, opened design-pass-first) is the
\emph{D-candidate crux} (KT~\S6.4.1, Lemma~6.10 strata~2--3, Claims~6.11/6.12) --- the
missing $+1$ row that lifts Phase~22c's $D(|V|-1)-1$ brick to full $D(|V|-1)$; its
verdict is that Claim~6.11 takes the axiomatize-as-hypothesis fallback while Claim~6.12
is de-risked on the green Lemma~2.1. The reducible-vertex producer nodes
(\cref{lem:case-II-realization,lem:case-III}) stay red until that construction lands.
The $d=3$ assembly --- the genericity-free \texttt{hub} partition brick and the
\cref{thm:theorem-55} flip --- is the deferred, unlettered further cut. Phases~23--26
remain to be opened.
