\section{Grassmann--Cayley meet and projective duality}
\label{sec:molecular-meet}

This chapter is the \textbf{Phase~21a} dep-graph: the \emph{dual half}
of the Grassmann--Cayley algebra of
Katoh--Tanigawa~\cite[\S2.1]{katohTanigawa2011}, a prerequisite
sub-phase of the algebraic induction (\cref{sec:molecular-algebraic-induction})
inserted by the panel-coplanarity re-scope. Where
\cref{sec:molecular-homog} built the \emph{join} (the progressive
product $A \vee B$, the symbolic exterior product on
$\bigwedge[\R] \R^{k+2}$) and a coordinatized
PlĂĽcker bridge, this chapter builds the \emph{meet} (the regressive
product) and the projective-duality dictionary on which it rests, on the
same concrete carrier. The meet is the device the panel-coplanarity
layer, the cycle-realization Lemma~5.4 (Crapo--Whiteley~\cite{crapoWhiteley1982}),
and the projective invariance of Phase~25 all consume.

The construction is \textbf{metric-free}: projective geometry needs no
inner product, only the top-power volume form (an orientation). Write
$N = k+2$ and $V = \R^{N}$ for the homogeneous-coordinate carrier
of $\cref{sec:molecular-homog}$; the top graded piece
$\bigwedge^{N} V$ is free of rank $\binom{N}{N} = 1$, hence
canonically $\cong \R$.

\medskip
\noindent\emph{Status.} In progress. The top-power volume-form iso
(\cref{def:meet-top-equiv}) and the duality-pairing iso
(\cref{def:meet-pairing-dual}) are green, formalized in
\texttt{Molecular/Meet.lean} (with the general facts mirrored under
\texttt{Mathlib/LinearAlgebra/ExteriorPower/Basis.lean}). The complement
iso and the regressive product itself remain to be formalized.

\begin{definition}[Top-power volume form $\bigwedge^{N} V \cong \R$]
  \label{def:meet-top-equiv}
  \lean{CombinatorialRigidity.Molecular.screwAlgebraTopEquiv,
    exteriorPower.topEquiv}
  \leanok
  The $N$-th exterior power of a free module of rank $N$ is free of rank
  $\binom{N}{N} = 1$, hence canonically isomorphic to the base ring. On
  the standard carrier $V = \R^{N}$ this is the \emph{volume
  form}: the linear isomorphism
  \[
    \mathrm{topEquiv} : \bigwedge^{N} V \xrightarrow{\ \sim\ } \R,
  \]
  sending the wedge $\mathbf{e}_1 \wedge \cdots \wedge \mathbf{e}_N$ of all
  standard basis vectors to $1$. It carries an orientation, not a metric:
  the meet is built from it without any inner-product structure.
\end{definition}
\begin{proof}
  \leanok
  Built from the standard-basis top-power basis
  $\mathrm{Module.Basis.exteriorPower}\,(\mathrm{Pi.basisFun}\ \R\ (\mathrm{Fin}\ N))$,
  whose index set $\mathrm{powersetCard}\,(\mathrm{Fin}\ N)\ N$ is a
  singleton (the full set is the unique $N$-element subset of an
  $N$-element type), composed with the unique-index evaluation
  $\mathrm{LinearEquiv.funUnique}$. The general fact (over any commutative
  ring, on $\mathrm{Fin}\ n \to R$) is the mirror lemma
  $\mathrm{exteriorPower.topEquiv}$; $\mathrm{screwAlgebraTopEquiv}$ is the
  $N = k+2$ specialization on the screw-algebra carrier of
  \cref{sec:molecular-homog}.
\end{proof}

\begin{definition}[Exterior-power duality iso $\bigwedge^{j} V^{*} \cong (\bigwedge^{j} V)^{*}$]
  \label{def:meet-pairing-dual}
  \lean{CombinatorialRigidity.Molecular.screwAlgebraPairingDualEquiv,
    exteriorPower.pairingDualEquiv}
  \leanok
  \uses{def:meet-top-equiv}
  For free finite $V$, mathlib's $\mathrm{exteriorPower.pairingDual}$ is a
  bare linear map $\bigwedge^{j} V^{*} \to (\bigwedge^{j} V)^{*}$; it
  upgrades to an isomorphism
  \[
    \mathrm{pairingDualEquiv} : \bigwedge^{j} V^{*}
      \xrightarrow{\ \sim\ } (\bigwedge^{j} V)^{*}.
  \]
  This is the \textbf{projective-duality dictionary entry} reused by the
  Crapo--Whiteley projective-invariance development of Phase~25.
\end{definition}
\begin{proof}
  \leanok
  Fix an ordered basis $b$ of $V$. The map $\mathrm{pairingDual}$ carries
  the exterior-power basis $\bigwedge^{j} b^{*}$ built from the dual basis
  onto the dual basis of the exterior-power basis $\bigwedge^{j} b$: the
  identification $b^{*} = b.\mathrm{coord}$ together with the coordinate
  computation $\mathrm{exteriorPower.basis\_coord}$ shows the image of each
  basis vector is the corresponding coordinate form. Hence
  $\mathrm{pairingDual}$ sends a basis to a basis and is an isomorphism;
  the mirror lemma $\mathrm{exteriorPower.pairingDualEquiv}$ packages this
  as a $\mathrm{Basis.equiv}$ and $\mathrm{coe\_pairingDualEquiv}$ identifies
  it with $\mathrm{pairingDual}$ in place.
  $\mathrm{screwAlgebraPairingDualEquiv}$ is the specialization at the
  standard basis of the screw-algebra carrier.
\end{proof}

\begin{definition}[Complement iso $\bigwedge^{j} V \cong \bigwedge^{N-j} V$]
  \label{def:meet-complement-iso}
  \uses{def:meet-top-equiv, def:meet-pairing-dual}
  The perfect wedge pairing
  $\bigwedge^{j} V \times \bigwedge^{N-j} V \to \bigwedge^{N} V
  \cong \R$ (the join of \cref{sec:molecular-homog} followed by the
  volume form of \cref{def:meet-top-equiv}) is nondegenerate for free $V$,
  inducing a linear isomorphism
  $\bigwedge^{j} V \cong \bigwedge^{N-j} V$. This is the genuinely
  new core of the chapter; the grade-swap sign is routed through the
  duality iso of \cref{def:meet-pairing-dual}.
\end{definition}

\begin{definition}[The regressive product (meet)]
  \label{def:meet}
  \uses{def:meet-complement-iso}
  The \emph{meet} (regressive product) $A \wedge B$ of two extensors is the
  composite
  $\mathrm{complementIso}^{-1} \circ \vee \circ (\mathrm{complementIso}
  \times \mathrm{complementIso})$ --- the join of \cref{sec:molecular-homog}
  transported across the complement iso of \cref{def:meet-complement-iso}.
  Its nonvanishing characterizes transversality (independence) of the two
  factors; geometrically, the meet of two hyperplane normals is the
  supporting extensor of their codimension-2 intersection, landing in the
  screw space of \cref{sec:molecular-rigidity-matrix}. This is the device
  the panel-coplanarity layer and Lemma~5.4 consume.
\end{definition}
