\section{Grassmann--Cayley extensor algebra}
\label{sec:molecular}

This chapter opens the \textbf{molecular-conjecture program}
(Phases~17--26), the project's longest-horizon target: the molecular
conjecture of Tay and Whiteley, proved by
Katoh--Tanigawa~\cite{katohTanigawa2011}. The program-level plan ---
target, five-strata architecture, and the phase breakdown --- is
described in the project's \texttt{notes/MolecularConjecture.md}; this
chapter is the forward-mode authoritative dep-graph and lemma index for
its \textbf{first phase}, Phase~17.

Phase~17 formalizes the \emph{Grassmann--Cayley / extensor algebra}
layer of Katoh--Tanigawa~\cite[\S2.1]{katohTanigawa2011}, the first of
the five strata. It is genuinely new linear algebra --- mathlib carries
the exterior algebra $\bigwedge[\R] M$ but not the Plücker/extensor
coordinatization, the join, or the affine-subspace--to-extensor map ---
and bottoms out the most difficult case of the conjecture's algebraic
induction (Case~III, Phases~22--23). The load-bearing target is
\textbf{Lemma~2.1} (\cref{lem:extensor-independence}): the
$D = \binom{d+1}{2}$ many $(d-1)$-extensors associated to $d+1$ affinely
independent points are linearly independent.

\medskip
\noindent\emph{Forward-mode chapter.} Following the convention for
Phases~6 onward, the nodes below are stated with their intended
\texttt{\textbackslash uses\{\}} dependency structure but \emph{without}
\texttt{\textbackslash lean\{\}} / \texttt{\textbackslash leanok}: the
dep-graph is red and \emph{is} the phase to-do list. Each node turns
green as its Lean lands, in the same commit that adds the pointer.

\subsection{Homogeneous coordinates and affine independence}
\label{sec:molecular-homog}

We work in $\R^{d}$ with points homogenized into $\R^{d+1}$. Throughout,
$D = \binom{d+1}{2} = \tfrac{d(d+1)}{2}$ is the dimension of the space of
infinitesimal screw motions of a rigid body in $\R^{d}$, matching the
$\delta = \mathrm{bodyBarDim}\,n$ of the body-bar / body-hinge chapters
(\cref{sec:body-bar,sec:body-hinge}).

\begin{definition}[Homogeneous coordinatization]
  \label{def:homogeneous-coords}
  \lean{CombinatorialRigidity.Molecular.homogenize}
  \leanok
  For a point $p \in \R^{d}$ write $\bar p = (p, 1) \in \R^{d+1}$ for its
  homogeneous coordinate vector. A finite family $p_1, \dots, p_k$ of
  points of $\R^{d}$ is \emph{affinely independent} when the homogenized
  family $\bar p_1, \dots, \bar p_k$ is linearly independent in
  $\R^{d+1}$.
\end{definition}

\begin{lemma}[Affine independence via homogenization and the top extensor]
  \label{lem:affine-indep-iff}
  \lean{CombinatorialRigidity.Molecular.affineIndependent_iff_linearIndependent_homogenize,
    CombinatorialRigidity.Molecular.affineIndependent_fin_iff_det_homogenize}
  \leanok
  \uses{def:homogeneous-coords}
  Points $p_1, \dots, p_k \in \R^{d}$ are affinely independent if and
  only if the homogenized family $\bar p_1, \dots, \bar p_k$ is linearly
  independent in $\R^{d+1}$ --- equivalently, once the join is in hand
  (\cref{def:join}), if and only if the join
  $\bar p_1 \vee \cdots \vee \bar p_k$ is nonzero. In particular $d+1$
  points are affinely independent iff their top extensor
  $\bar p_1 \vee \cdots \vee \bar p_{d+1}$ --- the determinant of the
  $(d+1)\times(d+1)$ matrix of their homogeneous coordinates --- is
  nonzero.
\end{lemma}
\begin{proof}
  \leanok
  Affine independence of $p_1, \dots, p_k$ means every coefficient
  family $w$ with $\sum_i w_i = 0$ and $\sum_i w_i p_i = 0$ vanishes; a
  vanishing combination $\sum_i w_i \bar p_i = 0$ of the homogenized
  vectors is exactly such a family (the last coordinate gives
  $\sum_i w_i = 0$, the first $d$ give $\sum_i w_i p_i = 0$), so affine
  independence is linear independence of the $\bar p_i$. Once the join is
  available it is the image in $\bigwedge^{k} \R^{d+1}$, nonzero exactly
  when the vectors are linearly independent. For $k = d+1$ the top
  exterior power is one-dimensional and the coordinate of the join is the
  determinant of the homogeneous coordinate matrix; that determinant is
  nonzero iff the rows are linearly independent.
\end{proof}

\subsection{Extensors, the join, and Plücker coordinates}
\label{sec:molecular-extensor}

\begin{definition}[Extensor]
  \label{def:extensor}
  A \emph{$j$-extensor} in $\R^{d+1}$ is a decomposable element of the
  $j$-th exterior power $\bigwedge^{j} \R^{d+1}$, i.e.\ one of the form
  $v_1 \wedge \cdots \wedge v_j$ for vectors $v_i \in \R^{d+1}$
  (Barnabei--Brini--Rota~\cite{barnabeiBriniRota1985}). We build the
  symbolic layer on mathlib's $\bigwedge[\R] \R^{d+1}$ and recover the
  geometric notion of the extensor of an affine subspace below.
\end{definition}

\begin{definition}[Join]
  \label{def:join}
  \uses{def:extensor}
  The \emph{join} $A \vee B$ of extensors is their exterior product
  $A \wedge B$ (the span of the corresponding subspaces, when they meet
  only at the origin). The join is alternating and vanishes whenever a
  vector is repeated, and is associative and graded-commutative as
  inherited from the exterior product.
\end{definition}

\begin{definition}[Plücker coordinates]
  \label{def:plucker-coords}
  \uses{def:extensor, def:join}
  The \emph{Plücker coordinate vector} of a $j$-extensor
  $v_1 \wedge \cdots \wedge v_j$ in $\R^{d+1}$ collects its $j \times j$
  minors over the increasing index sets
  $1 \le i_1 < \cdots < i_j \le d+1$, with the
  Katoh--Tanigawa~\cite[\S2.1]{katohTanigawa2011} sign convention
  $(-1)^{1 + \sum_\ell i_\ell}$. This is the coordinatized bridge from
  the symbolic exterior-algebra layer to the concrete minor vectors on
  which Lemma~2.1 and the downstream rank computations operate.
\end{definition}

\begin{definition}[Extensor of an affine subspace]
  \label{def:affine-subspace-extensor}
  \uses{def:homogeneous-coords, def:extensor, def:join}
  For an affine subspace spanned by affinely independent points
  $p_1, \dots, p_k \in \R^{d}$, its \emph{extensor} is
  $C(\cdot) = \bar p_1 \vee \cdots \vee \bar p_k \in \bigwedge^{k} \R^{d+1}$,
  well-defined up to a nonzero scalar. By \cref{lem:affine-indep-iff} it
  is nonzero exactly when the points are affinely independent, so $C$
  faithfully encodes the affine subspace.
\end{definition}

\subsection{Independence of the $(d-1)$-extensors (Lemma 2.1)}
\label{sec:molecular-lemma21}

\begin{lemma}[Extensor independence; Katoh--Tanigawa Lemma 2.1]
  \label{lem:extensor-independence}
  \uses{def:affine-subspace-extensor, def:plucker-coords, lem:affine-indep-iff}
  Let $p_1, \dots, p_{d+1} \in \R^{d}$ be affinely independent. The
  $D = \binom{d+1}{2}$ many $(d-1)$-extensors obtained by omitting two of
  the $d+1$ points,
  \[
    \bar p_1 \vee \cdots \vee \widehat{\bar p_a} \vee \cdots
      \vee \widehat{\bar p_b} \vee \cdots \vee \bar p_{d+1}
    \qquad (1 \le a < b \le d+1),
  \]
  are linearly independent in $\bigwedge^{d-1} \R^{d+1}$.
\end{lemma}
\begin{proof}
  \uses{def:join, lem:affine-indep-iff}
  Suppose a linear combination of the $D$ extensors vanishes. Fix a pair
  $\{a, b\}$ and join the relation with the $2$-extensor
  $\bar p_a \vee \bar p_b$. Every term whose omitted pair is not
  $\{a, b\}$ then repeats a vector $\bar p_a$ or $\bar p_b$, so by the
  alternating property of the join (\cref{def:join}) it vanishes; the
  single surviving term is a nonzero multiple of the top extensor
  $\bar p_1 \vee \cdots \vee \bar p_{d+1}$, which is nonzero by affine
  independence (\cref{lem:affine-indep-iff}). Hence the coefficient on
  the $\{a, b\}$-extensor is zero. As $\{a, b\}$ was arbitrary, all
  coefficients vanish.
\end{proof}

This independence is the linear-algebraic foundation that Case~III of
the conjecture's algebraic induction (Phases~22--23) bottoms out on: a
nonzero screw $r \in \R^{D}$ orthogonal to the extensors of $d+1$
distinct generic panels would, by \cref{lem:extensor-independence}, be
orthogonal to a spanning set of $\R^{D}$ and hence zero. The genuine
panel-hinge rigidity matrix that consumes this fact is built in
Phase~18 (the next chapter, not yet opened).
