\section{The combinatorial induction: graph operations and Theorem 4.9}
\label{sec:molecular-induction}

This chapter is the \textbf{Phase~20} dep-graph: the
\emph{combinatorial} half of Katoh--Tanigawa's proof of the molecular
conjecture --- \cite[\S3 (Lemmas 3.4, 3.5), \S4]{katohTanigawa2011},
the fourth stratum. Where \cref{sec:molecular-deficiency} built the
matroid $M(\tilde G)$, the $D$-deficiency, and the $k$-dof hierarchy,
this chapter develops the \emph{graph operations} that reduce a
minimal $k$-dof-graph to the two-vertex double edge, and assembles them
into Katoh--Tanigawa's Theorem~4.9. The algebraic induction of
Phases~21--23 then realizes each reduction step at the rigidity-matrix
rank. Throughout $D = \binom{d+1}{2} = \delta$ is the body-bar
dimension and $\tilde G = (D-1)\cdot G$ is the multiplied graph of
\cref{sec:body-hinge}.

\medskip
\noindent\emph{Status.} Forward mode: this chapter is the live to-do
list. Every node below is red until its Lean lands in
\texttt{Molecular/Induction.lean}; the leaf-most red node (statement
dependencies all green or mathlib) is the next target. The chapter
opens with two structural lemmas inherited from
\cref{sec:molecular-deficiency} --- the full form of KT Lemma 3.4 and
KT Lemma 3.5 --- whose lower bounds the Phase~19 def\,$=$\,corank bridge
(\cref{thm:def-eq-corank}) now unblocks, then builds the graph
operations, the degree-of-freedom tracking, and Theorem~4.9. The
forest-surgery lemmas KT~4.1/4.2 turn out to be \emph{deferred} (see
\cref{rem:kt-lemma-41}): KT's Lemma~4.1 is over-quantified and its proof
glosses a balanced-packing assumption, so the dof tracking is obtained
instead by a direct deficiency-count argument
(\cref{sec:molecular-induction-dof}).

\subsection{Inherited structural lemmas (KT 3.4 full form, KT 3.5)}
\label{sec:molecular-induction-inherited}

\begin{definition}[Vertex-induced subgraph from a fiber set]
  \label{def:induced-span}
  For a fiber set $X$ of the multiplied graph $\tilde G = (D-1)\cdot G$,
  the \emph{vertex span} $V(X)$ is the set of vertices of $G$ incident to
  some fiber of $X$ (\texttt{Graph.fiberSpan}), and the
  \emph{vertex-induced subgraph} $G[V(X)]$ is the induced subgraph of $G$
  on $V(X)$ (\texttt{Graph.inducedSpan}, mathlib's \texttt{Graph.induce}
  on $V(X)$). This is the induced-from-an-edge-set construction
  \cref{lem:circuit-induces-rigid} and the graph operations consume.
\end{definition}

\begin{lemma}[A circuit induces a rigid subgraph; KT Lemma 3.4, full form]
  \label{lem:circuit-induces-rigid}
  \lean{Graph.circuit_induces_isTight}
  \leanok
  \uses{lem:circuit-rigid, def:induced-span, def:rigid-subgraph}
  Let $X$ be a circuit of $M(\tilde G)$ and $e \in X$. Then $X - e$ is
  $(D,D)$-tight on its vertex span: $|X - e| = D(|V(X)| - 1)$, so $X - e$
  packs $D$ edge-disjoint spanning trees on $V(X)$ and the vertex-induced
  subgraph $G[V(X)]$ is rigid.
\end{lemma}
\begin{proof}
  \leanok
  \uses{def:matroid-MG}
  By \cref{lem:circuit-rigid}, $X - e$ is independent, hence $(D,D)$-sparse,
  giving the upper bound $|X - e| \le D(|V(X - e)| - 1) \le D(|V(X)| - 1)$.
  The matching lower bound $|X| > D(|V(X)| - 1)$ is forced by $X$ being a
  circuit: a circuit is a \emph{minimal} dependent set, so every proper
  subset of $X$ is independent (\,$(D,D)$-sparse\,), whence the sparsity
  failure of the dependent $X$ can only occur at $X$ itself
  (\cref{def:matroid-MG}, the boundary-regime cleanliness of $M(\tilde G)$).
  Together they give $|X - e| = D(|V(X)| - 1)$, i.e.\
  $\operatorname{def}(\widetilde{G[V(X)]}) = 0$.
\end{proof}

\begin{lemma}[A rigid subgraph attains full rank; KT Lemma 3.5, rank core]
  \label{lem:rigid-full-rank}
  \lean{Graph.rank_matroidMG_of_isKDof_zero}
  \leanok
  \uses{def:rigid-subgraph, thm:def-eq-corank, def:matroid-MG}
  For a rigid subgraph $H$ (so $\operatorname{def}(\tilde H) = 0$) with
  $V(H) \ne \emptyset$ and $D \ge 1$, the matroid $M(\tilde H)$ has full
  rank $\operatorname{rank} M(\tilde H) = D(|V(H)| - 1)$. This is the rank
  quantity that contracting $H$ removes from both $\operatorname{rank}
  M(\tilde G)$ and the ambient $D(|V| - 1)$.
\end{lemma}
\begin{proof}
  \leanok
  Immediate from the def\,$=$\,corank bridge (\cref{thm:def-eq-corank},
  $\operatorname{rank} M(\tilde H) + \operatorname{def}(\tilde H) = D(|V(H)|
  - 1)$) with $\operatorname{def}(\tilde H) = 0$ from rigidity.
\end{proof}

\begin{lemma}[Contraction rank bridge; KT Lemma 3.5, contraction arithmetic]
  \label{lem:contract-rank-bridge}
  \lean{Graph.contract_matroidMG_rank}
  \leanok
  \uses{def:rigid-contraction, def:matroid-MG, lem:matroid-restrict-subgraph}
  For a subgraph $H \le G$, contracting the edge-fibers $E(\tilde H)$ in
  $M(\tilde G)$ removes exactly $\operatorname{rank} M(\tilde H)$:
  \[
    \operatorname{rank}\bigl(M(\tilde G) / E(\tilde H)\bigr)
      + \operatorname{rank} M(\tilde H)
      = \operatorname{rank} M(\tilde G).
  \]
\end{lemma}
\begin{proof}
  \leanok
  The standard matroid identity $\operatorname{rank}(M / C) =
  \operatorname{rank} M - \operatorname{rank}_M(C)$ in additive form, on
  $M = M(\tilde G)$ and $C = E(\tilde H)$. The restriction $M(\tilde G)
  \restriction E(\tilde H)$ is $M(\tilde H)$
  (\cref{lem:matroid-restrict-subgraph}), whose rank is
  $\operatorname{rank}_{M(\tilde G)}(E(\tilde H))$.
\end{proof}

\begin{lemma}[Contracting a rigid subgraph conserves the deficiency; KT Lemma 3.5]
  \label{lem:contract-deficiency-conservation}
  \lean{Graph.contract_matroidMG_deficiency_eq}
  \leanok
  \uses{lem:rigid-full-rank, lem:contract-rank-bridge, thm:def-eq-corank}
  For a rigid subgraph $H \le G$ (so $\operatorname{def}(\tilde H) = 0$) with $V(H) \ne
  \emptyset$ and $D \ge 1$, the corank of the matroid contraction $M(\tilde G) / E(\tilde
  H)$ at the reduced ambient $D(|V(G)| - |V(H)|)$ --- the trivial-motion count of the
  contracted graph, which has $|V(G)| - |V(H)| + 1$ vertices --- equals
  $\operatorname{def}(\tilde G)$:
  \[
    D\bigl(|V(G)| - |V(H)|\bigr) - \operatorname{rank}\bigl(M(\tilde G) / E(\tilde H)\bigr)
      = \operatorname{def}(\tilde G).
  \]
\end{lemma}
\begin{proof}
  \leanok
  Pure matroid bookkeeping. \Cref{lem:contract-rank-bridge} with the rank core
  \cref{lem:rigid-full-rank} ($\operatorname{rank} M(\tilde H) = D(|V(H)| - 1)$) gives
  $\operatorname{rank}(M(\tilde G) / E(\tilde H)) = \operatorname{rank} M(\tilde G) -
  D(|V(H)| - 1)$. Substituting into the def\,$=$\,corank bridge
  (\cref{thm:def-eq-corank}, $\operatorname{rank} M(\tilde G) + \operatorname{def}(\tilde
  G) = D(|V(G)| - 1)$) and cancelling the $D(|V(H)| - 1)$ between the rank drop and the
  ambient drop leaves $\operatorname{def}(\tilde G)$. No graph$\leftrightarrow$matroid
  correspondence is needed --- the statement is against the matroid contraction directly.
\end{proof}

\begin{lemma}[Minimality transports along a contraction; KT Lemma 3.5, second half]
  \label{lem:contract-minimality-transport}
  \lean{Graph.contract_minimality_transport}
  \leanok
  \uses{lem:subgraph-minimality, def:k-dof, def:rigid-contraction}
  Let $G$ be a minimal $k$-dof-graph and $H$ a subgraph. Then every base
  $B'$ of the matroid contraction $M(\tilde G) / E(\tilde H)$ meets every
  \emph{surviving} edge-fiber $\tilde e$ of an edge $e \in E(G) \setminus
  E(H)$: $B' \cap \tilde e \ne \emptyset$. This is the contraction analogue
  of the restriction transport \cref{lem:subgraph-minimality}.
\end{lemma}
\begin{proof}
  \leanok
  \uses{def:matroid-MG}
  A base $B'$ of $M(\tilde G) / E(\tilde H)$ lies in the contraction's
  ground $E(\tilde G) \setminus E(\tilde H)$, so it is disjoint from
  $E(\tilde H)$. Pick an $M(\tilde G)$-basis $J$ of $E(\tilde H)$; then
  $B' \cup J$ is a base of $M(\tilde G)$ (the deleted part $E(\tilde H)
  \setminus J$ consists of loops of the contraction $M(\tilde G) / J$, so
  $B'$ is a base there and \cref{def:matroid-MG}'s independence lifts). By
  $G$'s minimality, $(B' \cup J) \cap \tilde e \ne \emptyset$; the surviving
  fiber $\tilde e$ ($e \notin E(H)$) is disjoint from $E(\tilde H) \supseteq
  J$, so the witness lands in $B'$. No graph$\leftrightarrow$matroid
  correspondence is needed.
\end{proof}

\begin{lemma}[Rigid-subgraph contraction preserves minimality; KT Lemma 3.5]
  \label{lem:contraction-minimality}
  \lean{Graph.contraction_isMinimalKDof}
  \leanok
  \uses{def:rigid-subgraph, lem:subgraph-minimality, def:k-dof, lem:rigid-full-rank, lem:contract-rank-bridge, lem:contract-deficiency-conservation, lem:contract-minimality-transport}
  Let $G$ be a minimal $k$-dof-graph and let $H$ be a proper rigid
  subgraph of $G$. Then the contraction $G / E(H)$ (collapse $V(H)$ to
  a single vertex, keeping the edges of $G$ incident to $V(H)$ from
  outside) is again a minimal $k$-dof-graph, with the same deficiency.
  Stated on the matroid side --- the matroid contraction $M(\tilde G) /
  E(\tilde H)$ is a minimal $k$-dof matroid at the reduced ambient
  $D(|V(G)| - |V(H)|)$ --- so no graph$\leftrightarrow$matroid
  correspondence is needed. This is the engine of Case~I of the
  algebraic induction.
\end{lemma}
\begin{proof}
  \leanok
  \uses{lem:contract-deficiency-conservation, lem:contract-minimality-transport}
  The deficiency is unchanged by
  \cref{lem:contract-deficiency-conservation}: contracting a rigid
  subgraph removes $D(|V(H)| - 1)$ from both the rank of $M(\tilde G)$
  and the ambient $D(|V| - 1)$, so the corank --- hence the deficiency
  --- is conserved. Minimality (every base of the contracted matroid
  meets every edge-fiber) transports from $G$ along the contraction,
  using the subgraph-minimality machinery of
  \cref{lem:subgraph-minimality}.
\end{proof}

\subsection{Graph operations}
\label{sec:molecular-induction-operations}

\begin{definition}[Vertex removal, splitting-off, and edge-splitting]
  \label{def:graph-operations}
  \lean{Graph.removeVertex, Graph.splitOff, Graph.edgeSplit}
  \leanok
  \uses{def:k-dof}
  For a multigraph $G$ and a vertex $v$ of degree~2 with the two
  incident edges joining $v$ to $a$ and $b$:
  \begin{itemize}
    \item \emph{Removal} $G_v := G - v$ deletes $v$ and its incident
      edges.
    \item \emph{Splitting-off} $G_v^{ab}$ deletes $v$ and adds the new
      edge $ab$ (the two edges through $v$ are short-circuited into a
      single edge between its neighbours).
    \item \emph{Edge-splitting} $H_{ab}^v$ is the inverse: subdivide an
      edge $ab$ of $H$ by a fresh degree-2 vertex $v$, replacing $ab$
      with the path $a$--$v$--$b$. It satisfies $(H_{ab}^v)_v^{ab} = H$.
  \end{itemize}
\end{definition}

\begin{definition}[Rigid-subgraph contraction]
  \label{def:rigid-contraction}
  \lean{Graph.rigidContract}
  \leanok
  \uses{def:rigid-subgraph}
  For a rigid subgraph $H \le G$, the \emph{contraction} $G / E(H)$
  collapses the vertex set $V(H)$ to a single vertex and discards the
  edges of $H$, retaining every other edge of $G$ (with endpoints in
  $V(H)$ redirected to the collapsed vertex). On the matroid side this
  is the matroid contraction $M(\tilde G) / E(\tilde H)$ restricted to
  the surviving fibers.
\end{definition}

\subsection{Forest surgery (deferred --- see the remark on KT Lemma 4.1)}
\label{sec:molecular-induction-forest}

\begin{remark}[On Katoh--Tanigawa Lemma 4.1]
  \label{rem:kt-lemma-41}
  Formalizing the forest surgery of \cite[Lemma 4.1]{katohTanigawa2011}
  (p.~660; the 2009 arXiv predecessor, Lemma 5.1, p.~11, is essentially
  verbatim) surfaced a three-layer subtlety. \emph{The molecular
  conjecture and Katoh--Tanigawa's proof are not in question}; the value
  is in stating exactly which layer is which.
  \begin{enumerate}
    \item \emph{Statement-as-quantified --- false.} Both versions read
      ``for \emph{any} independent set $I$ of $M(\tilde G)$ \dots there
      exists $I'$ \dots with $|I'| = |I| - D$''. For any independent $I$
      with $|I| < D$ --- e.g.\ $I = \emptyset$ --- this demands $|I'| =
      |I| - D < 0$, which is impossible. The intended quantifier is
      clearly over \emph{bases} of $M(\tilde G)$; the universal-over-$I$
      reading must be restricted. We record the literal disproof as
      \cref{ex:kt-41-overquantified}.
    \item \emph{Proof of the intended (base) case --- an unjustified
      step.} ``Convert each $F_i$ to $F'_i$ with $|F'_i| = |F_i| - 1$
      for each $i$'' (2011), and the equation $2h' + (D - h') = h$
      (2009), silently assume the chosen $D$-forest packing is
      \emph{balanced at $v$}: every forest uses at least one edge at the
      degree-2 vertex $v$. Neither version justifies this, and for a
      general packing it can fail (pigeonhole when $h < D$); when it
      fails, both the cardinality count and the slack clause break.
      Whether the balanced packing is achievable for a base we leave
      \emph{open} --- Katoh--Tanigawa omit a justification and we did not
      recover one.
    \item \emph{Intended content --- true.} What the induction
      (\cref{lem:dof-tracking}) actually consumes is the deficiency
      inequality $\operatorname{def}(\widetilde{G_v^{ab}}) \le
      \operatorname{def}(\tilde G)$. This is correct, and we establish it
      directly (\cref{lem:splitoff-deficiency}) by a partition-count
      comparison through the def\,$=$\,corank bridge
      (\cref{thm:def-eq-corank}), bypassing the forest surgery entirely.
  \end{enumerate}
  Accordingly the forest-surgery lemmas
  (\cref{lem:forest-surgery-split,lem:forest-surgery-unsplit}) are
  \textbf{deferred / off the Theorem-4.9 critical path}: their substrate
  is formalized but they complete only if the balanced-packing
  assumption of layer~(2) is settled. Theorem~4.9 is reached instead
  through the deficiency-counting route of
  \cref{sec:molecular-induction-dof}.
\end{remark}

\begin{example}[KT Lemma 4.1's universal quantification is unsatisfiable]
  \label{ex:kt-41-overquantified}
  % Formalized as an unnamed Lean `example`; no `\lean{}` pointer to pin.
  \leanok
  \uses{def:matroid-MG, rem:kt-lemma-41}
  Taking $I = \emptyset$ (independent in any matroid), the promised $I'$
  with $|I'| = |I| - D = -D$ would need $|I'| + D = 0$ in $\mathbb{N}$,
  impossible for $D = \binom{d+1}{2} \ge 1$. This is layer~(1) of
  \cref{rem:kt-lemma-41}: the universal-over-$I$ quantification is
  formally false, witnessing that the intended quantifier ranges over
  bases. Formalized as a one-line \texttt{example} in
  \texttt{Molecular/Induction.lean}.
\end{example}

\begin{lemma}[Forest-packing decomposition of an $M(\tilde G)$-independent set]
  \label{lem:forest-packing-decomp}
  \lean{Graph.matroidMG_indep_iff_exists_forest_packing}
  \leanok
  \uses{def:matroid-MG}
  A fiber set $I \subseteq E(\tilde G)$ is independent in $M(\tilde G)$ if and only
  if it is covered by $D = \binom{d+1}{2}$ cycle-matroid-independent fiber sets
  $F_0, \dots, F_{D-1}$ --- the \emph{$D$ edge-disjoint forests on $V(\tilde G)$} of
  Katoh--Tanigawa's surgery: $\bigcup_{i} F_i = I$ with each $F_i$ a forest of
  $\tilde G$. This is the matroidal reading of ``$I$ partitions into $D$ forests'',
  fixing the framing of the forest surgery below: a \emph{forest} is a
  cycle-matroid-independent fiber set, and the $D$-forest partition is the union
  decomposition of $M(\tilde G)$ --- no hand-rolled acyclicity predicate.
\end{lemma}
\begin{proof}
  \leanok
  $M(\tilde G)$ is the $D$-fold cycle-matroid union restricted to $E(\tilde G)$
  (\cref{def:matroid-MG}), so independence unfolds through matroid restriction and
  the matroid-union characterization (Nash--Williams~\cite{nashWilliams1966},
  Edmonds~\cite{edmonds1965}): independence in $\bigcup_{i < D}
  \operatorname{cycleMatroid}(\tilde G)$ is exactly a cover by $D$
  cycle-matroid-independent sets.
\end{proof}

\begin{lemma}[Forest surgery at a degree-2 vertex, splitting-off direction; KT Lemma 4.1]
  \label{lem:forest-surgery-split}
  \uses{def:graph-operations, def:matroid-MG, thm:def-eq-corank, lem:forest-packing-decomp, rem:kt-lemma-41}
  \emph{Deferred / off the Theorem-4.9 critical path}
  (\cref{rem:kt-lemma-41}): its substrate is formalized in
  \texttt{Molecular/Induction.lean}, but the lemma completes only if the
  balanced-packing assumption of \cref{rem:kt-lemma-41}~(2) is settled;
  the induction instead uses \cref{lem:splitoff-deficiency}.
  Let $v$ be a degree-2 vertex of $G$ with neighbours $a, b$. A
  packing of $\tilde G$ into $D$ edge-disjoint forests, restricted to
  the fibers not incident to $v$ and rerouted across $v$, induces a
  packing of $\widetilde{G_v^{ab}}$, with the deficiency changing in a
  controlled way. Concretely the surgery rewires the $D-1$ fibers of
  one of $v$'s edges through $v$ onto the new edge $ab$.
\end{lemma}
\begin{proof}
  Explicit forest surgery: each of the $D$ spanning forests of $\tilde
  G$ uses some subset of the $2(D-1)$ fibers at $v$; rerouting a
  $v$-traversing tree-path across the short-circuit $ab$ keeps each
  forest acyclic and edge-disjoint. The bookkeeping is the hard new
  combinatorial core of Phase~20; no existing analogue in the project
  or mathlib.
\end{proof}

\begin{lemma}[Forest surgery at a degree-2 vertex, edge-splitting direction; KT Lemma 4.2]
  \label{lem:forest-surgery-unsplit}
  \uses{def:graph-operations, lem:forest-surgery-split, rem:kt-lemma-41}
  \emph{Deferred} with \cref{lem:forest-surgery-split}
  (\cref{rem:kt-lemma-41}); off the Theorem-4.9 critical path.
  The inverse direction: a packing of $\widetilde{H}$ extends across an
  edge-split $H_{ab}^v$ to a packing of $\widetilde{H_{ab}^v}$,
  matching deficiencies under
  $(H_{ab}^v)_v^{ab} = H$. Together with
  \cref{lem:forest-surgery-split} this makes splitting-off and
  edge-splitting inverse operations on the deficiency.
\end{lemma}
\begin{proof}
  The fresh degree-2 vertex $v$ contributes $D$ to $D(|V|-1)$ and the
  subdivided edge contributes $D-1$ crossing fibers; the surgery of
  \cref{lem:forest-surgery-split} run in reverse distributes the new
  fibers across the $D$ forests. Explicit forest construction, the
  companion of the splitting-off surgery.
\end{proof}

\subsection{Degree-of-freedom tracking (via deficiency counting)}
\label{sec:molecular-induction-dof}

The dof-conservation laws Katoh--Tanigawa derive from the forest surgery
(KT~4.3--4.5) are obtained here directly from the deficiency
max-formula, on the green infrastructure of
\cref{sec:molecular-deficiency} --- the partition deficiency
$\operatorname{def}(\tilde G) = \max_P [\, D(|P| - 1) - (D - 1)\,d_G(P)\,]$
and the def\,$=$\,corank bridge (\cref{thm:def-eq-corank}). This is the
route that bypasses the deferred forest surgery
(\cref{rem:kt-lemma-41}).

\begin{lemma}[Splitting-off does not increase the deficiency; KT Lemma 4.3]
  \label{lem:splitoff-deficiency}
  \uses{def:graph-operations, def:matroid-MG, thm:def-eq-corank}
  Let $v$ be a degree-2 vertex of $G$ with neighbours $a, b$. Then
  $\operatorname{def}(\widetilde{G_v^{ab}}) \le \operatorname{def}(\tilde
  G)$, with the refinement that the deficiency drops by at most the
  $k$-vs-$(k-1)$ amount the induction tracks (KT~4.3(ii)). This is the
  substantive splitting-off fact the induction consumes, replacing the
  forest surgery of \cref{lem:forest-surgery-split}.
\end{lemma}
\begin{proof}
  Partition-count comparison. Take an optimal partition $P'$ of $V - v$
  realizing $\operatorname{def}(\widetilde{G_v^{ab}})$ and extend it to a
  partition $P$ of $V$ by dropping $v$ into $a$'s block; then $|P| =
  |P'|$. The edge $va$ no longer crosses $P$ while $vb$ crosses iff $ab$
  crosses $P'$, so the crossing counts agree, $d_G(P) =
  d_{G_v^{ab}}(P')$, whence the per-partition deficiency is unchanged and
  $\operatorname{def}(\tilde G) \ge \operatorname{def}(\widetilde{G_v^{ab}})$.
  No forests.
\end{proof}

\begin{lemma}[Vertex removal raises the deficiency; KT Lemma 4.4]
  \label{lem:removal-deficiency}
  \uses{def:graph-operations, def:matroid-MG, thm:def-eq-corank}
  Let $v$ be a degree-2 vertex of the minimal $k$-dof-graph $G$. Then the
  removal $G_v$ has $\operatorname{def}(\widetilde{G_v}) \ge k$: deleting
  $v$ and its two incident edges removes $D$ from the ambient and at most
  $2(D-1)$ from the rank, so the corank cannot decrease below $k$. This
  is the other half of the local dof bookkeeping at $v$.
\end{lemma}
\begin{proof}
  The def\,$=$\,corank bridge (\cref{thm:def-eq-corank}) at $G_v$: the
  ambient $D(|V| - 1)$ drops by $D$ and the rank of $M(\tilde G)$ drops
  by at most the $2(D-1)$ fibers at $v$, so $\operatorname{def}$ does not
  fall below the $k = \operatorname{def}(\tilde G)$ of the minimal graph.
\end{proof}

\begin{lemma}[Deficiency under vertex removal and splitting-off; KT Lemmas 4.3--4.5]
  \label{lem:dof-tracking}
  \uses{lem:splitoff-deficiency, lem:removal-deficiency, def:graph-operations}
  For a degree-2 vertex $v$ of $G$, the deficiencies of $G$, the
  removal $G_v$, and the splitting-off $G_v^{ab}$ are related by the
  KT~4.3--4.5 inequalities, pinning $\operatorname{def}(\tilde G)$ in
  terms of $\operatorname{def}(\widetilde{G_v^{ab}})$ and the local
  structure at $v$ (including the fundamental-circuit count of the new
  edge $ab$). These are the dof-conservation laws the induction tracks.
\end{lemma}
\begin{proof}
  \uses{lem:splitoff-deficiency, lem:removal-deficiency}
  Assemble the splitting-off bound
  (\cref{lem:splitoff-deficiency}) and the removal bound
  (\cref{lem:removal-deficiency}); each is a consequence of the
  def\,$=$\,corank bridge (\cref{thm:def-eq-corank}) under the
  corresponding rank change, and together they pin
  $\operatorname{def}(\tilde G)$ in terms of the local structure at $v$.
  This replaces Katoh--Tanigawa's forest-surgery derivation
  (\cref{rem:kt-lemma-41}).
\end{proof}

\begin{lemma}[Existence of a reducible degree-2 vertex; KT Lemma 4.6]
  \label{lem:reducible-vertex}
  \uses{lem:dof-tracking, lem:two-edge-conn, def:k-dof}
  A minimal $k$-dof-graph $G$ with $|V| \ge 3$ that has no proper rigid
  subgraph contains a degree-2 vertex whose splitting-off (or removal)
  preserves minimality and the degree of freedom. The proof is a
  maximal-chain / degree-sequence counting argument on the
  $2$-edge-connected graph (\cref{lem:two-edge-conn}).
\end{lemma}
\begin{proof}
  Degree-sequence counting: a minimal $k$-dof-graph has average degree
  bounded by the $(D,D)$-tightness, forcing a low-degree vertex; the
  $2$-edge-connectivity (\cref{lem:two-edge-conn}) rules out degree
  $\le 1$, and the no-proper-rigid-subgraph hypothesis makes a degree-2
  vertex reducible. This is the lemma that guarantees the induction can
  always take a step.
\end{proof}

\begin{lemma}[Reduction step preserves minimality; KT Lemmas 4.7--4.8]
  \label{lem:reduction-step}
  \uses{lem:reducible-vertex, lem:contraction-minimality, lem:dof-tracking}
  Each reduction --- splitting-off at the degree-2 vertex of
  \cref{lem:reducible-vertex}, or contraction of a proper rigid
  subgraph (\cref{lem:contraction-minimality}) --- carries a minimal
  $k$-dof-graph to a smaller minimal $k$-dof-graph with the same $k$.
  KT~4.8 is the capstone, run through two circuit-swap arguments on the
  fundamental circuits of $M(\tilde G)$.
\end{lemma}
\begin{proof}
  Minimality transport: the base / fiber-meeting condition of
  \cref{def:k-dof} is preserved by each operation, via fundamental
  circuit swaps in $M(\tilde G)$ (each base of the reduced matroid
  lifts to a base of $M(\tilde G)$ meeting every fiber). The deficiency
  is unchanged by \cref{lem:dof-tracking,lem:contraction-minimality}.
\end{proof}

\subsection{Theorem 4.9}
\label{sec:molecular-induction-thm49}

\begin{theorem}[Reduction of minimal $k$-dof-graphs; KT Theorem 4.9]
  \label{thm:minimal-kdof-reduction}
  \uses{lem:reduction-step, lem:reducible-vertex, lem:contraction-minimality, def:graph-operations}
  Every minimal $k$-dof-graph $G$ with $|V| \ge 2$ reduces to the
  two-vertex double edge by a sequence of splitting-off and
  proper-rigid-subgraph contraction operations. Equivalently, $G$ is
  generated from the two-vertex double edge by the inverse operations
  (edge-splitting and rigid-subgraph expansion).
\end{theorem}
\begin{proof}
  Induction on $|V|$. The base case $|V| = 2$ is the two-vertex double
  edge. For $|V| \ge 3$: if $G$ has a proper rigid subgraph, contract
  it (\cref{lem:contraction-minimality}); otherwise
  \cref{lem:reducible-vertex} supplies a reducible degree-2 vertex,
  split it off. Either way \cref{lem:reduction-step} produces a smaller
  minimal $k$-dof-graph with the same $k$, and the induction hypothesis
  applies. This is Katoh--Tanigawa~\cite[Theorem 4.9]{katohTanigawa2011};
  it is the combinatorial skeleton the algebraic induction
  (Phases~21--23) realizes at the rigidity-matrix rank.
\end{proof}
