\section{The \texorpdfstring{$(k, \ell)$}{(k, l)}-pebble game}
\label{sec:pebble-game}

The $(k, \ell)$-count matroid of \cref{sec:count-matroid} is defined
combinatorially: a subset of $E(K_V)$ is independent precisely when
the induced subgraph is $(k, \ell)$-sparse. \Cref{sec:count-matroid}
established that this collection of subsets satisfies the matroid
axioms via an augmentation / $I$-component argument, but the proof
is not constructive: given an arbitrary edge set, it does not
exhibit a procedure to decide whether the set is independent.

The \emph{$(k, \ell)$-pebble game} of
Lee--Streinu~\cite{leeStreinu2008} (generalising the original
$(2, 3)$ algorithm of Jacobs--Hendrickson~\cite{jacobsHendrickson1997})
is the missing algorithmic side. It processes a list of candidate
edges one at a time, maintaining a partial orientation of the
accepted edges together with a distribution of \emph{pebbles} on
the vertices, and decides at each step whether to insert the
candidate or reject it. We formalise the basic version of the
algorithm (Lee--Streinu \S 3) and prove its correctness theorem
(Lee--Streinu \S 4, Theorem 8) in the matroidal regime $\ell < 2k$
that matches \cref{sec:count-matroid}, in a certificate form: on
success, the algorithm returns the partial orientation as a
witness of sparsity; on failure, it returns a vertex subset $V'$
together with a proof that $|\edgesIn{V'}| > k|V'| - \ell$,
witnessing non-sparsity.

\emph{Out of scope.} The \emph{component pebble game} of
Lee--Streinu \S 5 maintains the $I$-components of the partial
orientation explicitly to reject dependent edges in $O(1)$ rather
than $O(n)$ time, achieving an overall $O(n^2)$ bound; the
underlying data structure (\emph{union pair-find}) was developed
by Lee--Streinu--Theran~\cite{leeStreinuTheran2005}. The Henneberg-
sequence application of Lee--Streinu \S 6 generalises the
Phase~3 Henneberg construction (\cref{sec:henneberg}) from Laman
graphs to all $(k, \ell)$-tight graphs. Neither is in scope for
this chapter; both are candidates for follow-up phases.

\emph{Multigraphs.} Lee--Streinu state results for multi-graphs
with loops, covering $\ell \in [0, 2k)$. Following the project
convention (\cref{sec:sparsity}), we work in $\mathrm{SimpleGraph}\,V$:
the loop-free assumption removes the $|V'| = 1$ single-vertex
block case of L-S Lemma 10 in the lower range $\ell < k$, but the
matroidal-regime theory carries over otherwise. The algorithm
state (\cref{def:partial-orientation} below) is decoupled from the
shape of the input edge container, so retargeting to a future
\texttt{MultiGraph} type is a relaxation of the no-parallel-pair
invariant on the partial orientation.

\subsection{State and moves}
\label{sec:pebble-game-state}

The pebble game maintains a partial orientation of a subset of
$E(K_V)$ together with a non-negative integer of \emph{free
pebbles} at each vertex. Both pieces of data are determined by
the orientation: pebbles are bookkeeping for the per-vertex
out-degree budget.

\begin{definition}[Partial orientation]
  \label{def:partial-orientation}
  \lean{CombinatorialRigidity.PebbleGame.PartialOrientation}
  \leanok
  Let $V$ be a finite type with decidable equality. A
  \emph{partial orientation} on $V$ is a finite set
  $D \subseteq V \times V$ of ordered pairs such that:
  \begin{enumerate}
    \item (no loops) $(v, v) \notin D$ for all $v \in V$;
    \item (no antiparallel pairs) $(u, v) \in D$ implies
      $(v, u) \notin D$.
  \end{enumerate}
  An edge $\{u, v\} \in \mathrm{Sym}_2 V$ is \emph{oriented in $D$}
  if either $(u, v) \in D$ or $(v, u) \in D$. The
  \emph{underlying edge set} of $D$, written $\underline{D} \subseteq
  E(K_V)$, is the set of oriented edges.
\end{definition}

\begin{definition}[Out-degree, pebble count, span]
  \label{def:pebble-counts}
  \lean{CombinatorialRigidity.PebbleGame.PartialOrientation.out,
    CombinatorialRigidity.PebbleGame.PartialOrientation.peb,
    CombinatorialRigidity.PebbleGame.PartialOrientation.span,
    CombinatorialRigidity.PebbleGame.PartialOrientation.outOn,
    CombinatorialRigidity.PebbleGame.PartialOrientation.pebOn}
  \leanok
  \uses{def:partial-orientation}
  Fix a partial orientation $D$ and an integer $k \ge 0$. For a
  vertex $v \in V$, write
  \[
    \mathrm{out}_D(v) := |\{u : (v, u) \in D\}|,
    \qquad
    \mathrm{peb}_{D,k}(v) := k - \mathrm{out}_D(v).
  \]
  Extend additively to subsets $V' \subseteq V$:
  $\mathrm{peb}_{D,k}(V') := \sum_{v \in V'} \mathrm{peb}_{D, k}(v)$,
  and write
  \[
    \mathrm{span}_D(V') := |\{(u, w) \in D : u, w \in V'\}|,
    \qquad
    \mathrm{out}_D(V') := |\{(u, w) \in D : u \in V',\ w \notin V'\}|.
  \]
  The pebble count is non-negative provided $\mathrm{out}_D(v) \le k$
  for all $v$, which is preserved by the algorithm
  (\cref{def:tryAddEdge}) as a structural invariant.
\end{definition}

\begin{definition}[Path-reversal move]
  \label{def:path-reversal}
  \lean{CombinatorialRigidity.PebbleGame.PartialOrientation.reverse}
  \leanok
  \uses{def:partial-orientation}
  Let $D$ be a partial orientation and $u_0 \to u_1 \to \cdots \to
  u_m$ a directed path in $D$ (a sequence of vertices with
  $(u_i, u_{i+1}) \in D$ for $0 \le i < m$, all $u_i$ distinct).
  The \emph{reversal} of this path replaces $D$ by
  $D \setminus \{(u_i, u_{i+1}) : 0 \le i < m\} \cup
  \{(u_{i+1}, u_i) : 0 \le i < m\}$. The result is again a partial
  orientation: no loops are introduced (the path is simple), and the
  no-antiparallel-pairs invariant is preserved by induction on $m$.
  Path reversal preserves $\underline{D}$, $\mathrm{out}_D(v)$ for
  all $v \neq u_0, u_m$, and decreases $\mathrm{out}_D(u_0)$ by $1$
  while increasing $\mathrm{out}_D(u_m)$ by $1$; equivalently, the
  pebble count $\mathrm{peb}_{D, k}$ rises by $1$ at $u_0$ (under
  the precondition $\mathrm{out}_D(u_0) \le k$ from
  \cref{lem:pebble-game-invariants}~(1)) and drops by $1$ at $u_m$
  (under $\mathrm{out}_D(u_m) < k$, the precondition for the
  reversal move to be valid). The invariant-preservation half,
  span invariance ($\mathrm{span}_D(V')$ unchanged on every $V'$),
  the per-vertex \texttt{out}/\texttt{peb} effects, and the
  subset-level $\mathrm{peb}_{D, k}(V') + \mathrm{out}_D(V')$
  invariance under reversal are formalised. The last is routed
  through the algebraic identity $\mathrm{peb}_{D, k}(V') +
  \mathrm{span}_D(V') + \mathrm{out}_D(V') = k|V'|$
  (\cref{lem:pebble-game-invariants}~(2)) applied to both $D$ and
  $D'$: since $\mathrm{span}_D(V')$ is invariant, $\mathrm{peb}(V')
  + \mathrm{out}(V')$ must match. (The subset-level
  $\mathrm{out}_D(V')$ itself is *not* invariant: it shifts by
  $[u_m \in V'] - [u_0 \in V']$, cancelling the corresponding shift
  in $\mathrm{peb}_{D, k}(V')$.)
\end{definition}

\subsection{Invariants}
\label{sec:pebble-game-invariants}

The four invariants of Lee--Streinu Lemma~10 are the soundness
backbone of the algorithm: they are maintained at every state
reachable from the initial empty orientation through edge
insertions and path reversals, and Invariant~(4) directly gives
the sparsity of $\underline{D}$.

\begin{lemma}[Pebble-game invariants]
  \label{lem:pebble-game-invariants}
  \uses{def:partial-orientation, def:pebble-counts}
  Let $V$ be a finite type, $k, \ell$ integers with $\ell < 2k$,
  and $D$ a partial orientation on $V$ reachable from the empty
  orientation by a sequence of \emph{accepted insertions} and
  \emph{path reversals}, where an accepted insertion of edge
  $\{u, v\}$ is permitted only when $\mathrm{peb}_{D, k}(u) +
  \mathrm{peb}_{D, k}(v) \ge \ell + 1$ at the moment of insertion.
  Then for every $v \in V$ and every $V' \subseteq V$ with
  $|V'| \ge 1$ when $\ell \le k$ and $|V'| \ge 2$ when $\ell > k$:
  \begin{enumerate}
    \item $\mathrm{peb}_{D, k}(v) + \mathrm{out}_D(v) = k$.
    \item $\mathrm{peb}_{D, k}(V') + \mathrm{span}_D(V') +
      \mathrm{out}_D(V') = k|V'|$.
    \item $\mathrm{peb}_{D, k}(V') + \mathrm{out}_D(V') \ge \ell$.
    \item $\mathrm{span}_D(V') \le k|V'| - \ell$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  \uses{def:partial-orientation, def:pebble-counts, def:path-reversal}
  Invariants (1) and (2) are bookkeeping: (1) is the definition of
  $\mathrm{peb}_{D, k}$; (2) follows by summing (1) over $v \in V'$
  and decomposing $\sum_{v \in V'} \mathrm{out}_D(v) =
  \mathrm{span}_D(V') + \mathrm{out}_D(V')$ along whether the
  arrow's head lies in $V'$. Invariant (4) is the algebraic
  rearrangement of (2) given (3). Invariant (3) is the substantive
  claim, proved by induction on the move sequence: the empty
  orientation has $\mathrm{peb}_{\emptyset, k}(V') = k|V'| \ge \ell$
  (using $|V'| \ge 1$ and $\ell \le k$, or $|V'| \ge 2$ and
  $\ell < 2k$); accepted insertions either preserve the LHS (edge
  inside $V'$ or outside $V'$ entirely) or trade one pebble in $V'$
  for one outgoing edge from $V'$, leaving the sum invariant; path
  reversals move a pebble along the path, changing
  $\mathrm{peb}(V')$ and $\mathrm{out}(V')$ by $\pm 1$ in lockstep.
  At an accepted insertion of an edge $\{u, v\}$ with both endpoints
  in $V'$, the precondition gives $\mathrm{peb}(V') \ge \ell + 1$,
  so the post-insertion $\mathrm{peb}(V') \ge \ell$. Compare
  Lee--Streinu Lemma 10.
\end{proof}

\subsection{The basic algorithm}
\label{sec:pebble-game-algorithm}

The algorithm processes a list of candidate edges one at a time.
For each edge it attempts to redirect pebbles in $D$ until both
endpoints together hold at least $\ell + 1$ free pebbles, then
inserts and consumes one pebble. If the pebble target cannot be
reached, the edge is rejected and the algorithm halts with a
failure witness.

\begin{definition}[Try-reach-pebble: DFS with path reversal]
  \label{def:tryReachPebble}
  \uses{def:partial-orientation, def:pebble-counts, def:path-reversal,
    thm:reachable-finding-correct}
  Given a partial orientation $D$, a source vertex $v$, and a
  predicate $P$ on $V$, the procedure $\mathrm{tryReachPebble}\,
  D\,v\,P$ runs $\mathrm{reachableFinding}$
  (\cref{def:reachable-finding}) with relation $R$ taken to be the
  outgoing-arc relation of $D$. On success it receives a witness
  walk $p$ from $v$ to a vertex $w$ with $P(w)$
  (\cref{thm:reachable-finding-correct}) and reverses every arc of
  $p$ in $D$ via \cref{def:path-reversal}, returning the updated
  $(D', w)$. On failure it propagates $\mathrm{none}$. The
  reachability closure $\mathrm{Reach}_D(v)$ used in later lemmas
  coincides with the visited set of a fully-exhausted DFS run.
\end{definition}

\begin{definition}[Try-add-edge]
  \label{def:tryAddEdge}
  \uses{def:tryReachPebble, def:partial-orientation}
  Given a partial orientation $D$, parameters $k, \ell$, and a
  candidate edge $\{u, v\} \notin \underline{D}$ (with $u \neq v$),
  the procedure $\mathrm{tryAddEdge}\,k\,\ell\,D\,\{u, v\}$
  repeatedly calls $\mathrm{tryReachPebble}\,D\,u\,(\mathrm{peb}_{D,k}
  > 0\ \text{at the vertex})$ and $\mathrm{tryReachPebble}\,D\,v\,
  (\mathrm{peb}_{D,k} > 0)$, accumulating successful reversals into
  $D$, until either:
  \begin{itemize}
    \item (accept) $\mathrm{peb}_{D, k}(u) + \mathrm{peb}_{D, k}(v)
      \ge \ell + 1$ is reached; insert $(u, v)$ into $D$ and return
      $\mathrm{Sum.inl}\,D'$ with $D'$ the updated orientation; or
    \item (reject) both DFS attempts simultaneously fail (no free
      pebble reachable from $u$ or $v$); return $\mathrm{Sum.inr}\,V'$
      with $V' := \mathrm{Reach}_D(u) \cup \mathrm{Reach}_D(v)$.
  \end{itemize}
  Termination of the outer loop: $(\ell + 1) - (\mathrm{peb}_{D,k}(u)
  + \mathrm{peb}_{D, k}(v))$ strictly decreases per successful
  reversal.
\end{definition}

\begin{definition}[Blocking witness]
  \label{def:blockingWitness}
  \uses{def:isSparse}
  Given a simple graph $G : \mathrm{SimpleGraph}\,V$ and parameters
  $k, \ell$, a \emph{blocking witness} for $G$ is a finite subset
  $V' \subseteq V$ together with a proof that
  $|\edgesIn[G]{V'}| > k|V'| - \ell$. Equivalently, $G$ is not
  $(k, \ell)$-sparse on $V'$.
\end{definition}

\begin{definition}[Run-pebble-game]
  \label{def:runPebbleGame}
  \uses{def:tryAddEdge, def:blockingWitness, def:partial-orientation}
  Given a simple graph $G$ on a finite vertex set $V$ and integer
  parameters $k, \ell$ with $\ell < 2k$, the procedure
  $\mathrm{runPebbleGame}\,k\,\ell\,G$ folds $\mathrm{tryAddEdge}$
  over an arbitrary enumeration of $E(G)$ starting from the empty
  partial orientation. The result is one of:
  \begin{itemize}
    \item (success) $\mathrm{Sum.inl}\,D$ where $D$ is a partial
      orientation with $\underline{D} = E(G)$ satisfying the four
      invariants of \cref{lem:pebble-game-invariants};
    \item (failure) $\mathrm{Sum.inr}\,W$ where $W$ is a blocking
      witness for $G$ (\cref{def:blockingWitness}).
  \end{itemize}
  Termination of the fold: the number of unprocessed edges of $E(G)$
  strictly decreases per iteration. The success branch carries the
  full $D$; partial-success states (where some edges of $E(G)$ have
  been processed and inserted but the next one fails) are merged
  into the failure branch, since the L-S Theorem~8 statement on
  $G$'s sparsity uses only the existence/non-existence of $D$ with
  $\underline{D} = E(G)$.
\end{definition}

\subsection{Soundness}
\label{sec:pebble-game-soundness}

The success branch of $\mathrm{runPebbleGame}$ is automatically a
witness of sparsity, by Invariant~(4) applied to $\underline{D} =
E(G)$. This is the easy half of Lee--Streinu Theorem 8
(corresponding to L-S Corollary 12).

\begin{theorem}[Soundness of the pebble game]
  \label{thm:pebble-game-soundness}
  \uses{def:runPebbleGame, lem:pebble-game-invariants, def:isSparse}
  Let $V$ be finite, $G : \mathrm{SimpleGraph}\,V$, and $k, \ell$
  integers with $\ell < 2k$. If $\mathrm{runPebbleGame}\,k\,\ell\,G
  = \mathrm{Sum.inl}\,D$, then $G$ is $(k, \ell)$-sparse.
\end{theorem}
\begin{proof}
  \uses{def:runPebbleGame, lem:pebble-game-invariants}
  By \cref{lem:pebble-game-invariants} Invariant~(4) applied to
  $D$ at termination, every subset $V' \subseteq V$ in the relevant
  size range satisfies $\mathrm{span}_D(V') \le k|V'| - \ell$.
  Since $\underline{D} = E(G)$, this is exactly $|\edgesIn[G]{V'}|
  \le k|V'| - \ell$, i.e.\ $G$ is $(k, \ell)$-sparse on $V'$.
\end{proof}

\subsection{Completeness}
\label{sec:pebble-game-completeness}

The harder direction shows that the algorithm cannot reject any
edge of a $(k, \ell)$-sparse graph. The argument routes through
Lee--Streinu Lemma~13: if the candidate edge $\{u, v\}$ is
independent (in the matroid sense), then a free pebble can be
brought to $u$ or $v$ from within $\mathrm{Reach}_D(u) \cup
\mathrm{Reach}_D(v)$ without disturbing the other endpoint's count.
Iterating produces the $\ell + 1$ pebbles required to insert.

\begin{lemma}[Independent edges yield reachable pebbles]
  \label{lem:pebble-game-independent-brings-pebble}
  \uses{def:partial-orientation, def:tryAddEdge,
    lem:pebble-game-invariants, def:isSparse}
  Let $D$ be a partial orientation on $V$ reachable as in
  \cref{lem:pebble-game-invariants}, $\{u, v\}$ a candidate edge
  not in $\underline{D}$, and $V' := \mathrm{Reach}_D(u) \cup
  \mathrm{Reach}_D(v)$. Suppose $\mathrm{fromEdgeSet}\,(\underline{D}
  \cup \{\{u, v\}\})$ is $(k, \ell)$-sparse and
  $\mathrm{peb}_{D, k}(u) + \mathrm{peb}_{D, k}(v) < \ell + 1$.
  Then there is a vertex $w \in V' \setminus \{u, v\}$ with
  $\mathrm{peb}_{D, k}(w) \ge 1$.
\end{lemma}
\begin{proof}
  \uses{lem:pebble-game-invariants, def:isSparse}
  By the definition of reachability closure, $\mathrm{out}_D(V')
  = 0$. Invariant~(3) of \cref{lem:pebble-game-invariants} gives
  $\mathrm{peb}_{D, k}(V') \ge \ell$. Independence of $\{u, v\}$
  with respect to $\underline{D}$ (the sparsity hypothesis after
  insertion) plus the failure to immediately satisfy the
  $\ge \ell + 1$ threshold strengthens this to $\mathrm{peb}_{D, k}
  (V') \ge \ell + 1$ by an $I$-component / block argument analogous
  to the Phase~7 maximal block scaffolding
  (\cref{lem:isTightOn-union-inter-of-pair}; here applied to the
  orientation's underlying graph in place of a static sparse graph).
  Subtracting $\mathrm{peb}_{D, k}(u) + \mathrm{peb}_{D, k}(v) <
  \ell + 1$ leaves at least one pebble distributed over
  $V' \setminus \{u, v\}$, supplying the required $w$. Compare
  Lee--Streinu Lemma 13.
\end{proof}

\begin{lemma}[Pebble game accepts iff edge is independent]
  \label{lem:pebble-game-tryAddEdge-iff-independent}
  \uses{def:tryAddEdge, lem:pebble-game-independent-brings-pebble,
    lem:pebble-game-invariants, def:isSparse}
  Let $D$ be a partial orientation reachable as above and $\{u, v\}$
  a candidate edge with $\{u, v\} \notin \underline{D}$. Then
  $\mathrm{tryAddEdge}\,k\,\ell\,D\,\{u, v\} = \mathrm{Sum.inl}\,D'$
  if and only if $\mathrm{fromEdgeSet}\,(\underline{D} \cup
  \{\{u, v\}\})$ is $(k, \ell)$-sparse.
\end{lemma}
\begin{proof}
  \uses{lem:pebble-game-independent-brings-pebble,
    lem:pebble-game-invariants}
  $(\Leftarrow)$: iterate
  \cref{lem:pebble-game-independent-brings-pebble} to bring pebbles
  into $\{u, v\}$ until the threshold is met; the outer-loop
  termination measure of \cref{def:tryAddEdge} bounds the number
  of iterations by $\ell + 1$, and each application preserves
  $(k, \ell)$-sparsity of $\underline{D} \cup \{\{u, v\}\}$.
  $(\Rightarrow)$: a successful insertion produces $D'$ with
  $\underline{D'} = \underline{D} \cup \{\{u, v\}\}$ and $D'$
  satisfies Invariant~(4) by \cref{lem:pebble-game-invariants}
  (which now applies to $D'$ since the insertion was performed
  under the threshold precondition). Compare Lee--Streinu Lemma 14.
\end{proof}

\begin{lemma}[Failure produces a blocking witness]
  \label{lem:pebble-game-failure-witness}
  \uses{def:tryAddEdge, def:blockingWitness,
    lem:pebble-game-invariants}
  Let $D$ be a partial orientation reachable as above, $\{u, v\}$
  a candidate edge, and $V' := \mathrm{Reach}_D(u) \cup
  \mathrm{Reach}_D(v)$. If $\mathrm{tryAddEdge}\,k\,\ell\,D\,\{u, v\}
  = \mathrm{Sum.inr}\,V'$, then $|\{e \in \underline{D} : e
  \subseteq V'\}| + 1 > k|V'| - \ell$ in $\mathrm{fromEdgeSet}\,
  (\underline{D} \cup \{\{u, v\}\})$.
\end{lemma}
\begin{proof}
  \uses{lem:pebble-game-invariants}
  At failure, both DFS searches return empty, so
  $\mathrm{out}_D(V') = 0$ and $\mathrm{peb}_{D, k}(V') \le \ell$
  (every reachable vertex from $u$ or $v$ has zero free pebbles
  except possibly $u, v$ themselves, which together hold at most
  $\ell$). Invariant~(2) gives $\mathrm{span}_D(V') = k|V'| -
  \mathrm{peb}(V') - \mathrm{out}(V') \ge k|V'| - \ell$; combined
  with Invariant~(4) ($\mathrm{span}_D(V') \le k|V'| - \ell$),
  equality holds: $\mathrm{span}_D(V') = k|V'| - \ell$ (i.e.\ $V'$
  is a \emph{block} of the orientation). Adding $\{u, v\}$ to
  $\underline{D}$ adds one edge inside $V'$, producing $|\{e :
  e \subseteq V'\}| = k|V'| - \ell + 1$.
\end{proof}

\subsection{Correctness theorem}
\label{sec:pebble-game-correctness}

The certificate-form statement of Lee--Streinu Theorem~8 in the
matroidal regime: $\mathrm{runPebbleGame}\,k\,\ell\,G$ returns
$\mathrm{Sum.inl}\,D$ exactly when $G$ is $(k, \ell)$-sparse, and
returns $\mathrm{Sum.inr}\,W$ with $W$ a genuine blocking witness
otherwise.

\begin{theorem}[Pebble game correctness, certificate form]
  \label{thm:pebble-game-correct}
  \uses{def:runPebbleGame, def:blockingWitness, def:isSparse,
    thm:pebble-game-soundness,
    lem:pebble-game-tryAddEdge-iff-independent,
    lem:pebble-game-failure-witness}
  Let $V$ be finite, $G : \mathrm{SimpleGraph}\,V$, and $k, \ell$
  integers with $\ell < 2k$. Then:
  \begin{enumerate}
    \item If $\mathrm{runPebbleGame}\,k\,\ell\,G = \mathrm{Sum.inl}\,
      D$, then $G$ is $(k, \ell)$-sparse and $\underline{D} = E(G)$.
    \item If $\mathrm{runPebbleGame}\,k\,\ell\,G = \mathrm{Sum.inr}\,
      W$, then $W$ is a blocking witness for $G$, so $G$ is not
      $(k, \ell)$-sparse.
  \end{enumerate}
  In particular, $G$ is $(k, \ell)$-sparse if and only if there
  exists $D$ with $\mathrm{runPebbleGame}\,k\,\ell\,G = \mathrm{Sum.inl}\,D$.
\end{theorem}
\begin{proof}
  \uses{thm:pebble-game-soundness,
    lem:pebble-game-tryAddEdge-iff-independent,
    lem:pebble-game-failure-witness}
  Part (1) is \cref{thm:pebble-game-soundness} together with the
  invariant $\underline{D} = E(G)$ enforced by the fold (every edge
  of $G$ is consumed). Part (2) follows from
  \cref{lem:pebble-game-failure-witness} applied at the failure
  point, together with the IH that the accepted edges before
  failure form a $(k, \ell)$-sparse subgraph $H \subseteq G$ (by
  iterated application of
  \cref{lem:pebble-game-tryAddEdge-iff-independent}); the union of
  $H$'s edges within $V'$ plus the rejected edge $\{u, v\}$
  saturates $\edgesIn[G]{V'}$ above the sparsity bound. The biconditional
  follows by case-splitting on the disjoint $\mathrm{Sum}$ output.
\end{proof}

\subsection{Matroidal-independence corollary}
\label{sec:pebble-game-matroid}

Translating the correctness theorem through the $(k, \ell)$-count
matroid identification of \cref{sec:count-matroid} gives the
algorithmic characterisation of independence in the count matroid:
the pebble game decides matroidal independence.

\begin{corollary}[Pebble-game certificate of matroidal independence]
  \label{cor:pebble-game-countMatroid-indep}
  \uses{thm:pebble-game-correct, def:countMatroid,
    thm:countMatroid-indep-iff}
  Let $V$ be finite and $k, \ell$ integers with $\ell < 2k$. For
  every $G : \mathrm{SimpleGraph}\,V$, the edge set $E(G)$ is
  independent in the $(k, \ell)$-count matroid
  $\mathrm{countMatroid}\,V\,k\,\ell$
  (\cref{def:countMatroid}) if and only if
  $\mathrm{runPebbleGame}\,k\,\ell\,G$ returns
  $\mathrm{Sum.inl}\,D$ for some partial orientation $D$.
\end{corollary}
\begin{proof}
  \uses{thm:pebble-game-correct, thm:countMatroid-indep-iff}
  By \cref{thm:countMatroid-indep-iff}, $E(G)$ is independent in
  the count matroid iff $G$ is $(k, \ell)$-sparse. By
  \cref{thm:pebble-game-correct}, the latter holds iff the pebble
  game succeeds.
\end{proof}

\emph{Special case: planar rigidity.} At $(k, \ell) = (2, 3)$,
\cref{cor:pebble-game-countMatroid-indep} composed with the
matroid identification of \cref{thm:rigidityMatroid-indep-iff-rowIndependent}
gives an algorithmic decision procedure for membership in the
planar rigidity matroid $\mathcal{R}(V)$
(\cref{def:rigidityMatroid}). Restricting further to spanning
subgraphs $H \subseteq G$ with $|E(H)| = 2|V| - 3$ decides the
Laman property (\cref{def:isLaman}). This is the Jacobs--Hendrickson
$(2, 3)$-pebble game~\cite{jacobsHendrickson1997} as the named
algorithmic specialisation of the present chapter's main theorem.
